---
title: "Analysis of 25 m data"
output:
  html_document:
    df_print: paged
---

8/17/23

Quick analysis of 25 m deployment data for meeting with JP and Kevin

Data summary: Measurements of water presence or absence along tributaries W1 and W2 in Watershed 3. Data was collected every 10 minutes, from 2023-06-06 14:00:00 to 2023-06-14 11:00:00. Presence or absence of water was determined using an absolute value threshold, when the relative conductivity measured over 40. 

```{r, warning=FALSE}
#setup
library(pacman)
p_load(tidyverse, raster, rgdal, sp, mapview,
       rgeos, spdplyr, whitebox, tmap, stars, STICr,
       lubridate, gganimate, animation, patchwork)

#read in previously formatted data
data <- read.csv("25mformatted.csv")

data$date <- day(data$datetime)


w3_streams <- readOGR(dsn = "./Map files 6_7_23", layer = "W3_stream")
w3_streams <- as(w3_streams, 'sf')



sample <-filter(data, datetime == "2023-06-08 14:40:00")
ggplot(sample)+
  geom_point(aes(x = coords.x1, y = coords.x2, color = wetdry), size = 2)+
  scale_color_manual(values = c("salmon3", "#377EB8"),
                     breaks = c("dry", "wet"),
                     name = "STIC Sensors")+
  geom_sf(data = w3_streams, aes(geometry = geometry, lty = StrType))+
  scale_linetype_manual(values = c(1,2,3),
                        breaks = c("P", "I", "E"),
                        name = "Stream Type")+
  #labs(title = "Wettest network, all sensors")+
  #theme_classic()+
  theme(panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_blank()#remove y axis ticks
  )
```

```{r}
#6/21/23 adding discharge and precip data
#read in data
q_raw <- read.csv("ws3_Q_6_21_2023.csv", skip = 2)
p_raw <- read.csv("ws3_P_6_21_2023.csv", skip = 3)
#remove dates outside of the extent of STIC deployment
q_raw$date.and.time <- mdy_hm(q_raw$date.and.time)
p_raw$date.time <- mdy_hm(p_raw$date.time)


q_date <- filter(q_raw, date.and.time >= "2023-06-06 14:00:00") %>% 
  filter(date.and.time <= "2023-06-14 11:00:00")

p_date <- filter(p_raw, date.time >= "2023-06-06 14:00:00") %>% 
  filter(date.time <= "2023-06-14 11:00:00")
#convert inches to mm
p_date$mm <- p_date$inches * 25.4
p_date$day <- date(p_date$date.time)

p2 <- p_date %>% 
  group_by(day) %>% 
  summarise(daily_mmhr = sum(mm)/24)



ggplot(p2, aes(x = day, y = daily_mmhr))+
  geom_bar(stat = "identity")+
  theme_classic()+
  #transition_time(day) +
  labs(x = "", y = "Precip (mm/hr)", title = "Precipitation from June 6 to June 14")

ggplot(q_date, aes(x = date.and.time, y = mm.hr))+
  geom_line()+
  theme_classic()+
  labs(x = "Date", y = "Discharge (mm/hr)", title = "Discharge from June 6 to June 14")
q_date2 <- q_date
q_date2$date.and.time <- as.Date(q_date$date.and.time)

ggplot()+
  geom_bar(data = p2, aes(x = day, y = daily_mmhr), stat = "identity")+
  geom_line(data = q_date2, aes(x = date.and.time, y = mm.hr))+
  theme_classic()+
  #transition_time(day) +
  labs(x = "", y = "mm/hr", title = "Discharge and Precip from June 6 to June 14")
```

```{r}

ggplot(data, 
       aes(x = datetime, 
           fill = wetdry)) + 
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("salmon3", "#377EB8"),
                     breaks = c("dry", "wet"),
                     name = "STIC Sensors")+
  #scale_fill_brewer(palette = "Set1") +

  labs(y = "Proportion", title = "Proportion of time flowing by day")

data$datetime <- ymd_hms(data$datetime)
q_date <- rename(q_date, "datetime" = "date.and.time")
data_q <- left_join(data, q_date, by = "datetime")


# ggplot(data_q, 
#        aes(x = ft.3.sec, 
#            fill = wetdry)) + 
#   geom_density(alpha = 0.4) +
#   scale_fill_manual(values = c("salmon3", "#377EB8"),
#                      breaks = c("dry", "wet"),
#                      name = "STIC Sensors")+
#   #scale_fill_brewer(palette = "Set1") +
# 
#   labs(y = "Proportion", title = "Proportion of time flowing by discharge")

#at any given time, how many are flowing? Can I plot that by discharge?
#loop to figure out when the wettest time was

#try and do what is happening below but without a for loop

times <- unique(data$datetime)

for(z in 1:length(times)){
  #filter to every 10 min increment
  onetime <- (filter(data, datetime == times[z]))
  #find the number that were wet over the total number of sensors
  proportion_wet <- length(which(onetime$wetdry == "wet"))/length(onetime$wetdry)
  #create table of results
  out <- data.frame("datetime" = times[z], "proportion_wet" = proportion_wet)
  if(z == 1) final <- out
  if(z > 1) final <- rbind(final, out)
  
}
q_date$datetime <- mdy_hm(q_date$datetime)

data_prop <- left_join(final, q_date, by = "datetime")
data_prop$day <- day(data_prop$datetime)

daily <- data_prop %>% group_by(day) %>% 
  summarize(proportion_wet_max = max(proportion_wet),
            proportion_wet_min = min(proportion_wet),
            proportion_wet_mean = mean(proportion_wet))
ggplot(daily)+
  geom_bar


ggplot(daily, 
       aes(x = day, 
           fill = proportion_wet_max )) + 
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("salmon3", "#377EB8"),
                     breaks = c("dry", "wet"),
                     name = "STIC Sensors")+
  #scale_fill_brewer(palette = "Set1") +

  labs(y = "Proportion", title = "Proportion of time flowing by day")
#plot(x = data_prop$ft.3.sec, y = data_prop$proportion_wet, type = 'l')
```


```{r}
#analysis based on changing the spacing
every100 <- filter(data, dist %in% c(0, 100, 200, 300, 400))
every50 <- filter(data, dist %in% c(0, 50 ,100, 150, 200, 250, 300, 350, 400, 450, 461))
data$spacing <- 25
every100$spacing <- 100
every50$spacing <- 50

allspaces <- rbind(data, every50)
allspaces <- rbind(allspaces, every100)

#one way to look at how the proportion of sensors that were flowing changed with differences in spatial resolution
ggplot(allspaces, 
       aes(x = date, 
           fill = wetdry)) + 
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("salmon3", "#377EB8"),
                     breaks = c("dry", "wet"),
                     name = "STIC Sensors")+
  facet_grid(cols = vars(spacing))+
  labs(y = "Proportion")+
  theme_classic()


#figuring out proportion of sensors that were flowing for each spacing
#data_prop is for the baseline scenario, every 25 m 
  onetime <- (filter(data, datetime == times[1000]))
number <- length(onetime$wetdry)


#every 50 m
# for(z in 1:length(times)){
#   #filter to every 10 min increment
#   onetime <- (filter(every50, datetime == times[z]))
#   #find the number that were wet over the total number of sensors
#   proportion_wet <- length(which(onetime$wetdry == "wet"))/length(onetime$wetdry)
#   #create table of results
#   out <- data.frame("datetime" = times[z], "proportion_wet" = proportion_wet)
#   if(z == 1) final <- out
#   if(z > 1) final <- rbind(final, out)
#   
# }
# data_prop_every50 <- left_join(final, q_date, by = "datetime")
# 
# #every100
# for(z in 1:length(times)){
#   #filter to every 10 min increment
#   onetime <- (filter(every100, datetime == times[z]))
#   #find the number that were wet over the total number of sensors
#   proportion_wet <- length(which(onetime$wetdry == "wet"))/length(onetime$wetdry)
#   #create table of results
#   out <- data.frame("datetime" = times[z], "proportion_wet" = proportion_wet)
#   if(z == 1) final <- out
#   if(z > 1) final <- rbind(final, out)
#   
# }
# data_prop_every100 <- left_join(final, q_date, by = "datetime")
```
```{r}
#fewer_times <- alldat[seq(1, nrow(alldat), 6), ]
#for time and spatial sampling- plot the residuals through time, and plot the residuals between the full resolution and coarser resolutions by precipitation or discharge

#also could plot by distance from outlet

```


Possible improvements: Classifying water using dampening of temperature signals in addition to conductivity. 