---
title: "Analysis of 25 m data"
output:
  html_document:
    df_print: paged
---

8/17/23

Quick analysis of 25 m deployment data for meeting with JP and Kevin

Data summary: Measurements of water presence or absence along tributaries W1 and W2 in Watershed 3. Data was collected every 10 minutes, from 2023-06-06 14:00:00 to 2023-06-14 11:00:00. Presence or absence of water was determined using an absolute value threshold, when the relative conductivity measured over 40. 

```{r, warning=FALSE}
#setup
library(pacman)
library(data.table)
p_load(tidyverse, raster, rgdal, sp, mapview,
       rgeos, spdplyr, tmap, stars, STICr,
       lubridate, gganimate, animation, patchwork)

#read in previously formatted data
data <- read.csv("25mformatted.csv")
data$datetime <- ymd_hms(data$datetime)
data$datetime <- with_tz(data$datetime, "US/Eastern")



data$date <- day(data$datetime)
#adding binary column
data$wetdry_binary <- 0
data$wetdry_binary[data$wetdry == "wet"] <- 1

w3_streams <- readOGR(dsn = "./Map files 6_7_23", layer = "W3_stream")
w3_streams <- as(w3_streams, 'sf')




sample <- filter(data, datetime >= "2023-06-13 16:10:00") %>% 
  filter(datetime <= "2023-06-13 17:00:00")
ggplot(sample)+
  #geom_point(aes(x = coords.x1, y = coords.x2, color = wetdry), size = 2)+
  scale_color_manual(values = c("salmon3", "#377EB8"),
                     breaks = c("dry", "wet"),
                     name = "STIC Sensors")+
  geom_text(aes(x = coords.x1, y = coords.x2, color = wetdry, label = dist), size = 3)+
    facet_wrap(~datetime)+
  geom_sf(data = w3_streams, aes(geometry = geometry, lty = StrType))+
  scale_linetype_manual(values = c(1,2,3),
                        breaks = c("P", "I", "E"),
                        name = "Stream Type")+
  #labs(title = "Wettest network, all sensors")+
  #theme_classic()+
  theme(panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_blank()#remove y axis ticks
  )
```

```{r}
#6/21/23 adding discharge and precip data
#read in data
q_raw <- read.csv("ws3_Q_6_21_2023.csv", skip = 2)
p_raw <- read.csv("ws3_P_6_21_2023.csv", skip = 3)
#remove dates outside of the extent of STIC deployment
q_raw$date.and.time <- mdy_hm(q_raw$date.and.time)
p_raw$date.time <- mdy_hm(p_raw$date.time)


q_date <- filter(q_raw, date.and.time >= "2023-06-06 14:00:00") %>% 
  filter(date.and.time <= "2023-06-14 11:00:00") %>% 
  rename("datetime" = "date.and.time")

p_date <- filter(p_raw, date.time >= "2023-06-06 14:00:00") %>% 
  filter(date.time <= "2023-06-14 11:00:00")
#convert inches to mm
p_date$mm <- p_date$inches * 25.4
p_date$day <- date(p_date$date.time)

p2 <- p_date %>% 
  group_by(day) %>% 
  summarise(daily_mmhr = sum(mm)/24)



ggplot(p2, aes(x = day, y = daily_mmhr))+
  geom_bar(stat = "identity")+
  theme_classic()+
  #transition_time(day) +
  labs(x = "", y = "Precip (mm/hr)", title = "Precipitation from June 6 to June 14")

ggplot(q_date, aes(x = datetime, y = mm.hr))+
  geom_line()+
  theme_classic()+
  labs(x = "Date", y = "Discharge (mm/hr)", title = "Discharge from June 6 to June 14")
q_date2 <- q_date
q_date2$datetime <- as.Date(q_date$date.and.time)


ggplot()+
  geom_bar(data = p2, aes(x = day, y = daily_mmhr), stat = "identity")+
  geom_line(data = q_date2, aes(x = datetime, y = mm.hr))+
  theme_classic()+
  #transition_time(day) +
  labs(x = "", y = "mm/hr", title = "Discharge and Precip from June 6 to June 14")
```

```{r}

ggplot(data, 
       aes(x = datetime, 
           fill = wetdry)) + 
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("salmon3", "#377EB8"),
                     breaks = c("dry", "wet"),
                     name = "STIC Sensors")+
  #scale_fill_brewer(palette = "Set1") +

  labs(y = "Proportion", title = "Proportion of time flowing by day")

data_q <- left_join(data, q_date, by = "datetime")




#at any given time, how many are flowing? Can I plot that by discharge?




#proportion of sensors that were flowing
daily2 <- data %>% group_by(datetime) %>% 
  mutate(proportion_wet = sum(wetdry_binary)/length(wetdry_binary)) %>% 
  dplyr::select(datetime, proportion_wet)

daily2$day <- day(daily2$datetime)

daily3 <- unique(daily2)

daily_stats <- daily3 %>% group_by(day) %>% 
  summarize(proportion_wet_max = max(proportion_wet),
            proportion_wet_min = min(proportion_wet),
            proportion_wet_mean = mean(proportion_wet))
#successfully made for loop obsolete

#data_prop <- left_join(final, q_date, by = "datetime")

#good blue color = #377EB8

#this plot shows the proportions of time flowing statistics at the daily time scale
daily_stats_tall <- gather(daily_stats, "stat", "value", -day)
ggplot(daily_stats_tall, 
       aes(x = day, 
           y = value)) + 
  geom_line(aes(color = stat))+
  theme_classic()+
  labs(y = "Proportion", title = "Proportion of time flowing by day")

#do the same thing but for hourly and every 10 minutes; these are the null hypothesis or control scenario

#removing the last row, because it makes the dataframes even numbered in length- they were odd (1135 rows), so when filtering out every other row, it would set some sensors to begin at 11L10 and others to begin at 11:00. Fixed now, with all of the last dates/rows removed
even_data <- filter(data, datetime != "2023-06-14 11:00:00")
  

#remove every other datetime, makes the measurements 20 minutes apart
  
mins20 <- even_data[seq(1, nrow(even_data), 2), ]


mins20_prop <- mins20 %>% group_by(datetime) %>% 
  mutate(proportion_wet = sum(wetdry_binary)/length(wetdry_binary)) %>% 
  dplyr::select(datetime, proportion_wet)

#unique to get rid of duplicate records
mins20_prop <- unique(mins20_prop)

daily_stats$day_date <- c("2023-06-06",
                          "2023-06-07",
                          "2023-06-08",
                          "2023-06-09",
                          "2023-06-10",
                          "2023-06-11",
                          "2023-06-12",
                          "2023-06-13",
                          "2023-06-14")

daily_stats$day_date <- as.POSIXct(ymd(daily_stats$day_date))

#plot showing daily stats, and every actual measurement

#make plot showing what would happen if we had one measurement every 120 minutes
mins120 <- even_data[seq(1, nrow(even_data), 12), ]


mins120_prop <- mins120 %>% group_by(datetime) %>% 
  mutate(proportion_wet = sum(wetdry_binary)/length(wetdry_binary)) %>% 
  dplyr::select(datetime, proportion_wet)

#unique to get rid of duplicate records
mins120_prop <- unique(mins120_prop)

ggplot() + 
  geom_line(data = daily_stats,
            aes(x = day_date, 
                y = proportion_wet_mean), color = "black")+
  geom_line(data = mins20_prop,
            aes(x = datetime,
                y = proportion_wet), color = "blue")
  # geom_line(data = mins120_prop,
  #           aes(x = datetime,
  #               y = proportion_wet), color = "red")


```
  
```{r}
#bootstrapping to subset observation points
sensors <- unique(data$ID)
n <- 20

less_sense <- function(n, iterations){
  #n = number of sensors, from total 34
  for(i in 1:iterations){
    sub <- sample(sensors, n)
    filt <- filter(data, ID %in% sub)

    daily_sub <- filt %>% group_by(datetime) %>% 
    mutate(proportion_wet = sum(wetdry_binary)/length(wetdry_binary)) %>% 
    dplyr::select(datetime, proportion_wet)

    daily_sub$day <- day(daily_sub$datetime)

    daily_sub_un <- unique(daily_sub)

    daily_sub_stats <- daily_sub_un %>% group_by(day) %>% 
      summarize(proportion_wet_max = max(proportion_wet),
              proportion_wet_min = min(proportion_wet),
              proportion_wet_mean = mean(proportion_wet))

    daily_sub_stats$day_date <- c("2023-06-06",
                          "2023-06-07",
                          "2023-06-08",
                          "2023-06-09",
                          "2023-06-10",
                          "2023-06-11",
                          "2023-06-12",
                          "2023-06-13",
                          "2023-06-14")

    daily_sub_stats$day_date <- as.POSIXct(ymd(daily_sub_stats$day_date))
  
    #create table of results
    out <- dplyr::select(daily_sub_stats, -day) %>% 
      gather(key = "stat", value = "value", -day_date)
    out$number_set <- n
    out$sensors <- paste(sort(sub), collapse=', ' )
  
  
    if(i == 1) final <- out
    if(i > 1) final <- rbind(final, out)
  }
  return(final)
}

#subset of 5 sensors
n5 <- less_sense(5, 100)
final2 <- filter(n5, stat == "proportion_wet_mean")
ggplot() + 
  geom_boxplot(data = final2,
            aes(x = day_date, 
                y = value,
                group = day_date), color = "black")+
  geom_line(data = daily_stats,
            aes(x = day_date, 
                y = proportion_wet_mean), color = "black")+
  labs(x = "Date", y = "Proportion flowing", title = "5 random sensors bootstrap, n = 100")+
  lims(y = c(0, 1))+
  theme_classic()
#subset of 10 sensors
n1 <- less_sense(1, 100)
#n5 <- less_sense(5, 100)
n10 <- less_sense(10, 100)
n20 <- less_sense(20, 100)
n30 <- less_sense(30, 100)

all1 <- rbind(n1, n5)
all2 <- rbind(n5, n10)
all3 <- rbind(all2, n20)
all4 <- rbind(all3, n30)

#format daily stats correctly
actual <- filter(daily_stats, day_date == "2023-06-13 20:00:00")
value <- actual$proportion_wet_mean

actual2 <- filter(daily_stats, day_date == "2023-06-05 20:00:00")
value2 <- actual2$proportion_wet_mean


# This for loop will work, but will take about 20 mins
# for(z in 1:34){
#   raw <- less_sense(z, 100)
#   
#   if(z == 1) final <- raw
#   if(z > 1) final <- rbind(final, raw)
# }

date_wet <- filter(all4, day_date == "2023-06-13 20:00:00") %>% 
    filter(stat == "proportion_wet_mean")

date_dry <- filter(all4, day_date == "2023-06-05 20:00:00") %>% 
    filter(stat == "proportion_wet_mean")
#plot that will show a boxplot for each randomly chosen number of sensors for June 14, the wettest day
ggplot() + 
  geom_boxplot(data = date_wet,
            aes(x = number_set, 
                y = value,
                group = number_set), color = "black")+
  geom_hline(yintercept = value, linetype = "dashed")+
  labs(x = "Number of randomly chosen sensors", y = "Proportion flowing", title = "Random sensors bootstrap on the wettest day")+
  lims(y = c(0, 1))+
  theme_classic()

ggplot() + 
  geom_boxplot(data = date_dry,
            aes(x = number_set, 
                y = value,
                group = number_set), color = "black")+
    geom_hline(yintercept = value2, linetype = "dashed")+

  labs(x = "Number of randomly chosen sensors", y = "Proportion flowing", title = "Random sensors bootstrap on the driest day")+
  lims(y = c(0, 1))+
  theme_classic()

#when I write this in a for loop, need to save the output somehow
#out put needs to have the proportion wet, date, number of sensors used, and list the sensors used
```

```{r}
#find all unique combinations of wet and dry sensors
hush <- dplyr::select(data, datetime, wetdry_binary, ID) 
hush2 <- pivot_wider(hush, names_from = ID, values_from = wetdry_binary)

xy.list <- as.list(as.data.frame(t(hush2[,-1])))
length(xy.list)
length(unique(data$datetime))
#unique combinations of flow
length(unique(xy.list))



```


```{r}
#analysis based on changing the spacing
every100 <- filter(data, dist %in% c(0, 100, 200, 300, 400))
every50 <- filter(data, dist %in% c(0, 50 ,100, 150, 200, 250, 300, 350, 400, 450, 461))
data$spacing <- 25
every100$spacing <- 100
every50$spacing <- 50

allspaces <- rbind(data, every50)
allspaces <- rbind(allspaces, every100)

#one way to look at how the proportion of sensors that were flowing changed with differences in spatial resolution
ggplot(allspaces, 
       aes(x = date, 
           fill = wetdry)) + 
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("salmon3", "#377EB8"),
                     breaks = c("dry", "wet"),
                     name = "STIC Sensors")+
  facet_grid(cols = vars(spacing))+
  labs(y = "Proportion")+
  theme_classic()


#figuring out proportion of sensors that were flowing for each spacing
#data_prop is for the baseline scenario, every 25 m 
  onetime <- (filter(data, datetime == times[1000]))
number <- length(onetime$wetdry)


#every 50 m
# for(z in 1:length(times)){
#   #filter to every 10 min increment
#   onetime <- (filter(every50, datetime == times[z]))
#   #find the number that were wet over the total number of sensors
#   proportion_wet <- length(which(onetime$wetdry == "wet"))/length(onetime$wetdry)
#   #create table of results
#   out <- data.frame("datetime" = times[z], "proportion_wet" = proportion_wet)
#   if(z == 1) final <- out
#   if(z > 1) final <- rbind(final, out)
#   
# }
# data_prop_every50 <- left_join(final, q_date, by = "datetime")
# 
# #every100
# for(z in 1:length(times)){
#   #filter to every 10 min increment
#   onetime <- (filter(every100, datetime == times[z]))
#   #find the number that were wet over the total number of sensors
#   proportion_wet <- length(which(onetime$wetdry == "wet"))/length(onetime$wetdry)
#   #create table of results
#   out <- data.frame("datetime" = times[z], "proportion_wet" = proportion_wet)
#   if(z == 1) final <- out
#   if(z > 1) final <- rbind(final, out)
#   
# }
# data_prop_every100 <- left_join(final, q_date, by = "datetime")
```
```{r}
#fewer_times <- alldat[seq(1, nrow(alldat), 6), ]
#for time and spatial sampling- plot the residuals through time, and plot the residuals between the full resolution and coarser resolutions by precipitation or discharge

#also could plot by distance from outlet

```


Possible improvements: Classifying water using dampening of temperature signals in addition to conductivity. 

```{r}
#HB air temperature record goes to May 15
sample <- filter(data, datetime >= "2023-06-12 16:10:00") %>% 
  filter(datetime <= "2023-06-13 17:00:00") 

#find where wet starts and stops
test <- filter(data, ID == "w2_175")
test$start_stop <- c(NA, diff(test$wetdry_binary, lag = 1))
ends <- dplyr::select(test, datetime, ID, wetdry_binary, start_stop)

ends_start <- filter(ends, start_stop == 1)
ends_stop <- filter(ends, start_stop == -1)

ends_final <- data.frame("start" = ends_start$datetime,
                         "stop" = c(ends_stop$datetime, NA),
                         
                         "ID" = "w1_0")

sensors <- unique(data$ID)

for(i in 1:length(sensors)){
  test <- filter(data, ID == sensors[i])
  if(max(test$wetdry_binary) == 0) next
  
  test$start_stop <- c(NA, diff(test$wetdry_binary, lag = 1))
  ends <- dplyr::select(test, datetime, ID, wetdry_binary, start_stop)

  ends_start <- filter(ends, start_stop == 1)
  ends_stop <- filter(ends, start_stop == -1)
  if(length(ends_start$datetime) > length(ends_stop$datetime)) {
    ends_final <- data.frame("start" = ends_start$datetime,
                           "stop" = c(ends_stop$datetime, "2023-06-14 11:00:00"),
                           "ID" = sensors[i])
  } else {
    ends_final <- data.frame("start" = ends_start$datetime,
                           "stop" = ends_stop$datetime,
                           "ID" = sensors[i])
  }
  
  
 
  
  if(i == 1) alldat <- ends_final
  if(i > 1) alldat <- rbind(alldat, ends_final)
}

ggplot() + 
  geom_rect(data = alldat,
            mapping = aes(xmin = start,
                          xmax = stop,
                          ymin = - Inf,
                          ymax = Inf),
              fill = "blue",
              alpha = 0.5) +
  geom_line(data = data, 
            mapping = aes(x = datetime, 
            y = tempC,
            color = stream))+
  theme_classic()+
  facet_wrap(~ID)+
  labs(y = "Temp (C)", 
       x = "",
       title = "Temperature versus threshold determined flow")



ggplot() + 
  geom_rect(data = alldat,
            mapping = aes(xmin = start,
                          xmax = stop,
                          ymin = - Inf,
                          ymax = Inf),
              fill = "blue",
              alpha = 0.5) +
  geom_line(data = data, 
            mapping = aes(x = datetime, 
            y = condUncal  ,
            color = stream))+
  geom_vline(xintercept = ymd_hms("2023-06-07 13:30:00"))+
  geom_vline(xintercept = ymd_hms("2023-06-13 15:30:00"))+
  theme_classic()+
  facet_wrap(~ID, scales = "free_y")+
  labs(y = "Conductivity (us)",
       x = "",
       title = "Conductivity versus threshold determined flow")

```

