---
title: "Final figures, sequential wetting and drying publication"
format:
  html:
    code-fold: true
    fig-width: 9
    fig-height: 7
editor_options:
  chunk_output_type: console
---

Publication Figures 9/20/26
- Fig 1: Maps of the three watersheds (flow persistence) with an inset map of 
         United states and NH
- Fig 2: Conceptual diagram of how test sequences were defined (made by hand in illustrator).
- Fig 3:
- Fig 4: 
- Fig 5: 
- Fig 6: 
- Fig 7:
- Fig 8: 
- Table 1:
- Table 2:
- Table 3: 
                                            (RQ1 + RQ2)
  Fig 3  Ordered deviation heatmap: consensus column + one column per event
         (event-based sequences only; no FP/TWI/BFS). Tan = node missing from
         that event. One row of panels per watershed.                   (RQ2)
  Fig 4  Wetting vs. drying sequence as a rank-rank plot (x = wetting, y =
         drying, y reversed so the dashed line = drying is the reverse of
         wetting). One panel per watershed.                             (RQ2)
  Fig 5  Conceptual diagram — NOT generated here.
  Fig 6  Rank-rank agreement of Flow Persistence and TWI vs. the event-based
         sequence (no BFS), high-contrast diverging colour. Row per shed. (RQ3)
  Fig 7  W3 hydrograph (cropped to the two STIC seasons): points only during
         events, coloured by whether Flow Persistence or TWI is more accurate
         at that step.                                                   (RQ3)
  Fig 8  Maps of per-node Flow Persistence and TWI model accuracy over the whole
         record, all three watersheds (Fig 1 style).                     (RQ3)

# Setup and inputs

```{r setup}
#| message: false
#| warning: false
library(pacman)
p_load(tidyverse, terra, tidyterra, whitebox, scales, wesanderson, caret,
       ggnewscale, sf, patchwork, ggspatial, zoo, igraph, TSP, knitr,
       hydroEvents, fuzzyjoin, rempsyc, ggpubr, maps)

# Colour palettes used throughout ---------------------------------------------
binary_colors  <- c("0" = "#DB995A", "1" = "#586BA4")     # dry / wet
seq_colors     <- c(
  "Flow Persistence"          = "#7ebdc2",
  "Topographic Wetness Index" = "#f3dfa2",
  "BFS"                       = "#bb4430",
  "Event based sequence"      = "#231f20"
)
dir_colors     <- c("wetting" = "#93C2F1", "drying" = "#FFA400")
shed_colors    <- c("FB" = "#397367", "W3" = "#FFD166", "ZZ" = "#7E6B8F")

# Read final-format STIC data -------------------------------------------------
data_23 <- read_csv("./DataForMary/HB_stic.csv")          # summer 2023
data_24 <- read_csv("./summer2024/STICS2024.csv")         # summer 2024

# binary wet (1) / dry (0) ----------------------------------------------------
data_23$binary <- ifelse(data_23$wetdry == "dry", 0, 1)
data_24$binary <- ifelse(data_24$wetdry == "dry", 0, 1)

data_23$mins <- minute(data_23$datetime)
data_24$mins <- minute(data_24$datetime)
```

```{r prepare-inputs}
#| message: false
#| warning: false
# Sensors deployed in BOTH campaigns, per watershed ---------------------------
W3_IDs <- intersect(unique(filter(data_24, wshed == "W3")$number),
                    unique(filter(data_23, wshed == "W3")$ID))
FB_IDs <- intersect(unique(filter(data_24, wshed == "FB")$number),
                    unique(filter(data_23, wshed == "FB")$ID))
ZZ_IDs <- intersect(unique(filter(data_24, wshed == "ZZ")$number),
                    unique(filter(data_23, wshed == "ZZ")$ID))

# 30-minute records, common columns ------------------------------------------
bind24 <- data_24 %>%
  dplyr::select(datetime, number, wshed, binary, mins) %>%
  dplyr::rename(ID = number) %>%
  filter(mins %in% c(0, 30))

bind23 <- data_23 %>%
  filter(ID %in% intersect(bind24$ID, data_23$ID)) %>%
  dplyr::select(datetime, ID, wshed, binary, mins) %>%
  filter(mins %in% c(0, 30))

# Per-watershed inputs (only sensors common to both years) --------------------
input_w3 <- rbind(bind23, bind24) %>% filter(wshed == "W3", ID %in% W3_IDs) %>% drop_na()
input_fb <- rbind(bind23, bind24) %>% filter(wshed == "FB", ID %in% FB_IDs)
input_zz <- rbind(bind23, bind24) %>% filter(wshed == "ZZ", ID %in% ZZ_IDs)
input_all <- rbind(input_w3, input_fb, input_zz)
```

## Functions

```{r helper-functions}
`%not_in%` <- purrr::negate(`%in%`)

# up/down  <->  ID/sequence converters ----------------------------------------
convert_to_IDseq <- function(input){
  tibble(ID = c(input$down[1], input$up),
         sequence = seq_len(length(input$up) + 1))
}

convert_to_updown <- function(input){
  input %>% mutate(up = ID, down = lag(up)) %>% select(up, down) %>% drop_na()
}

convert_to_IDseq2 <- function(input){
  # first row is the most-upstream / last-in-sequence pair
  IDs <- c(input$up[1], input$down)
  tibble(ID = as.numeric(substr(IDs, 3, 4)),
         sequence = seq_len(length(input$up) + 1))
}
```

```{r support-functions}
# ---- Count 2-step transition motifs for an up/down pair ---------------------
# "supports" are the four motifs consistent with parent->child ordering.
calc_support_combos <- function(up, down, input){
  output <- data.frame(up, down)
  no_dupes <- input %>%
    select(up, down, datetime) %>%
    filter(row_number() == 1 |
             !apply(select(., up, down) == lag(select(., up, down)), 1, all)) %>%
    drop_na()
  check <- nrow(no_dupes)
  if (check <= 2){
    return(data.frame(Sequence = NA, Frequency = NA, up = up, down = down))
  }
  windows <- rollapply(select(no_dupes, -datetime), width = 2, by.column = FALSE,
                       FUN = function(x) paste(as.vector(t(x)), collapse = ""))
  sequence_df <- as.data.frame(sort(table(windows), decreasing = TRUE),
                               stringsAsFactors = FALSE)
  colnames(sequence_df) <- c("Sequence", "Frequency")
  sequence_df$up <- up; sequence_df$down <- down
  sequence_df
}

# ---- Break a record into gap-free groups, then count motifs -----------------
iterate_groups_combos <- function(up, down, input, timestep){
  input$group <- cumsum(c(TRUE, diff(input$datetime) != timestep))
  out <- NULL
  for (u in unique(input$group)){
    filtered_input <- input %>% filter(group == u)
    out <- rbind(out, calc_support_combos(up, down, filtered_input))
  }
  out
}

# ---- Iterate over a routes table (a candidate sequence) ---------------------
calculate_hierarchy_combos <- function(routes, input, timestep){
  alldat <- NULL
  for (x in seq_along(routes$up)){
    alldat <- rbind(alldat,
                    iterate_groups_combos(paste0("r_", routes$up[x]),
                                          paste0("r_", routes$down[x]),
                                          input, timestep))
  }
  alldat %>% drop_na() %>%
    group_by(up, down, Sequence) %>%
    summarise(Frequency = sum(Frequency), .groups = "drop")
}

# ---- Build the 30-min wide input for a watershed and run the above ----------
fantastic_four_combos <- function(routes, shed){
  input <- input_all %>%
    filter(wshed == shed, mins %in% c(0, 30)) %>%
    select(datetime, binary, ID) %>%
    mutate(ID = paste0("r_", ID)) %>%
    pivot_wider(names_from = ID, values_from = binary)
  out <- calculate_hierarchy_combos(routes, input, minutes(30))
  out$timescale <- "30mins"; out$shed <- shed
  out
}

# ---- Proportion of state changes that follow the parent->child order --------
# (UNCHANGED behaviour — this is the version the model/figure pipeline relies on.)
calc_props <- function(routes, shed){
  full_combos <- fantastic_four_combos(routes, shed)
  supports <- c("0001", "0111", "1101", "0100")
  total_state_changes <- full_combos %>%
    filter(Sequence != "0011", Sequence != "1100") %>%
    group_by(up, down, timescale, shed) %>%
    summarise(totals = sum(Frequency), .groups = "drop")
  hierarchical_changes <- full_combos %>%
    filter(Sequence != "0011", Sequence != "1100", Sequence %in% supports) %>%
    group_by(up, down, timescale, shed) %>%
    summarise(hierarchical = sum(Frequency), .groups = "drop")
  total_state_changes %>%
    left_join(hierarchical_changes, by = c("up", "down", "shed", "timescale")) %>%
    mutate(prop = hierarchical / totals) %>%
    mutate(across(everything(), ~replace(., is.na(.), 0)))
}
```

```{r consensus-function}
#| message: false
#| warning: false
# Consensus ("average") activation order from a set of event sequences.
# Uses pairwise precedence: P[i,j] = fraction of events where i precedes j.
average_activation_sequence <- function(data,
                                        sensor_col = "sensor_id",
                                        event_col  = "event_id") {
  data <- data[order(data[[event_col]]), ]
  seqs <- split(data[[sensor_col]], data[[event_col]])
  all_sensors <- sort(unique(unlist(seqs)))
  n <- length(all_sensors)
  P <- matrix(NA, n, n, dimnames = list(all_sensors, all_sensors))
  for (s1 in all_sensors) for (s2 in all_sensors) {
    if (s1 == s2) next
    total <- 0; wins <- 0
    for (sq in seqs) if (s1 %in% sq && s2 %in% sq) {
      total <- total + 1
      if (match(s1, sq) < match(s2, sq)) wins <- wins + 1
    }
    if (total > 0) P[as.character(s1), as.character(s2)] <- wins / total
  }
  score <- rowMeans(P, na.rm = TRUE)
  list(precedence_matrix = P,
       precedence_scores = score,
       consensus_sequence = names(sort(score, decreasing = TRUE)))
}
```

```{r model-functions}
# Botter-style model: given the NUMBER of flowing nodes at each timestep and a
# fixed activation order, activate that many nodes in order; compare to obs.
calc_model_result <- function(input_sequence, shed){
  input <- switch(shed, "W3" = input_w3, "FB" = input_fb, "ZZ" = input_zz)

  activation_schedule <- input %>%
    group_by(datetime) %>%
    select(-c(wshed, mins)) %>%
    summarise(number_flowing = sum(binary == 1),
              deployed = list(unique(ID)), .groups = "drop")

  full_node_sequence <- input_sequence
  '%!in%' <- function(x, y) !('%in%'(x, y))

  generate_activation_row <- function(datetime, n_active, available_nodes){
    ordered_available <- full_node_sequence[full_node_sequence %in% available_nodes]
    activated_nodes   <- head(ordered_available, n_active)
    tibble(datetime = datetime, ID = full_node_sequence, binary = 0) %>%
      mutate(binary = case_when(ID %in% activated_nodes      ~ 1,
                                ID %!in% ordered_available    ~ NA_real_,
                                TRUE                          ~ 0))
  }

  Map(generate_activation_row,
      datetime        = activation_schedule$datetime,
      n_active        = activation_schedule$number_flowing,
      available_nodes = activation_schedule$deployed) %>%
    dplyr::bind_rows() %>%
    dplyr::rename(pred_out = binary)
}

# correct / omission / commission label ---------------------------------------
get_eval_label <- function(true, pred){
  if (is.na(true) || is.na(pred)) return(NA_character_)
  if (true == pred) return("correct")
  if (true == 1 && pred == 0) return("omission")
  "commission"
}

# Accuracy metrics for one sequence (whole record) ----------------------------
produce_metrics <- function(chain, shed, method){
  input <- switch(shed, "W3" = input_w3, "FB" = input_fb, "ZZ" = input_zz)
  model_result <- calc_model_result(chain$ID, shed)
  comparison <- input %>%
    filter(mins %in% c(0, 30)) %>%
    left_join(model_result, by = c("datetime", "ID")) %>%
    select(-wshed, -mins) %>%
    drop_na()
  cm <- confusionMatrix(factor(comparison$pred_out, levels = c(0, 1)),
                        factor(comparison$binary,   levels = c(0, 1)),
                        positive = "1")
  tibble(Accuracy    = cm$overall["Accuracy"],
         Kappa       = cm$overall["Kappa"],
         Sensitivity = cm$byClass["Sensitivity"],
         Specificity = cm$byClass["Specificity"],
         shed = shed, method = method)
}
```

```{r wetting-drying-functions}
# ---------------------------------------------------------------------------
# DIRECTION-SPLIT support analysis (wetting vs. drying).
# Returns, for each up/down pair, the proportion of WETTING transitions and of
# DRYING transitions that follow the parent->child order.
# NOTE: these functions return their own object and DO NOT touch calc_props().
# ---------------------------------------------------------------------------
all_transitions <- function(up, down, input){
  output <- data.frame(up, down)
  no_dupes <- input %>%
    select(up, down, datetime) %>%
    filter(row_number() == 1 |
             !apply(select(., up, down) == lag(select(., up, down)), 1, all)) %>%
    drop_na()
  if (nrow(no_dupes) <= 2){
    output$wetting <- NA; output$drying <- NA; return(output)
  }
  windows <- rollapply(select(no_dupes, -datetime), width = 2, by.column = FALSE,
                       FUN = function(x) paste(as.vector(t(x)), collapse = ""))
  sdf <- as.data.frame(sort(table(windows), decreasing = TRUE),
                       stringsAsFactors = FALSE)
  sdf$sum_t1 <- as.numeric(substr(sdf$windows, 1, 1)) +
                as.numeric(substr(sdf$windows, 2, 2))
  sdf$sum_t2 <- as.numeric(substr(sdf$windows, 3, 3)) +
                as.numeric(substr(sdf$windows, 4, 4))
  sdf$direction <- ifelse(sdf$sum_t1 < sdf$sum_t2, "wetting", "drying")
  supports <- c("0001", "0111", "1101", "0100")
  totals <- sdf %>% group_by(direction) %>% summarise(totals = sum(Freq), .groups = "drop")
  hier   <- sdf %>% filter(windows %in% supports) %>%
    group_by(direction) %>% summarise(hierarchical = sum(Freq), .groups = "drop")
  sub <- totals %>% left_join(hier, by = "direction") %>%
    mutate(prop = hierarchical / totals,
           across(everything(), ~replace(., is.na(.), 0)))
  output$drying  <- sub$prop[sub$direction == "drying"][1]
  output$wetting <- sub$prop[sub$direction == "wetting"][1]
  output
}

iterate_groups_wd <- function(up, down, input, timestep){
  input$group <- cumsum(c(TRUE, diff(input$datetime) != timestep))
  out <- NULL
  for (u in unique(input$group)){
    out <- rbind(out, all_transitions(up, down, input %>% filter(group == u)))
  }
  out
}

calculate_hierarchy_wd <- function(routes, input, timestep){
  alldat <- NULL
  for (x in seq_along(routes$up)){
    alldat <- rbind(alldat,
                    iterate_groups_wd(paste0("r_", routes$up[x]),
                                      paste0("r_", routes$down[x]),
                                      input, timestep))
  }
  alldat %>% drop_na()
}

# Wetting/drying proportions for a sequence's routes.
# `restrict_dt` (optional) limits the analysis to a set of datetimes (e.g. events).
wetting_drying_props <- function(routes, shed, restrict_dt = NULL){
  input <- input_all %>%
    filter(wshed == shed, mins %in% c(0, 30)) %>%
    { if (!is.null(restrict_dt)) filter(., datetime %in% restrict_dt) else . } %>%
    select(datetime, binary, ID) %>%
    mutate(ID = paste0("r_", ID)) %>%
    pivot_wider(names_from = ID, values_from = binary)
  out <- calculate_hierarchy_wd(routes, input, minutes(30))
  out$shed <- shed
  out
}

# Unique datetimes that fall inside an identified event (drop baseflow). -------
event_dt <- function(event_df) event_df %>%
  filter(event_type != "baseflow", !is.na(event_id)) %>% pull(datetime) %>% unique()
```

```{r sequence-similarity-functions}
# ---------------------------------------------------------------------------
# How similar are two orderings of the same nodes? Simple string-matching /
# sequence-comparison metrics from CS and computational biology:
#   * Kendall's tau, Spearman's rho  - rank correlation of the orderings
#   * Levenshtein similarity         - 1 - (edit distance / longer length)
#   * LCS similarity                 - longest common subsequence / longer length
#   * Hamming similarity             - share of positions that match (equal len)
# Inputs `a`, `b` are ordered vectors of node IDs (1st = activates first).
# ---------------------------------------------------------------------------
lcs_length <- function(x, y){
  n <- length(x); m <- length(y)
  if (n == 0 || m == 0) return(0L)
  d <- matrix(0L, n + 1, m + 1)
  for (i in seq_len(n)) for (j in seq_len(m))
    d[i + 1, j + 1] <- if (x[i] == y[j]) d[i, j] + 1L
                        else max(d[i, j + 1], d[i + 1, j])
  d[n + 1, m + 1]
}

seq_similarity <- function(a, b){
  a <- as.character(a); b <- as.character(b)
  common <- intersect(a, b)
  ra <- match(common, a); rb <- match(common, b)
  tau <- if (length(common) > 2) suppressWarnings(cor(ra, rb, method = "kendall")) else NA_real_
  rho <- if (length(common) > 2) suppressWarnings(cor(ra, rb, method = "spearman")) else NA_real_
  # map every node to a distinct single character, then run edit distance
  uni <- union(a, b)
  key <- setNames(intToUtf8(64 + seq_along(uni), multiple = TRUE), uni)
  sa <- paste(key[a], collapse = ""); sb <- paste(key[b], collapse = "")
  lev <- as.numeric(utils::adist(sa, sb))
  lev_sim <- 1 - lev / max(nchar(sa), nchar(sb))
  lcs_sim <- lcs_length(a, b) / max(length(a), length(b))
  ham_sim <- if (length(a) == length(b)) mean(a == b) else NA_real_
  tibble(kendall_tau     = round(tau, 3),
         spearman        = round(rho, 3),
         levenshtein_sim = round(lev_sim, 3),
         lcs_sim         = round(lcs_sim, 3),
         hamming_sim     = round(ham_sim, 3),
         n_common        = length(common))
}
```

## Discharge and event identification

```{r prepare-discharge-w3}
#| message: false
#| warning: false
# W3 discharge (input to event detection) -------------------------------------
q <- read_csv("HB_5minQ.csv")

start <- min(input_w3$datetime); stop <- max(input_w3$datetime)
q_23 <- q %>%
  filter(DATETIME > start & DATETIME < stop) %>%
  mutate(Q_mm_day = Discharge_ls * 0.001 * 86400 / 420000 * 1000,
         mins = minute(DATETIME))
q_23_f <- filter(q_23, mins %in% c(0, 30))

q_24 <- read_csv("w3_discharge_24.csv") %>%
  mutate(datetime = mdy_hm(datetime))
start24 <- ymd_hms("2024-05-15 00:00:00 UTC")
stop24  <- max(input_w3$datetime)
q_24 <- q_24 %>% filter(datetime > start24 & datetime < stop24) %>%
  mutate(mins = minute(datetime))
q_24_f <- filter(q_24, mins %in% c(0, 30))

# A single discharge series for plotting --------------------------------------
w3_discharge_bind <- bind_rows(
  q_23_f %>% select(DATETIME, Q_mm_day) %>% rename(datetime = DATETIME),
  q_24_f %>% select(datetime, Q_mmperday) %>% rename(Q_mm_day = Q_mmperday)
)

# date windows for the two STIC seasons ---------------------------------------
start_23 <- ymd_hms("2023-07-20 00:00:00"); stop_23 <- ymd_hms("2023-10-22 00:00:00")
q_23_plotting <- q_23_f %>% filter(DATETIME > start_23 & DATETIME < stop_23) %>%
  rename(datetime = DATETIME)

# Continuous W3 discharge restricted to the two STIC seasons (no winter gap).
# Used as the hydrograph line for Figs 2, 6, 7.
w3_q_seasonal <- w3_discharge_bind %>%
  filter((datetime >= start_23 & datetime <= stop_23) |
         (datetime >= start24 & datetime <= stop24))

# Hourly precipitation for both STIC seasons (Fig 2 event panels). -------------
# NB: aggregate with floor_date, NOT group_by(day, hour) — the latter merges the
# same day-of-month/hour across different months.
precip1 <- read_csv("./HB/HBEF_W3precipitation_15min.csv") %>%
  mutate(datetime = floor_date(as_datetime(DateTime), "hour")) %>%
  filter((datetime >= start_23 & datetime <= stop_23) |
         (datetime >= start24 & datetime <= stop24)) %>%
  group_by(datetime) %>%
  summarise(hourly_precip = sum(precip, na.rm = TRUE), .groups = "drop")

# Tag a datetime column with its STIC season and drop the winter gap. ----------
add_season <- function(df, dt = "datetime"){
  df %>%
    filter((.data[[dt]] >= start_23 & .data[[dt]] <= stop_23) |
           (.data[[dt]] >= start24 & .data[[dt]] <= stop24)) %>%
    mutate(season = if_else(.data[[dt]] <= stop_23, "Summer 2023", "Summer 2024"))
}
```

```{r identify-events-w3}
#| message: false
#| warning: false
# Helper that turns an eventPOT/limbs result into a per-timestep event table.
label_events <- function(discharge_df, events){
  discharge_df$event_type <- NA_character_
  discharge_df$event_id   <- NA_integer_
  for (i in seq_len(nrow(events))){
    ev <- events[i, ]
    discharge_df$event_id[ev$ris.srt:ev$fal.end] <- i
    if (!is.na(ev$ris.srt) && !is.na(ev$ris.end) && ev$ris.srt <= ev$ris.end)
      discharge_df$event_type[ev$ris.srt:ev$ris.end] <- "rising"
    if (!is.na(ev$fal.srt) && !is.na(ev$fal.end) && ev$fal.srt <= ev$fal.end)
      discharge_df$event_type[ev$fal.srt:ev$fal.end] <- "falling"
    if (i < nrow(events)){
      gap <- (ev$fal.end + 1):(events$ris.srt[i + 1] - 1)
      if (ev$fal.end + 1 <= events$ris.srt[i + 1] - 1){
        discharge_df$event_type[gap] <- "baseflow"
        discharge_df$event_id[gap]   <- NA
      }
    }
  }
  discharge_df$event_type[is.na(discharge_df$event_type)] <- "baseflow"
  discharge_df
}

## 2023 ------------------------------------------------------------------------
discharge_df <- q_23_plotting %>% select(datetime, Q_mm_day)
bf <- baseflowB(discharge_df$Q_mm_day, alpha = 0.925)
PoT_res <- eventPOT(discharge_df$Q_mm_day - bf$bf, threshold = 0.5, min.diff = 96)
events <- limbs(data = discharge_df$Q_mm_day, dates = NULL, events = PoT_res, to.plot = FALSE) %>%
  mutate(ris.srt = ris.srt - 48, fal.end = fal.end + 48)
events[1, 6] <- 1; events[nrow(events), 9] <- min(events[nrow(events), 9], nrow(discharge_df))
discharge_df_23 <- label_events(discharge_df, events)

## 2024 ------------------------------------------------------------------------
discharge_df <- w3_discharge_bind %>%
  filter(datetime >= start24 & datetime <= stop24) %>% select(datetime, Q_mm_day)
bf <- baseflowB(discharge_df$Q_mm_day, alpha = 0.925)
PoT_res <- eventPOT(discharge_df$Q_mm_day - bf$bf, threshold = 0.25, min.diff = 30)
events <- limbs(data = discharge_df$Q_mm_day, dates = NULL, events = PoT_res, to.plot = FALSE) %>%
  mutate(ris.srt = ris.srt - 48, fal.end = fal.end + 48)
events[1, 6] <- 1
discharge_df_24 <- label_events(discharge_df, events)

# ---- GLOBALLY UNIQUE EVENT IDS ---------------------------------------------
# Each year's events are numbered from 1 independently, so 2023 event 1 and
# 2024 event 1 collide when the years are combined. Offset the 2024 IDs so they
# continue after the last 2023 event: event 1 is now the first 2023 event and
# 2024 events follow chronologically. (event_id is NA during baseflow; NA + k
# stays NA.) If you applied your own fix, keep that and delete this block.
offset_23 <- max(discharge_df_23$event_id, na.rm = TRUE)
discharge_df_24 <- discharge_df_24 %>% mutate(event_id = event_id + offset_23)

# combined W3 event table (used for shading hydrographs / Figs 2 & 6) ---------
w3_events_all <- bind_rows(discharge_df_23, discharge_df_24)
```

```{r build-event-sequences-w3}
#| message: false
#| warning: false
# For each event, order sensors by FIRST time they flow -> activation sequence.
build_event_sequences <- function(input, event_df){
  ttt <- input %>%
    inner_join(event_df, by = "datetime") %>%
    filter(binary == 1) %>%
    group_by(ID, event_id) %>%
    summarise(start_of_flow = min(datetime), .groups = "drop") %>%
    arrange(start_of_flow) %>% drop_na()
  out <- NULL
  for (e in unique(ttt$event_id)){
    out <- bind_rows(out,
                     ttt %>% filter(event_id == e) %>%
                       arrange(start_of_flow) %>%
                       rowid_to_column("Sequence"))
  }
  out %>% select(-start_of_flow)
}

all_events23_sequences <- build_event_sequences(input_w3, discharge_df_23)
all_events24_sequences <- build_event_sequences(input_w3, discharge_df_24)
write_csv(all_events23_sequences, "W3event_sequences23.csv")
write_csv(all_events24_sequences, "W3event_sequences24.csv")
```

```{r events-fb-zz}
#| message: false
#| warning: false
# FB / ZZ event tables (pre-computed timestamps with event_id + event_type) ----
FBall_events <- read_csv("FB_events.csv") %>% dplyr::rename(datetime = DATETIME)
ZZall_events <- read_csv("ZZ_events.csv") %>% dplyr::rename(datetime = DATETIME)

all_eventsFB_sequences <- build_event_sequences(input_fb, FBall_events)
all_eventsZZ_sequences <- build_event_sequences(input_zz, ZZall_events)
write_csv(all_eventsFB_sequences, "FBevent_sequences.csv")
write_csv(all_eventsZZ_sequences, "ZZevent_sequences.csv")
```

## Sequence definitions

```{r flow-permanence}
#| message: false
#| warning: false
# Flow permanence (pk) = proportion of time each sensor is wet, both seasons ---
flow_permanence <- function(shed_name){
  precalc_24 <- data_24 %>% mutate(mins = minute(datetime)) %>%
    filter(wshed == shed_name, mins %in% c(0, 30)) %>%
    select(datetime, number, lat, long, binary) %>%
    group_by(number) %>% rename(DATETIME = datetime, ID = number)
  data_23 %>% mutate(mins = minute(datetime)) %>%
    filter(wshed == shed_name, mins %in% c(0, 30)) %>%
    select(datetime, ID, lat, long, binary) %>%
    group_by(ID) %>% rename(DATETIME = datetime) %>%
    rbind(precalc_24) %>%
    summarise(pk = sum(binary) / length(binary), .groups = "drop") %>%
    select(ID, pk) %>% mutate(wshed = shed_name)
}
pks_w3 <- flow_permanence("W3")
pks_fb <- flow_permanence("FB")
pks_zz <- flow_permanence("ZZ")

make_pk_seq <- function(pks) pks %>% arrange(desc(pk)) %>%
  mutate(sequence = row_number()) %>% select(ID, sequence)
W3_pk_seq <- make_pk_seq(pks_w3)
FB_pk_seq <- make_pk_seq(pks_fb)
ZZ_pk_seq <- make_pk_seq(pks_zz)

# routes (up/down adjacency) implied by the pk ordering -----------------------
pk_routes <- function(pks, ids, drop_one = FALSE){
  pks %>% filter(ID %in% ids) %>%
    { if (drop_one) filter(., pk != 1) else . } %>%
    arrange(desc(pk)) %>% mutate(down = lag(ID)) %>%
    rename(up = ID) %>% drop_na() %>% select(up, down)
}
routes_w3 <- pk_routes(pks_w3, W3_IDs)
routes_fb <- pk_routes(pks_fb, FB_IDs, drop_one = TRUE)
routes_zz <- pk_routes(pks_zz, ZZ_IDs, drop_one = TRUE)
```

```{r twi-sequences}
#| message: false
#| warning: false
# ---------------------------------------------------------------------------
# Topographic Wetness Index from a 10 m DEM (WhiteboxTools). This block writes
# intermediate rasters; once run, you can cache by setting `recompute_twi`.
# ---------------------------------------------------------------------------
recompute_twi <- TRUE   # set FALSE after first run to reuse the .tif outputs

# snapped sensor locations (extracted in ArcGIS) ------------------------------
w3_locs <- read_csv("./STIC_uaa/w32.csv") %>% select(ID, POINT_X, POINT_Y)
fb_locs <- read_csv("./STIC_uaa/fb2.csv") %>% select(ID, POINT_X, POINT_Y)
zz_locs <- read_csv("./STIC_uaa/zz1.csv") %>% select(ID, POINT_X, POINT_Y)

dem <- "./HB/1m hydro enforced DEM/dem1m.tif"

compute_twi <- function(locs, dem_path, out_dir){
  if (recompute_twi){
    m10 <- aggregate(rast(dem_path), 10)
    writeRaster(m10, file.path(out_dir, "10mdem.tif"), overwrite = TRUE)
    wbt_breach_depressions_least_cost(file.path(out_dir, "10mdem.tif"),
                                      file.path(out_dir, "10mdem_breach.tif"),
                                      dist = 10, fill = TRUE)
    wbt_fill_depressions_wang_and_liu(file.path(out_dir, "10mdem_breach.tif"),
                                      file.path(out_dir, "10mdem_fill.tif"))
    wbt_d_inf_flow_accumulation(file.path(out_dir, "10mdem_fill.tif"),
                                file.path(out_dir, "10mdem_flowacc.tif"),
                                out_type = "Specific Contributing Area")
    wbt_slope(file.path(out_dir, "10mdem_fill.tif"),
              file.path(out_dir, "10mdem_slope.tif"), units = "degrees")
    wbt_wetness_index(sca = file.path(out_dir, "10mdem_flowacc.tif"),
                      slope = file.path(out_dir, "10mdem_slope.tif"),
                      output = file.path(out_dir, "10mdem_twi.tif"))
  }
  locs_shape <- vect(locs, geom = c("POINT_X", "POINT_Y"),
                     crs = crs(rast(file.path(out_dir, "10mdem_twi.tif"))))
  terra::extract(rast(file.path(out_dir, "10mdem_twi.tif")), locs_shape,
                 ID = FALSE, bind = TRUE) %>%
    as_tibble() %>% rename(value = last_col()) %>%
    select(ID, value)
}

w3_twi <- compute_twi(w3_locs, dem, "./w3_dems")
fb_twi <- compute_twi(fb_locs, dem, "./w3_dems")   # paths kept as in original
zz_twi <- compute_twi(zz_locs, dem, "./zz_dems")

make_twi_seq <- function(twi, ids) twi %>% filter(ID %in% ids) %>%
  arrange(desc(value)) %>% mutate(sequence = row_number()) %>% select(ID, sequence)
w3_twi_sequence <- make_twi_seq(w3_twi, W3_IDs)
fb_twi_sequence <- make_twi_seq(fb_twi, FB_IDs)
zz_twi_sequence <- make_twi_seq(zz_twi, ZZ_IDs)

twi_routes <- function(twi, ids) twi %>% filter(ID %in% ids) %>%
  rename(up = ID) %>% arrange(desc(value)) %>% mutate(down = lag(up)) %>%
  drop_na() %>% select(up, down)
routes_w3_twi <- twi_routes(w3_twi, W3_IDs)
routes_fb_twi <- twi_routes(fb_twi, FB_IDs)
routes_zz_twi <- twi_routes(zz_twi, ZZ_IDs)
```

```{r event-consensus-sequences}
#| message: false
#| warning: false
# Combined per-event sequences for each watershed -----------------------------
# 2024 sequences already carry globally-unique IDs (discharge_df_24 was offset
# above), so simply bind. Event 1 = first 2023 event.
event_seqsW3 <- bind_rows(all_events23_sequences, all_events24_sequences)
event_seqsFB <- all_eventsFB_sequences
event_seqsZZ <- all_eventsZZ_sequences

# Consensus ("Event based") sequence from the pairwise precedence -------------
consensus_seq <- function(event_seqs){
  d <- event_seqs %>% dplyr::select(event_id, ID) %>%
    filter(event_id != 1) %>% dplyr::rename(sensor_id = ID) %>% as.data.frame()
  cs <- as.numeric(average_activation_sequence(d)$consensus_sequence)
  tibble(ID = cs, sequence = seq_along(cs))
}
avg_event_w3 <- consensus_seq(event_seqsW3)
avg_event_fb <- consensus_seq(event_seqsFB)
avg_event_zz <- consensus_seq(event_seqsZZ)
```

```{r bfs-sequences}
#| message: false
#| warning: false
# BFS ("best path") results were produced once and saved; read them back. -----
bfsW3 <- read_csv("breadthfs_results.csv")
bfsFB <- read_csv("breadthfs_resultsFB.csv")
bfsZZ <- read_csv("breadthfs_resultsZZ.csv")

bfs_consensus <- function(bfs){
  allout <- NULL
  for (i in unique(bfs$start)){
    allout <- bind_rows(allout,
      bfs %>% filter(start == i) %>% select(up, down) %>%
        convert_to_IDseq2() %>%
        transmute(event_id = i, sensor_id = ID))
  }
  cs <- as.numeric(average_activation_sequence(data.frame(allout))$consensus_sequence)
  tibble(ID = cs, sequence = seq_along(cs))
}
best_bfs_w3 <- bfs_consensus(bfsW3)
best_bfs_fb <- bfs_consensus(bfsFB)
best_bfs_zz <- bfs_consensus(bfsZZ)
```

## Run the model for every sequence

```{r run-model}
#| message: false
#| warning: false
# For each watershed, run the Botter-style model with each candidate sequence
# and build (a) a per-node comparison table and (b) accuracy-through-time.
run_all_models <- function(shed, input, pk, twi, bfs, ev){
  pk_m  <- calc_model_result(pk$ID,  shed) %>% dplyr::rename(pk    = pred_out)
  twi_m <- calc_model_result(twi$ID, shed) %>% dplyr::rename(twi   = pred_out)
  bfs_m <- calc_model_result(bfs$ID, shed) %>% dplyr::rename(bfs   = pred_out)
  ev_m  <- calc_model_result(ev$ID,  shed) %>% dplyr::rename(event = pred_out)

  comparison <- input %>% filter(mins %in% c(0, 30)) %>%
    left_join(pk_m,  by = c("datetime", "ID")) %>%
    left_join(twi_m, by = c("datetime", "ID")) %>%
    left_join(bfs_m, by = c("datetime", "ID")) %>%
    left_join(ev_m,  by = c("datetime", "ID")) %>%
    select(-wshed, -mins) %>% drop_na()

  comparison <- comparison %>%
    mutate(pk_eval    = mapply(get_eval_label, binary, pk),
           twi_eval   = mapply(get_eval_label, binary, twi),
           bfs_eval   = mapply(get_eval_label, binary, bfs),
           event_eval = mapply(get_eval_label, binary, event))

  df_long <- comparison %>%
    group_by(datetime) %>%
    summarise(total = n(),
              pk_accuracy    = sum(pk_eval    == "correct") / total,
              twi_accuracy   = sum(twi_eval   == "correct") / total,
              bfs_accuracy   = sum(bfs_eval   == "correct") / total,
              event_accuracy = sum(event_eval == "correct") / total,
              .groups = "drop") %>%
    select(datetime, ends_with("accuracy")) %>%
    pivot_longer(-datetime, names_to = "model", values_to = "accuracy")

  list(comparison = comparison, df_long = df_long)
}

w3_mod <- run_all_models("W3", input_w3, W3_pk_seq, w3_twi_sequence, best_bfs_w3, avg_event_w3)
fb_mod <- run_all_models("FB", input_fb, FB_pk_seq, fb_twi_sequence, best_bfs_fb, avg_event_fb)
zz_mod <- run_all_models("ZZ", input_zz, ZZ_pk_seq, zz_twi_sequence, best_bfs_zz, avg_event_zz)

df_long_w3 <- w3_mod$df_long
df_long_fb <- fb_mod$df_long
df_long_zz <- zz_mod$df_long
```

## Map layers

```{r map-layers}
#| message: false
#| warning: false
# Shared DEM / valley hillshade -----------------------------------------------
m1 <- rast(dem)
valley_hill <- rast("./HB/1m hydro enforced DEM/wall_hillshade.tif")

# Watershed outlines ----------------------------------------------------------
w3_outline <- as.polygons(rast("./w3_dems/w3_shed.tif"), extent = FALSE)
fb_outline <- as.polygons(rast("./fb_dems/fb_shed.tif"), extent = FALSE)
zz_outline <- as.polygons(rast("./zz_dems/zz_shed.tif"), extent = FALSE)

# Whole-valley boundary (all HBEF sheds dissolved) ----------------------------
sheds <- terra::project(vect("./HB/hbef_wsheds/hbef_wsheds.shp"), crs(m1))
hb_outline <- aggregate(sheds)

# Stream networks -------------------------------------------------------------
w3_net <- vect("./carrieZigZag/w3_network.shp")
fb_net <- vect("./carrieZigZag/FB_network.shp")
zz_net <- vect("./carrieZigZag/zigzag_streams.shp")

# Per-watershed hillshades ----------------------------------------------------
w3_hill <- rast("./w3_dems/1mdem_hillshade.tif")
fb_hill <- rast("./fb_dems/1mdem_hillshade.tif")
zz_hill <- rast("./zz_dems/1mdem_hillshade.tif")

# Snapped sensor locations, joined to flow permanence -------------------------
w3_stic_locs_r <- vect("w3_stic_locs_snap.shp") %>% left_join(pks_w3, by = "ID")

locs_shape_fb <- vect(data_23 %>% filter(wshed == "FB") %>%
                        select(ID, lat, long) %>% unique(),
                      geom = c("long", "lat"), crs = "+proj=longlat +datum=WGS84")
lcc_fb <- terra::project(locs_shape_fb, crs(m1)) %>% left_join(pks_fb, by = "ID")

locs_shape_zz <- vect(data_23 %>% filter(wshed == "ZZ") %>%
                        select(ID, lat, long) %>% unique(),
                      geom = c("long", "lat"), crs = "+proj=longlat +datum=WGS84")
lcc <- terra::project(locs_shape_zz, crs(m1)) %>% left_join(pks_zz, by = "ID")
```

# Figure 1 — Watershed maps with relocated NH/USA locator

```{r fig1-maps}
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 9
hill_fill <- function() scale_fill_gradientn(
  colors = c("gray9", "gray48", "lightgray", "white"), guide = "none")
pk_fill <- function() binned_scale(
  aesthetics = "fill", scale_name = "stepsn",
  palette = function(x) c("#F0F2FA", "#B3C0E5", "#4969C1", "#2A3F7A"),
  breaks = c(0.25, 0.5, 0.75), limits = c(0, 1),
  show.limits = TRUE, guide = "colorsteps", name = "Flow\npersistence")

# (a) valley context
valley_plot <- ggplot() +
  geom_spatraster(data = valley_hill, maxcell = 5e5) + hill_fill() +
  new_scale_fill() +
  geom_spatvector(data = hb_outline, fill = NA, color = "black", alpha = .6, lwd = 1, lty = 2) +
  geom_spatvector(data = fb_outline, fill = shed_colors["FB"], color = shed_colors["FB"], alpha = .6, lwd = 1) +
  geom_spatvector(data = w3_outline, fill = shed_colors["W3"], color = shed_colors["W3"], alpha = .6, lwd = 1) +
  geom_spatvector(data = zz_outline, fill = shed_colors["ZZ"], color = shed_colors["ZZ"], alpha = .6, lwd = 1) +
  ggspatial::annotation_scale(location = "tr") +
  theme_void() + theme(legend.position = "none")

# (b) locator inset: white states / black outlines, NH highlighted ------------
hbef_pt <- data.frame(long = -71.74, lat = 43.94)
usa_df <- map_data("state"); nh_df <- subset(usa_df, region == "new hampshire")
inset_plot <- ggplot() +
  geom_polygon(data = usa_df, aes(long, lat, group = group),
               fill = "white", color = "black", linewidth = 0.25) +
  geom_polygon(data = nh_df, aes(long, lat, group = group),
               fill = shed_colors["W3"], color = "black", linewidth = 0.45) +
  geom_point(data = hbef_pt, aes(long, lat), color = "#bb4430", size = 1.8) +
  coord_quickmap() + theme_void() +
  theme(plot.background = element_rect(fill = "white", color = "grey70"),
        plot.margin = margin(4, 4, 4, 4))

# (c) per-watershed flow-persistence maps
shed_map <- function(hill, outline, net, locs, col, title){
  ggplot() + geom_spatraster(data = hill) + hill_fill() + new_scale_fill() +
    geom_spatvector(data = outline, fill = NA, color = col, lwd = 1.5) +
    geom_spatvector(data = net, color = "#9AC0CD", lwd = 1) +
    geom_spatvector(data = locs, aes(fill = pk), size = 3, pch = 21) +
    pk_fill() + labs(title = title) + theme_void() +
    theme(plot.title = element_text(hjust = .5))
}
w3_map <- shed_map(w3_hill, w3_outline, w3_net, w3_stic_locs_r, shed_colors["W3"], "W3")
fb_map <- shed_map(fb_hill, fb_outline, fb_net, lcc_fb, shed_colors["FB"], "FB")
zz_map <- shed_map(zz_hill, zz_outline, zz_net, lcc, shed_colors["ZZ"], "ZZ")

# inset placed BESIDE the valley map (no overlap)
top <- valley_plot + inset_plot + plot_layout(widths = c(3, 1))
figure1 <- top /
  (w3_map + fb_map + zz_map + plot_layout(guides = "collect")) +
  plot_layout(heights = c(1.2, 1)) + plot_annotation(tag_levels = "a")
figure1
ggsave("figure1_maps.pdf", figure1, width = 12, height = 9)
```

# Figure 2 - Conceptual diagram of how sequences were defined (made in illustrator)
# Figure 2 — W3: event-specific vs. consensus activation order (events 7 & 10)
*Changes- remove lower row
```{r fig2-functions}
#| message: false
#| warning: false
# Colours: white = missing data, light grey = dry, black = flowing.
flow_tile_cols <- c("0" = "grey80", "1" = "black")

train_sec <- function(primary, secondary, na.rm = TRUE){
  from <- range(secondary, na.rm = na.rm); to <- range(primary, na.rm = na.rm)
  list(fwd = function(x) rescale(x, from = from, to = to),
       rev = function(x) rescale(x, from = to,   to = from))
}

event_specific_order <- function(ev_dat, min_run_hours = 4){
  ev_dat %>% arrange(ID, datetime) %>% group_by(ID) %>%
    mutate(gap = as.numeric(difftime(datetime, lag(datetime, default = first(datetime)), units = "hours")),
           new_run = ifelse(row_number() == 1 | binary == 0 | gap > 1, 1, 0),
           run_id  = cumsum(new_run)) %>% ungroup() %>%
    filter(binary == 1) %>% group_by(ID, run_id) %>%
    summarise(run_start = min(datetime),
              run_dur = as.numeric(difftime(max(datetime), min(datetime), units = "hours")),
              .groups = "drop") %>%
    filter(run_dur >= min_run_hours) %>% group_by(ID) %>%
    summarise(start_of_flow = min(run_start), .groups = "drop") %>%
    arrange(start_of_flow) %>% mutate(event_rank = row_number()) %>% select(ID, event_rank)
}

fig2_compare_events <- function(shed, input, event_df, consensus, ids, event_ids,
                                precip = NULL, discharge = NULL, discharge_col = "Q_mm_day",
                                hydro_height = 2, tile_height = 5){
  if (length(event_ids) != 2) stop("Please supply exactly two event IDs.")
  has_hydro <- !is.null(precip) && !is.null(discharge)
  cons_rank <- consensus %>% dplyr::rename(consensus_rank = sequence) %>% select(ID, consensus_rank)
  dat_all <- input %>% filter(ID %in% ids) %>%
    inner_join(event_df %>% select(datetime, event_id, event_type), by = "datetime") %>%
    filter(!is.na(event_id), event_type != "baseflow") %>%
    left_join(cons_rank, by = "ID")

  tile_layer <- function(ymax) list(
    geom_tile(aes(fill = factor(binary))),
    scale_fill_manual(values = flow_tile_cols, breaks = c("0", "1"),
                      labels = c("Dry", "Flowing"), name = "", drop = FALSE, na.value = "white"),
    scale_y_continuous(limits = c(0.5, ymax + 0.5), expand = c(0, 0)),
    scale_x_datetime(date_breaks = "1 day", date_labels = "%m/%d"),
    theme_classic(),
    theme(legend.position = "right", strip.background = element_blank(),
          panel.background = element_rect(fill = "white", colour = NA),
          axis.title.y = element_text(angle = 90, size = 12), axis.text = element_text(size = 12),
          plot.title = element_text(size = 9, hjust = 0.5, face = "bold"),
          plot.margin = unit(c(2, 2, 2, 2), "pt")))

  make_top <- function(ev){
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
    ev_ranks <- event_specific_order(ev_dat)
    if (nrow(ev_ranks) == 0) stop("Event ", ev, " has no sustained flow.")
    ev_ranked <- ev_dat %>% inner_join(ev_ranks, by = "ID")
    p_tiles <- ggplot(ev_ranked, aes(x = datetime, y = event_rank)) +
      tile_layer(nrow(ev_ranks)) +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", y = "Event order (1 = wets first)")
    if (!has_hydro) return(p_tiles + labs(title = paste("Event", ev, "- event order")))

    ev_water <- precip %>% select(datetime, hourly_precip) %>%
      filter(datetime >= ev_s, datetime <= ev_e) %>%
      left_join(discharge %>% select(datetime, all_of(discharge_col)), by = "datetime")
    sec <- train_sec(primary = c(0, max(ev_water$hourly_precip, na.rm = TRUE)),
                     secondary = range(ev_water[[discharge_col]], na.rm = TRUE))
    ev_water$.q <- sec$fwd(ev_water[[discharge_col]])
    ytop <- max(c(ev_water$hourly_precip, ev_water$.q), na.rm = TRUE) * 1.10  # headroom
    p_hydro <- ggplot(ev_water, aes(x = datetime)) +
      geom_col(aes(y = hourly_precip), width = 3600, fill = "grey40") +
      geom_line(aes(y = .q), colour = "blue", linewidth = 0.6) +
      scale_y_continuous("Precip (mm)", limits = c(0, ytop), expand = c(0, 0),
                         sec.axis = sec_axis(~ sec$rev(.), name = "Q (mm/day)")) +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", title = paste("Event", ev, "- event-specific order")) +
      theme_classic() +
      theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
            axis.text.y = element_text(size = 12),
            axis.title.y = element_text(angle = 90, size = 12),
            axis.title.y.right = element_text(angle = 90, size = 12),
            plot.title = element_text(size = 9, hjust = 0.5, face = "bold"),
            plot.margin = unit(c(2, 2, 2, 2), "pt"))
    p_hydro / p_tiles + plot_layout(heights = c(hydro_height, tile_height))
  }
  make_bottom <- function(ev){
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
    ggplot(ev_dat, aes(x = datetime, y = consensus_rank)) +
      tile_layer(max(cons_rank$consensus_rank, na.rm = TRUE)) +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", y = "Consensus order (1 = wets first)",
           title = paste("Event", ev, "- consensus order"))
  }
  (make_top(event_ids[1]) | make_top(event_ids[2])) /
  (make_bottom(event_ids[1]) | make_bottom(event_ids[2])) +
    plot_layout(heights = c(hydro_height + tile_height, tile_height), guides = "collect") +
    plot_annotation(title = paste0(shed, ": event-specific vs. consensus activation order"),
      caption = "Top: each event in its own order (with precip bars + discharge line). Bottom: same events forced into the consensus order. White = missing, grey = dry, black = flowing.")
}
```

```{r}

flow_tile_cols <- c("0" = "grey80", "1" = "#586BA4")

train_sec <- function(primary, secondary, na.rm = TRUE){
  from <- range(secondary, na.rm = na.rm); to <- range(primary, na.rm = na.rm)
  list(fwd = function(x) rescale(x, from = from, to = to),
       rev = function(x) rescale(x, from = to,   to = from))
}

event_specific_order <- function(ev_dat, min_run_hours = 4){
  ev_dat %>% arrange(ID, datetime) %>% group_by(ID) %>%
    mutate(gap = as.numeric(difftime(datetime, lag(datetime, default = first(datetime)), units = "hours")),
           new_run = ifelse(row_number() == 1 | binary == 0 | gap > 1, 1, 0),
           run_id  = cumsum(new_run)) %>% ungroup() %>%
    filter(binary == 1) %>% group_by(ID, run_id) %>%
    summarise(run_start = min(datetime),
              run_dur = as.numeric(difftime(max(datetime), min(datetime), units = "hours")),
              .groups = "drop") %>%
    filter(run_dur >= min_run_hours) %>% group_by(ID) %>%
    summarise(start_of_flow = min(run_start), .groups = "drop") %>%
    arrange(start_of_flow) %>% mutate(event_rank = row_number()) %>% select(ID, event_rank)
}

fig2_compare_events <- function(shed, input, event_df, consensus, ids, event_ids,
                                precip = NULL, discharge = NULL, discharge_col = "Q_mm_day",
                                hydro_height = 2, tile_height = 5){

  if (length(event_ids) != 2) stop("Please supply exactly two event IDs.")
  has_hydro <- !is.null(precip) && !is.null(discharge)

  cons_rank <- consensus %>% dplyr::rename(consensus_rank = sequence) %>% select(ID, consensus_rank)

  dat_all <- input %>% filter(ID %in% ids) %>%
    inner_join(event_df %>% select(datetime, event_id, event_type), by = "datetime") %>%
    filter(!is.na(event_id), event_type != "baseflow") %>%
    left_join(cons_rank, by = "ID")

  # Global hydro ranges ------------------------------------------------------
  hydro_list <- list()
  sec_global <- NULL
  global_ytop <- NULL
  if (has_hydro) {
    for (ev in event_ids) {
      ev_dat <- dat_all %>% filter(event_id == ev)
      ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
      hydro_list[[as.character(ev)]] <- precip %>%
        select(datetime, hourly_precip) %>%
        filter(datetime >= ev_s, datetime <= ev_e) %>%
        left_join(discharge %>% select(datetime, all_of(discharge_col)), by = "datetime")
    }
    all_hydro <- bind_rows(hydro_list)
    global_precip_max <- max(all_hydro$hourly_precip, na.rm = TRUE)
    global_q_range    <- range(all_hydro[[discharge_col]], na.rm = TRUE)
    sec_global  <- train_sec(primary = c(0, global_precip_max), secondary = global_q_range)
    global_ytop <- global_precip_max * 1.18
  }

  # Common tile layers -------------------------------------------------------
  tile_layer <- function(ymax) list(
    geom_tile(aes(fill = factor(binary))),
    scale_fill_manual(values = flow_tile_cols, breaks = c("0", "1"),
                      labels = c("Dry", "Flowing"), name = "", drop = FALSE,
                      na.value = "white"),
    scale_y_continuous(limits = c(0.5, ymax + 0.5), expand = c(0, 0)),
    scale_x_datetime(date_breaks = "1 day", date_labels = "%m/%d"),
    theme_classic(),
    theme(legend.position = "right", strip.background = element_blank(),
          panel.background = element_rect(fill = "white", colour = NA),
          axis.title.y = element_text(angle = 90, size = 12),
          axis.text    = element_text(size = 12),
          plot.title   = element_text(size = 9, hjust = 0.5, face = "bold"),
          plot.margin  = unit(c(2, 2, 2, 2), "pt")))

  # Hydro panel builder (modified legend + black line) -----------------------
  build_hydro <- function(ev) {
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
    ev_water <- hydro_list[[as.character(ev)]]
    ev_water$.q <- sec_global$fwd(ev_water[[discharge_col]])

    ggplot(ev_water, aes(x = datetime)) +
      # Mapped fill aesthetic for legend entry
      geom_col(aes(y = hourly_precip, fill = "Precipitation"),
               width = 3600, show.legend = TRUE) +
      # Mapped color aesthetic for legend entry; line is now black
      geom_line(aes(y = .q, color = "Discharge"),
                linewidth = 0.6, show.legend = TRUE) +
      scale_fill_manual(values = c("Precipitation" = "grey40"), name = "") +
      scale_color_manual(values = c("Discharge" = "black"), name = "") +
      scale_y_continuous("Precip (mm)", limits = c(0, global_ytop), expand = c(0, 0),
                         sec.axis = sec_axis(~ sec_global$rev(.), name = "Q (mm/day)")) +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", title = paste("Event", ev)) +
      theme_classic() +
      theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
            axis.text.y = element_text(size = 12),
            axis.title.y = element_text(angle = 90, size = 12),
            axis.title.y.right = element_text(angle = 90, size = 12),
            plot.title = element_text(size = 9, hjust = 0.5, face = "bold"),
            plot.margin = unit(c(2, 2, 2, 2), "pt"))
  }

  # Event-specific tile builder ----------------------------------------------
  build_event_tiles <- function(ev) {
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
    ev_ranks <- event_specific_order(ev_dat)
    if (nrow(ev_ranks) == 0) stop("Event ", ev, " has no sustained flow.")
    ev_ranked <- ev_dat %>% inner_join(ev_ranks, by = "ID")

    ggplot(ev_ranked, aes(x = datetime, y = event_rank)) +
      tile_layer(nrow(ev_ranks)) +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", y = "Event order (1 = wets first)")
  }

  # Consensus tile builder ---------------------------------------------------
  build_cons_tiles <- function(ev) {
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)

    ggplot(ev_dat, aes(x = datetime, y = consensus_rank)) +
      tile_layer(max(cons_rank$consensus_rank, na.rm = TRUE)) +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", y = "Consensus order (1 = wets first)",
           title = paste("Event", ev, "- consensus order"))
  }

  # Assemble flat grid -------------------------------------------------------
  if (has_hydro) {
    plots <- list(
      build_hydro(event_ids[1]),       build_hydro(event_ids[2]),
      build_event_tiles(event_ids[1]), build_event_tiles(event_ids[2]),
      build_cons_tiles(event_ids[1]),  build_cons_tiles(event_ids[2])
    )
    heights <- c(hydro_height, tile_height, tile_height)
  } else {
    plots <- list(
      build_event_tiles(event_ids[1]), build_event_tiles(event_ids[2]),
      build_cons_tiles(event_ids[1]),  build_cons_tiles(event_ids[2])
    )
    heights <- c(tile_height, tile_height)
  }

  wrap_plots(plots, ncol = 2, byrow = TRUE) +
    plot_layout(heights = heights, widths = 1, guides = "collect") +
    plot_annotation(
      title = paste0(shed, ": event-specific vs. consensus activation order"),
      caption = paste(
        "Top row: precip + discharge.",
        "Middle row: event-specific activation order.",
        "Bottom row: consensus activation order.",
        "White = missing, grey = dry, black = flowing."
      )
    )
}
```

```{r drying version}

flow_tile_cols <- c("0" = "grey80", "1" = "black")

train_sec <- function(primary, secondary, na.rm = TRUE){
  from <- range(secondary, na.rm = na.rm); to <- range(primary, na.rm = na.rm)
  list(fwd = function(x) rescale(x, from = from, to = to),
       rev = function(x) rescale(x, from = to,   to = from))
}

# -------------------------------------------------------------------------
# Drying rank for a single event: last flowing time after peak discharge
# -------------------------------------------------------------------------
event_drying_rank <- function(ev_dat){
  counts <- ev_dat %>% 
    group_by(datetime) %>% 
    summarise(nw = sum(binary == 1, na.rm = TRUE), .groups = "drop")
  
  if (max(counts$nw) == 0) return(tibble(ID = character(), dry_rank = integer()))
  
  peak_t <- counts$datetime[which.max(counts$nw)]
  
  ev_dat %>%
    filter(datetime >= peak_t, binary == 1) %>%
    group_by(ID) %>%
    summarise(dry_time = max(datetime), .groups = "drop") %>%
    arrange(dry_time) %>%                # dries first = smallest dry_time
    mutate(dry_rank = row_number()) %>%
    select(ID, dry_rank)
}

# -------------------------------------------------------------------------
fig2_compare_events_drying <- function(shed, input, event_df, consensus_dry, ids, event_ids,
                                       precip = NULL, discharge = NULL, 
                                       discharge_col = "Q_mm_day",
                                       hydro_height = 2, tile_height = 5){
  
  if (length(event_ids) != 2) stop("Please supply exactly two event IDs.")
  has_hydro <- !is.null(precip) && !is.null(discharge)
  
  # consensus_dry should have columns: ID, sequence (1 = dries first)
  cons_rank <- consensus_dry %>% 
    dplyr::rename(consensus_rank = sequence) %>% 
    select(ID, consensus_rank)
  
  # core data --------------------------------------------------------------
  dat_all <- input %>% 
    filter(ID %in% ids) %>%
    inner_join(event_df %>% select(datetime, event_id, event_type), by = "datetime") %>%
    filter(!is.na(event_id), event_type != "baseflow") %>%
    left_join(cons_rank, by = "ID")
  
  # global hydro ranges across both events ---------------------------------
  hydro_list <- list()
  sec_global <- NULL
  global_ytop <- NULL
  
  if (has_hydro) {
    for (ev in event_ids) {
      ev_dat <- dat_all %>% filter(event_id == ev)
      ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
      hydro_list[[as.character(ev)]] <- precip %>%
        select(datetime, hourly_precip) %>%
        filter(datetime >= ev_s, datetime <= ev_e) %>%
        left_join(discharge %>% select(datetime, all_of(discharge_col)), by = "datetime")
    }
    all_hydro <- bind_rows(hydro_list)
    global_precip_max <- max(all_hydro$hourly_precip, na.rm = TRUE)
    global_q_range    <- range(all_hydro[[discharge_col]], na.rm = TRUE)
    sec_global  <- train_sec(primary = c(0, global_precip_max), secondary = global_q_range)
    global_ytop <- global_precip_max * 1.18
  }
  
  # common tile layers -----------------------------------------------------
  tile_layer <- function(ymax) list(
    geom_tile(aes(fill = factor(binary))),
    scale_fill_manual(values = flow_tile_cols, breaks = c("0", "1"),
                      labels = c("Dry", "Flowing"), name = "", drop = FALSE,
                      na.value = "white"),
    scale_y_continuous(limits = c(0.5, ymax + 0.5), expand = c(0, 0)),
    scale_x_datetime(date_breaks = "1 day", date_labels = "%m/%d"),
    theme_classic(),
    theme(legend.position = "right", strip.background = element_blank(),
          panel.background = element_rect(fill = "white", colour = NA),
          axis.title.y = element_text(angle = 90, size = 12),
          axis.text    = element_text(size = 12),
          plot.title   = element_text(size = 9, hjust = 0.5, face = "bold"),
          plot.margin  = unit(c(2, 2, 2, 2), "pt")))
  
  # hydro panel ------------------------------------------------------------
  build_hydro <- function(ev) {
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
    ev_water <- hydro_list[[as.character(ev)]]
    ev_water$.q <- sec_global$fwd(ev_water[[discharge_col]])
    
    ggplot(ev_water, aes(x = datetime)) +
      geom_col(aes(y = hourly_precip, fill = "Precipitation"),
               width = 3600, show.legend = TRUE) +
      geom_line(aes(y = .q, color = "Discharge"),
                linewidth = 0.6, show.legend = TRUE) +
      scale_fill_manual(values = c("Precipitation" = "grey40"), name = "") +
      scale_color_manual(values = c("Discharge" = "black"), name = "") +
      scale_y_continuous("Precip (mm)", limits = c(0, global_ytop), expand = c(0, 0),
                         sec.axis = sec_axis(~ sec_global$rev(.), name = "Q (mm/day)")) +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", title = paste("Event", ev)) +
      theme_classic() +
      theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
            axis.text.y = element_text(size = 12),
            axis.title.y = element_text(angle = 90, size = 12),
            axis.title.y.right = element_text(angle = 90, size = 12),
            plot.title = element_text(size = 9, hjust = 0.5, face = "bold"),
            plot.margin = unit(c(2, 2, 2, 2), "pt"))
  }
  
  # event-specific drying order tiles --------------------------------------
  build_event_tiles <- function(ev) {
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
    dry_ranks <- event_drying_rank(ev_dat)
    
    if (nrow(dry_ranks) == 0) stop("Event ", ev, " has no drying sequence.")
    
    ev_ranked <- ev_dat %>% inner_join(dry_ranks, by = "ID")
    
    ggplot(ev_ranked, aes(x = datetime, y = dry_rank)) +
      tile_layer(nrow(dry_ranks)) +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", y = "Drying order (1 = dries first)")
  }
  
  # consensus drying order tiles -------------------------------------------
  build_cons_tiles <- function(ev) {
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
    
    ggplot(ev_dat, aes(x = datetime, y = consensus_rank)) +
      tile_layer(max(cons_rank$consensus_rank, na.rm = TRUE)) +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", y = "Consensus drying order (1 = dries first)",
           title = paste("Event", ev, "- consensus drying"))
  }
  
  # flat 3×2 grid assembly -------------------------------------------------
  if (has_hydro) {
    plots <- list(
      build_hydro(event_ids[1]),       build_hydro(event_ids[2]),
      build_event_tiles(event_ids[1]), build_event_tiles(event_ids[2]),
      build_cons_tiles(event_ids[1]),  build_cons_tiles(event_ids[2])
    )
    heights <- c(hydro_height, tile_height, tile_height)
  } else {
    plots <- list(
      build_event_tiles(event_ids[1]), build_event_tiles(event_ids[2]),
      build_cons_tiles(event_ids[1]),  build_cons_tiles(event_ids[2])
    )
    heights <- c(tile_height, tile_height)
  }
  
  wrap_plots(plots, ncol = 2, byrow = TRUE) +
    plot_layout(heights = heights, widths = 1, guides = "collect") +
    plot_annotation(
      title = paste0(shed, ": event-specific vs. consensus drying order"),
      caption = paste(
        "Top row: precip + discharge.",
        "Middle row: drying sequence for each event (rank 1 at bottom = dries first).",
        "Bottom row: same events forced into the consensus drying order.",
        "White = missing, grey = dry, black = flowing."
      )
    )
}
```


```{r fig2-w3}
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 9
figure2 <- fig2_compare_events(
  shed = "W3", input = input_w3, event_df = w3_events_all,
  consensus = avg_event_w3, ids = W3_IDs, event_ids = c(10, 12),
  precip = precip1, discharge = w3_q_seasonal, discharge_col = "Q_mm_day")
#figure2 <- figure2 + theme(text = element_text(size = 12))
ggsave("figure_coop.pdf", figure2, width = 12, height = 9)

fig2_compare_events_drying(
  shed          = "W3",
  input         = input_w3,
  event_df      = w3_events_all,
  consensus_dry = dry_w3,          # from your direction_sequences / cons_dir code
  ids           = W3_IDs,
  event_ids     = c(12, 10),
  precip        = precip1,
  discharge     = w3_q_seasonal,
  discharge_col = "Q_mm_day"
)

ggsave("figure2_drying_events_w3.pdf", figure2_drying, width = 12, height = 9)
```

## Figure 2: one wetting and one drying
```{r}

flow_tile_cols <- c("0" = "grey80", "1" = "black")

train_sec <- function(primary, secondary, na.rm = TRUE){
  from <- range(secondary, na.rm = na.rm); to <- range(primary, na.rm = na.rm)
  list(fwd = function(x) rescale(x, from = from, to = to),
       rev = function(x) rescale(x, from = to,   to = from))
}

event_specific_order <- function(ev_dat, min_run_hours = 4){
  ev_dat %>% arrange(ID, datetime) %>% group_by(ID) %>%
    mutate(gap = as.numeric(difftime(datetime, lag(datetime, default = first(datetime)), units = "hours")),
           new_run = ifelse(row_number() == 1 | binary == 0 | gap > 1, 1, 0),
           run_id  = cumsum(new_run)) %>% ungroup() %>%
    filter(binary == 1) %>% group_by(ID, run_id) %>%
    summarise(run_start = min(datetime),
              run_dur = as.numeric(difftime(max(datetime), min(datetime), units = "hours")),
              .groups = "drop") %>%
    filter(run_dur >= min_run_hours) %>% group_by(ID) %>%
    summarise(start_of_flow = min(run_start), .groups = "drop") %>%
    arrange(start_of_flow) %>% mutate(event_rank = row_number()) %>% select(ID, event_rank)
}

# FIXED: arrange(dry_time) so dry_rank 1 = dries FIRST (earliest max flowing time after peak)
event_drying_rank <- function(ev_dat){
  counts <- ev_dat %>% 
    group_by(datetime) %>% 
    summarise(nw = sum(binary == 1, na.rm = TRUE), .groups = "drop")
  if(max(counts$nw) == 0) return(tibble(ID = character(), dry_rank = integer()))
  peak_t <- counts$datetime[which.max(counts$nw)]
  
  ev_dat %>%
    filter(datetime >= peak_t, binary == 1) %>%
    group_by(ID) %>%
    summarise(dry_time = max(datetime), .groups = "drop") %>%
    arrange(dry_time) %>%      # earliest dry_time = dries first = rank 1
    mutate(dry_rank = row_number()) %>%
    select(ID, dry_rank)
}

fig2_wet_dry_columns <- function(shed, input, event_df,
                                 consensus_wet, consensus_dry,
                                 ids, event_wet, event_dry,
                                 precip = NULL, discharge = NULL,
                                 discharge_col = "Q_mm_day",
                                 hydro_height = 2, tile_height = 5){
  
  has_hydro <- !is.null(precip) && !is.null(discharge)
  global_n  <- length(ids)
  
  cons_wet <- consensus_wet %>% dplyr::rename(consensus_rank = sequence) %>% select(ID, consensus_rank)
  cons_dry <- consensus_dry %>% dplyr::rename(consensus_rank = sequence) %>% select(ID, consensus_rank)
  max_cons_dry <- max(cons_dry$consensus_rank, na.rm = TRUE)
  
  dat_all <- input %>% 
    filter(ID %in% ids) %>%
    inner_join(event_df %>% select(datetime, event_id, event_type), by = "datetime") %>%
    filter(!is.na(event_id), event_type != "baseflow")
  
  # global hydro limits across both events
  hydro_list <- list(); sec_global <- NULL; global_ytop <- NULL
  if(has_hydro){
    for(ev in unique(c(event_wet, event_dry))){
      ev_dat <- dat_all %>% filter(event_id == ev)
      ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
      hydro_list[[as.character(ev)]] <- precip %>%
        select(datetime, hourly_precip) %>%
        filter(datetime >= ev_s, datetime <= ev_e) %>%
        left_join(discharge %>% select(datetime, all_of(discharge_col)), by = "datetime")
    }
    all_hydro <- bind_rows(hydro_list)
    global_precip_max <- max(all_hydro$hourly_precip, na.rm = TRUE)
    global_q_range    <- range(all_hydro[[discharge_col]], na.rm = TRUE)
    sec_global  <- train_sec(primary = c(0, global_precip_max), secondary = global_q_range)
    global_ytop <- global_precip_max * 1.18
  }
  
  # tile layer with fixed global y-extent and white NA padding
  tile_layer <- function() list(
    geom_tile(aes(fill = factor(binary))),
    scale_fill_manual(values = flow_tile_cols, breaks = c("0","1"),
                      labels = c("Dry","Flowing"), name = "", drop = FALSE,
                      na.value = "white"),
    scale_y_continuous("Activation / Drying order",
                       limits = c(0.5, global_n + 0.5),
                       expand = c(0, 0),
                       breaks = function(lims) pretty(lims, n = 5)),
    scale_x_datetime(date_breaks = "1 day", date_labels = "%m/%d"),
    theme_classic() +
      theme(legend.position = "right",
            strip.background = element_blank(),
            panel.background = element_rect(fill = "white", colour = NA),
            axis.title.y = element_text(angle = 90, size = 12),
            axis.text    = element_text(size = 12),
            plot.title   = element_text(size = 9, hjust = 0.5, face = "bold"),
            plot.margin  = unit(c(2, 2, 2, 2), "pt")))
  
  # --- hydro panels --------------------------------------------------------
  build_hydro <- function(ev){
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
    ev_water <- hydro_list[[as.character(ev)]]
    ev_water$.q <- sec_global$fwd(ev_water[[discharge_col]])
    
    ggplot(ev_water, aes(x = datetime)) +
      geom_col(aes(y = hourly_precip, fill = "Precipitation"), width = 3600, show.legend = TRUE) +
      geom_line(aes(y = .q, color = "Discharge"), linewidth = 0.6, show.legend = TRUE) +
      scale_fill_manual(values = c("Precipitation" = "grey40"), name = "") +
      scale_color_manual(values = c("Discharge" = "black"), name = "") +
      scale_y_continuous("Precip (mm)", limits = c(0, global_ytop), expand = c(0, 0),
                         sec.axis = sec_axis(~ sec_global$rev(.), name = "Q (mm/day)")) +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", title = paste("Event", ev)) +
      theme_classic() +
      theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
            axis.text.y = element_text(size = 12),
            axis.title.y = element_text(angle = 90, size = 12),
            axis.title.y.right = element_text(angle = 90, size = 12),
            plot.title = element_text(size = 9, hjust = 0.5, face = "bold"),
            plot.margin = unit(c(2, 2, 2, 2), "pt"))
  }
  
  # --- wetting event-specific tiles (1 = wets first, at bottom) -----------
  build_wet_tiles <- function(ev){
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
    ev_ranks <- event_specific_order(ev_dat)
    if(nrow(ev_ranks) == 0) stop("Event ", ev, " has no sustained wetting sequence.")
    
    ev_full <- ev_dat %>%
      inner_join(ev_ranks, by = "ID") %>%
      mutate(y_rank = event_rank) %>%
      select(datetime, y_rank, binary) %>%
      ungroup() %>%
      complete(datetime, y_rank = 1:global_n)
    
    ggplot(ev_full, aes(x = datetime, y = y_rank)) +
      tile_layer() +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", y = "Wetting order (1 = wets first)")
  }
  
  # --- wetting consensus tiles ---------------------------------------------
  build_wet_cons <- function(ev){
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
    
    ev_full <- ev_dat %>%
      inner_join(cons_wet, by = "ID") %>%
      mutate(y_rank = consensus_rank) %>%
      select(datetime, y_rank, binary) %>%
      ungroup() %>%
      complete(datetime, y_rank = 1:global_n)
    
    ggplot(ev_full, aes(x = datetime, y = y_rank)) +
      tile_layer() +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", y = "Consensus wetting",
           title = paste("Event", ev, "- consensus wetting"))
  }
  
  # --- drying event-specific tiles (N at top = dries first) ----------------
  build_dry_tiles <- function(ev){
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
    dry_ranks <- event_drying_rank(ev_dat)
    if(nrow(dry_ranks) == 0) stop("Event ", ev, " has no drying sequence.")
    n_dry <- nrow(dry_ranks)
    
    # dry_rank 1 = dries first
    # n_dry - dry_rank + 1 puts first-to-dry at the TOP (largest y-value)
    ev_full <- ev_dat %>%
      inner_join(dry_ranks, by = "ID") %>%
      mutate(y_rank = n_dry - dry_rank + 1) %>%
      select(datetime, y_rank, binary) %>%
      ungroup() %>%
      complete(datetime, y_rank = 1:global_n)
    
    ggplot(ev_full, aes(x = datetime, y = y_rank)) +
      tile_layer() +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", y = "Drying order (N = dries first)")
  }
  
  # --- drying consensus tiles (N at top = dries first) ---------------------
  build_dry_cons <- function(ev){
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
    
    # consensus_rank 1 = dries first
    # max_cons_dry - consensus_rank + 1 puts first-to-dry at top
    ev_full <- ev_dat %>%
      inner_join(cons_dry, by = "ID") %>%
      mutate(y_rank = max_cons_dry - consensus_rank + 1) %>%
      select(datetime, y_rank, binary) %>%
      ungroup() %>%
      complete(datetime, y_rank = 1:global_n)
    
    ggplot(ev_full, aes(x = datetime, y = y_rank)) +
      tile_layer() +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", y = "Consensus drying (N = dries first)",
           title = paste("Event", ev, "- consensus drying"))
  }
  
  # --- assemble flat grid --------------------------------------------------
  if(has_hydro){
    plots <- list(
      build_hydro(event_wet),    build_hydro(event_dry),
      build_wet_tiles(event_wet), build_dry_tiles(event_dry),
      build_wet_cons(event_wet),  build_dry_cons(event_dry)
    )
    heights <- c(hydro_height, tile_height, tile_height)
  } else {
    plots <- list(
      build_wet_tiles(event_wet), build_dry_tiles(event_dry),
      build_wet_cons(event_wet),  build_dry_cons(event_dry)
    )
    heights <- c(tile_height, tile_height)
  }
  
  wrap_plots(plots, ncol = 2, byrow = TRUE) +
    plot_layout(heights = heights, widths = 1, guides = "collect") +
    plot_annotation(
      title = paste0(shed, ": wetting vs. drying activation order"),
      caption = paste(
        "Left column: wetting sequence (1 = wets first, bottom).",
        "Right column: drying sequence (largest number = dries first, top).",
        "All tile panels share the same y-axis extent; missing rows are white."
      )
    )
}
```

```{r}
figure_combo <- fig2_wet_dry_columns(
  shed          = "W3",
  input         = input_w3,
  event_df      = w3_events_all,
  consensus_wet = avg_event_w3,
  consensus_dry = dry_w3,
  ids           = W3_IDs,
  event_wet     = 7,
  event_dry     = 12,
  precip        = precip1,
  discharge     = w3_q_seasonal,
  discharge_col = "Q_mm_day"
)

ggsave("figure2_wet_vs_dry.pdf", figure_combo, width = 7.5, units ="in")
```

## without legend
```{r}
flow_tile_cols <- c("0" = "grey80", "1" = "black")

train_sec <- function(primary, secondary, na.rm = TRUE){
  from <- range(secondary, na.rm = na.rm); to <- range(primary, na.rm = na.rm)
  list(fwd = function(x) rescale(x, from = from, to = to),
       rev = function(x) rescale(x, from = to,   to = from))
}

event_specific_order <- function(ev_dat, min_run_hours = 4){
  ev_dat %>% arrange(ID, datetime) %>% group_by(ID) %>%
    mutate(gap = as.numeric(difftime(datetime, lag(datetime, default = first(datetime)), units = "hours")),
           new_run = ifelse(row_number() == 1 | binary == 0 | gap > 1, 1, 0),
           run_id  = cumsum(new_run)) %>% ungroup() %>%
    filter(binary == 1) %>% group_by(ID, run_id) %>%
    summarise(run_start = min(datetime),
              run_dur = as.numeric(difftime(max(datetime), min(datetime), units = "hours")),
              .groups = "drop") %>%
    filter(run_dur >= min_run_hours) %>% group_by(ID) %>%
    summarise(start_of_flow = min(run_start), .groups = "drop") %>%
    arrange(start_of_flow) %>% mutate(event_rank = row_number()) %>% select(ID, event_rank)
}

event_drying_rank <- function(ev_dat){
  counts <- ev_dat %>% 
    group_by(datetime) %>% 
    summarise(nw = sum(binary == 1, na.rm = TRUE), .groups = "drop")
  if(max(counts$nw) == 0) return(tibble(ID = character(), dry_rank = integer()))
  peak_t <- counts$datetime[which.max(counts$nw)]
  ev_dat %>%
    filter(datetime >= peak_t, binary == 1) %>%
    group_by(ID) %>%
    summarise(dry_time = max(datetime), .groups = "drop") %>%
    arrange(dry_time) %>%
    mutate(dry_rank = row_number()) %>%
    select(ID, dry_rank)
}

fig2_wet_dry_columns <- function(shed, input, event_df,
                                 consensus_wet, consensus_dry,
                                 ids, event_wet, event_dry,
                                 precip = NULL, discharge = NULL,
                                 discharge_col = "Q_mm_day",
                                 hydro_height = 2, tile_height = 5){
  
  has_hydro <- !is.null(precip) && !is.null(discharge)
  global_n  <- length(ids)
  
  cons_wet <- consensus_wet %>% dplyr::rename(consensus_rank = sequence) %>% select(ID, consensus_rank)
  cons_dry <- consensus_dry %>% dplyr::rename(consensus_rank = sequence) %>% select(ID, consensus_rank)
  max_cons_dry <- max(cons_dry$consensus_rank, na.rm = TRUE)
  
  dat_all <- input %>% 
    filter(ID %in% ids) %>%
    inner_join(event_df %>% select(datetime, event_id, event_type), by = "datetime") %>%
    filter(!is.na(event_id), event_type != "baseflow")
  
  hydro_list <- list(); sec_global <- NULL; global_ytop <- NULL
  if(has_hydro){
    for(ev in unique(c(event_wet, event_dry))){
      ev_dat <- dat_all %>% filter(event_id == ev)
      ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
      hydro_list[[as.character(ev)]] <- precip %>%
        select(datetime, hourly_precip) %>%
        filter(datetime >= ev_s, datetime <= ev_e) %>%
        left_join(discharge %>% select(datetime, all_of(discharge_col)), by = "datetime")
    }
    all_hydro <- bind_rows(hydro_list)
    global_precip_max <- max(all_hydro$hourly_precip, na.rm = TRUE)
    global_q_range    <- range(all_hydro[[discharge_col]], na.rm = TRUE)
    sec_global  <- train_sec(primary = c(0, global_precip_max), secondary = global_q_range)
    global_ytop <- global_precip_max * 1.18
  }
  
  # -- tile layer with NO LEGEND -------------------------------------------
  tile_layer <- function() list(
    geom_tile(aes(fill = factor(binary))),
    scale_fill_manual(values = flow_tile_cols, breaks = c("0","1"),
                      labels = c("Dry","Flowing"), name = "", drop = FALSE,
                      na.value = "white", guide = "none"),
    scale_y_continuous("Activation / Drying order",
                       limits = c(0.5, global_n + 0.5),
                       expand = c(0, 0),
                       breaks = function(lims) pretty(lims, n = 5)),
    scale_x_datetime(date_breaks = "1 day", date_labels = "%m/%d"),
    theme_classic() +
      theme(legend.position = "none",
            strip.background = element_blank(),
            panel.background = element_rect(fill = "white", colour = NA),
            axis.title.y = element_text(angle = 90, size = 12),
            axis.text    = element_text(size = 12),
            plot.title   = element_text(size = 9, hjust = 0.5, face = "bold"),
            plot.margin  = unit(c(2, 2, 2, 2), "pt")))
  
  # -- hydro panel with NO LEGEND ------------------------------------------
  build_hydro <- function(ev){
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
    ev_water <- hydro_list[[as.character(ev)]]
    ev_water$.q <- sec_global$fwd(ev_water[[discharge_col]])
    
    ggplot(ev_water, aes(x = datetime)) +
      geom_col(aes(y = hourly_precip, fill = "Precipitation"), 
               width = 3600, show.legend = FALSE) +
      geom_line(aes(y = .q, color = "Discharge"), 
                linewidth = 0.6, show.legend = FALSE) +
      scale_fill_manual(values = c("Precipitation" = "grey40"), name = "", guide = "none") +
      scale_color_manual(values = c("Discharge" = "black"), name = "", guide = "none") +
      scale_y_continuous("Precip (mm)", limits = c(0, global_ytop), expand = c(0, 0),
                         sec.axis = sec_axis(~ sec_global$rev(.), name = "Q (mm/day)")) +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", title = paste("Event", ev)) +
      theme_classic() +
      theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
            axis.text.y = element_text(size = 12),
            axis.title.y = element_text(angle = 90, size = 12),
            axis.title.y.right = element_text(angle = 90, size = 12),
            plot.title = element_text(size = 9, hjust = 0.5, face = "bold"),
            plot.margin = unit(c(2, 2, 2, 2), "pt"),
            legend.position = "none")
  }
  
  build_wet_tiles <- function(ev){
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
    ev_ranks <- event_specific_order(ev_dat)
    if(nrow(ev_ranks) == 0) stop("Event ", ev, " has no sustained wetting sequence.")
    ev_full <- ev_dat %>%
      inner_join(ev_ranks, by = "ID") %>%
      mutate(y_rank = event_rank) %>%
      select(datetime, y_rank, binary) %>%
      ungroup() %>%
      complete(datetime, y_rank = 1:global_n)
    ggplot(ev_full, aes(x = datetime, y = y_rank)) +
      tile_layer() +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", y = "Wetting order (1 = wets first)")
  }
  
  build_wet_cons <- function(ev){
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
    ev_full <- ev_dat %>%
      inner_join(cons_wet, by = "ID") %>%
      mutate(y_rank = consensus_rank) %>%
      select(datetime, y_rank, binary) %>%
      ungroup() %>%
      complete(datetime, y_rank = 1:global_n)
    ggplot(ev_full, aes(x = datetime, y = y_rank)) +
      tile_layer() +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", y = "Consensus wetting",
           title = paste("Event", ev, "- consensus wetting"))
  }
  
  build_dry_tiles <- function(ev){
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
    dry_ranks <- event_drying_rank(ev_dat)
    if(nrow(dry_ranks) == 0) stop("Event ", ev, " has no drying sequence.")
    n_dry <- nrow(dry_ranks)
    ev_full <- ev_dat %>%
      inner_join(dry_ranks, by = "ID") %>%
      mutate(y_rank = n_dry - dry_rank + 1) %>%
      select(datetime, y_rank, binary) %>%
      ungroup() %>%
      complete(datetime, y_rank = 1:global_n)
    ggplot(ev_full, aes(x = datetime, y = y_rank)) +
      tile_layer() +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", y = "Drying order (N = dries first)")
  }
  
  build_dry_cons <- function(ev){
    ev_dat <- dat_all %>% filter(event_id == ev)
    ev_s <- min(ev_dat$datetime); ev_e <- max(ev_dat$datetime)
    ev_full <- ev_dat %>%
      inner_join(cons_dry, by = "ID") %>%
      mutate(y_rank = max_cons_dry - consensus_rank + 1) %>%
      select(datetime, y_rank, binary) %>%
      ungroup() %>%
      complete(datetime, y_rank = 1:global_n)
    ggplot(ev_full, aes(x = datetime, y = y_rank)) +
      tile_layer() +
      coord_cartesian(xlim = c(ev_s, ev_e), expand = FALSE) +
      labs(x = "", y = "Consensus drying (N = dries first)",
           title = paste("Event", ev, "- consensus drying"))
  }
  
  if(has_hydro){
    plots <- list(
      build_hydro(event_wet),    build_hydro(event_dry),
      build_wet_tiles(event_wet), build_dry_tiles(event_dry),
      build_wet_cons(event_wet),  build_dry_cons(event_dry)
    )
    heights <- c(hydro_height, tile_height, tile_height)
  } else {
    plots <- list(
      build_wet_tiles(event_wet), build_dry_tiles(event_dry),
      build_wet_cons(event_wet),  build_dry_cons(event_dry)
    )
    heights <- c(tile_height, tile_height)
  }
  
  wrap_plots(plots, ncol = 2, byrow = TRUE) +
    plot_layout(heights = heights, widths = 1) +   # removed guides = "collect"
    plot_annotation(
      title = paste0(shed, ": wetting vs. drying activation order"),
      caption = paste(
        "Left column: wetting sequence (1 = wets first, bottom).",
        "Right column: drying sequence (N = dries first, top).",
        "All tile panels share the same y-axis extent; missing rows are white."
      )
    )
}

figure_combo <- fig2_wet_dry_columns(
  shed          = "W3",
  input         = input_w3,
  event_df      = w3_events_all,
  consensus_wet = avg_event_w3,
  consensus_dry = dry_w3,
  ids           = W3_IDs,
  event_wet     = 7,
  event_dry     = 10,
  precip        = precip1,
  discharge     = w3_q_seasonal,
  discharge_col = "Q_mm_day"
)

ggsave("figure2_wet_vs_dry.pdf", figure_combo, width = 6.5, units ="in")
```


# Figure 3 — Ordered deviation heatmap: consensus vs. each event's sequence

```{r fig3-heatmap}
#| message: false
#| warning: false
#| fig-width: 11
#| fig-height: 12
# RQ2. First column = consensus order; each remaining column = one event's
# activation order. Cell colour = that event's rank minus the consensus rank
# (blue = earlier than consensus, red = later). Nodes that have no sustained
# flow in an event (missing from that sequence) are drawn in tan. One row of
# panels per watershed. Only the EVENT-BASED sequences are shown (no FP/TWI/BFS).

dev_heatmap_data <- function(consensus, event_seqs, shed){
  ref <- consensus %>% transmute(ID, ref_rank = sequence)
  ev  <- event_seqs %>% transmute(col = paste0("E", event_id), ord = event_id,
                                  ID, rank = Sequence)
  cons <- ref %>% transmute(col = "Consensus", ord = -1, ID, rank = ref_rank)
  bind_rows(cons, ev) %>%
    complete(nesting(col, ord), ID) %>%      # fill grid so missing nodes show
    left_join(ref, by = "ID") %>%
    mutate(dev = rank - ref_rank, shed = shed)
}

plot_dev_heatmap <- function(d, shed){
  ref_order <- d %>% filter(col == "Consensus") %>% arrange(ref_rank) %>%
    pull(ID) %>% unique() %>% as.character()
  col_order <- d %>% distinct(col, ord) %>% arrange(ord) %>% pull(col)
  d <- d %>% mutate(ID = factor(as.character(ID), levels = ref_order),
                    col = factor(col, levels = col_order))
  ggplot(d, aes(col, ID)) +
    geom_tile(data = filter(d, is.na(dev)), fill = "#E8D8A0", color = "white") +
    geom_tile(data = filter(d, !is.na(dev)), aes(fill = dev), color = "white") +
    scale_fill_gradient2(low = "#2166AC", mid = "grey95", high = "#B2182B",
                         midpoint = 0, name = "rank −\nconsensus") +
    labs(x = "", y = "Node (consensus order; wets first at bottom)", title = shed) +
    theme_minimal() +
    theme(panel.grid = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
          axis.text.y = element_text(size = 6),
          plot.title = element_text(face = "bold"))
}

h_w3 <- plot_dev_heatmap(dev_heatmap_data(avg_event_w3, event_seqsW3, "W3"), "W3")
h_fb <- plot_dev_heatmap(dev_heatmap_data(avg_event_fb, event_seqsFB, "FB"), "FB")
h_zz <- plot_dev_heatmap(dev_heatmap_data(avg_event_zz, event_seqsZZ, "ZZ"), "ZZ")

figure3 <- (h_w3 / h_fb / h_zz) + plot_layout(guides = "collect") +
  plot_annotation(caption = "Tan cells = node had no sustained flow in that event (missing from its sequence).")
figure3
ggsave("figure3_deviation_heatmap.pdf", figure3, width = 11, height = 12)
```
```{r}
#| message: false
#| warning: false
#| fig-width: 11
#| fig-height: 12
# Ordered deviation heatmap: consensus column FIRST + one column per kept event.
# Events with data for < half the deployed sensors are dropped (per watershed);
# `drop_events` removes specific events manually. Missing cells = medium-dark grey.

dev_heatmap_data <- function(consensus, event_seqs, input, event_df, ids, shed,
                             drop_events = NULL){
  data_counts <- input %>% filter(ID %in% ids) %>%
    inner_join(event_df %>% select(datetime, event_id, event_type), by = "datetime") %>%
    filter(!is.na(event_id), event_type != "baseflow") %>%
    group_by(event_id) %>% summarise(n_with_data = n_distinct(ID), .groups = "drop")

  keep_events <- data_counts %>%
    filter(n_with_data >= length(ids) / 2) %>% pull(event_id)
  keep_events <- setdiff(keep_events, drop_events)         # manual removal
  message(shed, ": ", length(keep_events), " events kept; dropped manually: ",
          paste(drop_events, collapse = ", "))

  ref  <- consensus %>% transmute(ID, ref_rank = sequence)
  ev   <- event_seqs %>% filter(event_id %in% keep_events) %>%
    transmute(col = paste0("E", event_id), ord = as.numeric(event_id),
              ID, rank = as.numeric(Sequence))
  cons <- ref %>% transmute(col = "Consensus", ord = -Inf,
                            ID, rank = as.numeric(ref_rank))

  bind_rows(cons, ev) %>%
    complete(nesting(col, ord), ID) %>%
    left_join(ref, by = "ID") %>%
    mutate(dev = rank - ref_rank, shed = shed)
}

plot_dev_heatmap <- function(d, shed){
  ref_order  <- d %>% filter(col == "Consensus") %>% arrange(ref_rank) %>%
    pull(ID) %>% unique() %>% as.character()
  event_cols <- d %>% filter(col != "Consensus") %>% distinct(col, ord) %>%
    arrange(ord) %>% pull(col)
  col_order  <- c("Consensus", event_cols)

  d <- d %>% mutate(ID = factor(as.character(ID), levels = ref_order))

  ggplot(d, aes(col, ID)) +
    geom_tile(data = filter(d, is.na(dev)), fill = "grey40", color = "white") +
    geom_tile(data = filter(d, !is.na(dev)), aes(fill = dev), color = "white") +
    scale_x_discrete(limits = col_order) +                 # <- forces consensus first
    scale_fill_gradient2(low = "#2166AC", mid = "grey95", high = "#B2182B",
                         midpoint = 0, name = "rank −\nconsensus") +
    labs(x = "", y = "Node (consensus order; wets first at bottom)", title = shed) +
    theme_minimal() +
    theme(panel.grid = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
          axis.text.y = element_text(size = 6),
          plot.title = element_text(face = "bold"))
}

h_w3 <- plot_dev_heatmap(dev_heatmap_data(avg_event_w3, event_seqsW3, input_w3, w3_events_all, W3_IDs, "W3"), "W3")
h_fb <- plot_dev_heatmap(dev_heatmap_data(avg_event_fb, event_seqsFB, input_fb, FBall_events, FB_IDs, "FB"), "FB")
h_zz <- plot_dev_heatmap(dev_heatmap_data(avg_event_zz, event_seqsZZ, input_zz, ZZall_events, ZZ_IDs, "ZZ",
                                          drop_events = c(9, 14)), "ZZ")

figure3 <- (h_w3 / h_fb / h_zz) + plot_layout(guides = "collect") +
  plot_annotation(caption = "Events with data for fewer than half the deployed sensors, plus manually flagged events, are removed. Medium-dark grey = node missing from that event's sequence.")
figure3
ggsave("figure3_deviation_heatmap.pdf", figure3, width = 6.5, units = "in")
```
```{r}
#| message: false
#| warning: false
#| fig-width: 6.5
#| fig-height: 9
# Ordered deviation heatmap: consensus column FIRST + one column per kept event.
# Events with data for < half the deployed sensors are dropped (per watershed);
# `drop_events` removes specific events manually. Missing cells = medium-dark grey.
# Sized for an 8.5x11 page (<= 6.5 in wide, <= 9 in tall, all text >= 10 pt).

dev_heatmap_data <- function(consensus, event_seqs, input, event_df, ids, shed,
                             drop_events = NULL){
  data_counts <- input %>% filter(ID %in% ids) %>%
    inner_join(event_df %>% select(datetime, event_id, event_type), by = "datetime") %>%
    filter(!is.na(event_id), event_type != "baseflow") %>%
    group_by(event_id) %>% summarise(n_with_data = n_distinct(ID), .groups = "drop")

  keep_events <- data_counts %>%
    filter(n_with_data >= length(ids) / 2) %>% pull(event_id)
  keep_events <- setdiff(keep_events, drop_events)         # manual removal
  message(shed, ": ", length(keep_events), " events kept; dropped manually: ",
          paste(drop_events, collapse = ", "))

  ref  <- consensus %>% transmute(ID, ref_rank = sequence)
  ev   <- event_seqs %>% filter(event_id %in% keep_events) %>%
    transmute(col = paste0("E", event_id), ord = as.numeric(event_id),
              ID, rank = as.numeric(Sequence))
  cons <- ref %>% transmute(col = "Consensus", ord = -Inf,
                            ID, rank = as.numeric(ref_rank))

  bind_rows(cons, ev) %>%
    complete(nesting(col, ord), ID) %>%
    left_join(ref, by = "ID") %>%
    mutate(dev = rank - ref_rank, shed = shed)
}

plot_dev_heatmap <- function(d, shed){
  ref_order  <- d %>% filter(col == "Consensus") %>% arrange(ref_rank) %>%
    pull(ID) %>% unique() %>% as.character()
  event_cols <- d %>% filter(col != "Consensus") %>% distinct(col, ord) %>%
    arrange(ord) %>% pull(col)
  col_order  <- c("Consensus", event_cols)

  d <- d %>% mutate(ID = factor(as.character(ID), levels = ref_order))

  ggplot(d, aes(col, ID)) +
    geom_tile(data = filter(d, is.na(dev)), fill = "grey40", color = "white") +
    geom_tile(data = filter(d, !is.na(dev)), aes(fill = dev), color = "white") +
    scale_x_discrete(limits = col_order) +                 # forces consensus first
    scale_fill_gradient2(high = "#2166AC", mid = "grey95", low = "#B2182B",
                         midpoint = 0, name = "rank −\nconsensus") +
    labs(x = "", y = "Node (consensus order; wets first at bottom)", title = shed) +
    theme_minimal(base_size = 11) +
    theme(panel.grid       = element_blank(),
          axis.text.x      = element_text(angle = 45, hjust = 1, size = 10),
          axis.text.y      = element_text(size = 10),
          axis.title       = element_text(size = 11),
          plot.title       = element_text(size = 12, face = "bold"),
          legend.title     = element_text(size = 10),
          legend.text      = element_text(size = 10))
}

h_w3 <- plot_dev_heatmap(dev_heatmap_data(avg_event_w3, event_seqsW3, input_w3, w3_events_all, W3_IDs, "W3"), "W3")
h_fb <- plot_dev_heatmap(dev_heatmap_data(avg_event_fb, event_seqsFB, input_fb, FBall_events, FB_IDs, "FB"), "FB")
h_zz <- plot_dev_heatmap(dev_heatmap_data(avg_event_zz, event_seqsZZ, input_zz, ZZall_events, ZZ_IDs, "ZZ",
                                          drop_events = c(9, 14)), "ZZ")

figure3 <- (h_w3 / h_fb / h_zz) + plot_layout(guides = "collect") +
  plot_annotation(
    caption = "Events with data for fewer than half the deployed sensors, plus manually flagged events, are removed. Medium-dark grey = node missing from that event's sequence.")
figure3
ggsave("figure3_deviation_heatmap.pdf", figure3, width = 6.5, height = 9)
```

```{r classified by whether it changes or not during event}
#| message: false
#| warning: false
#| fig-width: 6.5
#| fig-height: 9
# Ordered deviation heatmap: consensus column FIRST + one column per kept event.
# Cells now reflect sensor behaviour during each event:
#   * changed state (activated)      -> coloured by rank - consensus rank
#   * no change, stayed flowing      -> solid green
#   * no change, stayed dry          -> solid orange (the true gaps in a partial seq.)
#   * no observations / not deployed -> medium-dark grey
# Events with data for < half the deployed sensors are dropped; `drop_events`
# removes specific events. Sized for an 8.5x11 page (<=6.5 in, all text >=10 pt).

dev_heatmap_data <- function(consensus, event_seqs, input, event_df, ids, shed,
                             drop_events = NULL){
  ev_obs <- input %>% filter(ID %in% ids) %>%
    inner_join(event_df %>% select(datetime, event_id, event_type), by = "datetime") %>%
    filter(!is.na(event_id), event_type != "baseflow")

  data_counts <- ev_obs %>% group_by(event_id) %>%
    summarise(n_with_data = n_distinct(ID), .groups = "drop")
  keep_events <- data_counts %>%
    filter(n_with_data >= length(ids) / 2) %>% pull(event_id)
  keep_events <- setdiff(keep_events, drop_events)
  message(shed, ": ", length(keep_events), " events kept; dropped manually: ",
          paste(drop_events, collapse = ", "))

  # per (event, node) state: did it change, stay wet, or stay dry?
  states <- ev_obs %>% filter(event_id %in% keep_events) %>%
    group_by(event_id, ID) %>%
    summarise(n = n(), n_wet = sum(binary == 1), .groups = "drop") %>%
    mutate(state = case_when(n_wet == 0 ~ "always_dry",
                             n_wet == n ~ "always_wet",
                             TRUE       ~ "changed"))

  ref     <- consensus %>% transmute(ID, ref_rank = sequence)
  ev_rank <- event_seqs %>% filter(event_id %in% keep_events) %>%
    transmute(event_id, ID, rank = as.numeric(Sequence))

  ev <- states %>%
    left_join(ev_rank, by = c("event_id", "ID")) %>%
    transmute(col = paste0("E", event_id), ord = as.numeric(event_id), ID, rank, state)
  cons <- ref %>% transmute(col = "Consensus", ord = -Inf, ID,
                            rank = as.numeric(ref_rank), state = "changed")

  bind_rows(cons, ev) %>%
    complete(nesting(col, ord), ID) %>%             # unobserved cells -> NA
    left_join(ref, by = "ID") %>%
    mutate(dev   = rank - ref_rank,
           state = ifelse(is.na(state), "no_data", state),
           shed  = shed)
}

plot_dev_heatmap <- function(d, shed){
  ref_order  <- d %>% filter(col == "Consensus") %>% arrange(ref_rank) %>%
    pull(ID) %>% unique() %>% as.character()
  event_cols <- d %>% filter(col != "Consensus") %>% distinct(col, ord) %>%
    arrange(ord) %>% pull(col)
  col_order  <- c("Consensus", event_cols)

  d <- d %>% mutate(ID = factor(as.character(ID), levels = ref_order))
  d_changed <- filter(d, state == "changed")
  d_cat     <- filter(d, state != "changed")

  state_cols <- c(always_wet = "grey5", always_dry = "#DB995A", no_data = "grey40")
  state_labs <- c(always_wet = "No change: flowing",
                  always_dry = "No change: dry",
                  no_data    = "No data")

  ggplot() +
    geom_tile(data = d_cat, aes(col, ID, fill = state), color = "white") +
    scale_fill_manual(values = state_cols, labels = state_labs,
                      breaks = c("always_wet", "always_dry", "no_data"),
                      name = "Sensor state", drop = FALSE) +
    ggnewscale::new_scale_fill() +
    geom_tile(data = d_changed, aes(col, ID, fill = dev), color = "white") +
    scale_fill_gradient2(low = "#B2182B", mid = "grey95", high = "#2166AC",
                         midpoint = 0, name = "rank −\nconsensus") +
    scale_x_discrete(limits = col_order) +
    labs(x = "", y = "Node (consensus order; wets first at bottom)", title = shed) +
    theme_minimal(base_size = 11) +
    theme(panel.grid   = element_blank(),
          axis.text.x  = element_text(angle = 45, hjust = 1, size = 10),
          axis.text.y  = element_text(size = 10),
          axis.title   = element_text(size = 11),
          plot.title   = element_text(size = 12, face = "bold"),
          legend.title = element_text(size = 10),
          legend.text  = element_text(size = 10))
}

h_w3 <- plot_dev_heatmap(dev_heatmap_data(avg_event_w3, event_seqsW3, input_w3, w3_events_all, W3_IDs, "W3"), "W3")
h_fb <- plot_dev_heatmap(dev_heatmap_data(avg_event_fb, event_seqsFB, input_fb, FBall_events, FB_IDs, "FB"), "FB")
h_zz <- plot_dev_heatmap(dev_heatmap_data(avg_event_zz, event_seqsZZ, input_zz, ZZall_events, ZZ_IDs, "ZZ",
                                          drop_events = c(9, 14)), "ZZ")

figure3 <- (h_w3 / h_fb / h_zz) + plot_layout(guides = "collect") +
  plot_annotation(
    caption = "Coloured cells = sensor activated (rank vs. consensus). Green = stayed flowing, orange = stayed dry (a gap in the partial sequence), grey = no data.",
    theme = theme(plot.caption = element_text(size = 10)))
figure3
ggsave("figure3_deviation_heatmap.pdf", figure3, width = 6.5, height = 9)
```
```{r}
#| message: false
#| warning: false
#| fig-width: 6.5
#| fig-height: 9
# No ggpattern needed: categorical states shown as glyphs over neutral tiles.

dev_heatmap_data <- function(consensus, event_seqs, input, event_df, ids, shed,
                             drop_events = NULL){
  ev_obs <- input %>% filter(ID %in% ids) %>%
    inner_join(event_df %>% select(datetime, event_id, event_type), by = "datetime") %>%
    filter(!is.na(event_id), event_type != "baseflow")

  data_counts <- ev_obs %>% group_by(event_id) %>%
    summarise(n_with_data = n_distinct(ID), .groups = "drop")
  keep_events <- data_counts %>%
    filter(n_with_data >= length(ids) / 2) %>% pull(event_id)
  keep_events <- setdiff(keep_events, drop_events)
  message(shed, ": ", length(keep_events), " events kept; dropped manually: ",
          paste(drop_events, collapse = ", "))

  states <- ev_obs %>% filter(event_id %in% keep_events) %>%
    group_by(event_id, ID) %>%
    summarise(n = n(), n_wet = sum(binary == 1), .groups = "drop") %>%
    mutate(state = case_when(n_wet == 0 ~ "always_dry",
                             n_wet == n ~ "always_wet",
                             TRUE       ~ "changed"))

  ref     <- consensus %>% transmute(ID, ref_rank = sequence)
  ev_rank <- event_seqs %>% filter(event_id %in% keep_events) %>%
    transmute(event_id, ID, rank = as.numeric(Sequence))

  ev <- states %>%
    left_join(ev_rank, by = c("event_id", "ID")) %>%
    transmute(col = paste0("E", event_id), ord = as.numeric(event_id), ID, rank, state)
  cons <- ref %>% transmute(col = "Consensus", ord = -Inf, ID,
                            rank = as.numeric(ref_rank), state = "changed")

  bind_rows(cons, ev) %>%
    complete(nesting(col, ord), ID) %>%
    left_join(ref, by = "ID") %>%
    mutate(dev   = rank - ref_rank,
           state = ifelse(is.na(state), "no_data", state),
           shed  = shed)
}

plot_dev_heatmap <- function(d, shed){
  ref_order  <- d %>% filter(col == "Consensus") %>% arrange(ref_rank) %>%
    pull(ID) %>% unique() %>% as.character()
  event_cols <- d %>% filter(col != "Consensus") %>% distinct(col, ord) %>%
    arrange(ord) %>% pull(col)
  col_order  <- c("Consensus", event_cols)

  d <- d %>% mutate(ID = factor(as.character(ID), levels = ref_order))
  d_changed <- filter(d, state == "changed")
  d_cat     <- filter(d, state != "changed") %>%
    mutate(state = factor(state, levels = c("always_wet", "always_dry", "no_data")))

  glyphs <- c(always_wet = 16, always_dry = 1, no_data = 4)   # filled dot, open dot, X
  glabs  <- c(always_wet = "No change: flowing",
              always_dry = "No change: dry",
              no_data    = "No data")

  ggplot() +
    geom_tile(data = d_cat, aes(col, ID), fill = "grey92", colour = "white") +
    geom_point(data = d_cat, aes(col, ID, shape = state), size = 1, stroke = 0.6) +
    scale_shape_manual(values = glyphs, labels = glabs,
                       breaks = c("always_wet", "always_dry", "no_data"),
                       name = "Sensor state") +
    geom_tile(data = d_changed, aes(col, ID, fill = dev), colour = "white") +
    scale_fill_gradient2(low = "#2166AC", mid = "grey95", high = "#B2182B",
                         midpoint = 0, name = "rank −\nconsensus") +
    scale_x_discrete(limits = col_order) +
    labs(x = "", y = "Node (consensus order; wets first at bottom)", title = shed) +
    theme_minimal(base_size = 11) +
    theme(panel.grid   = element_blank(),
          axis.text.x  = element_text(size = 8),
          axis.text.y  = element_text(size = 8),
          axis.title   = element_text(size = 11),
          plot.title   = element_text(size = 12, face = "bold"),
          legend.title = element_text(size = 10),
          legend.text  = element_text(size = 10))
}

h_w3 <- plot_dev_heatmap(dev_heatmap_data(avg_event_w3, event_seqsW3, input_w3, w3_events_all, W3_IDs, "W3"), "W3")
h_fb <- plot_dev_heatmap(dev_heatmap_data(avg_event_fb, event_seqsFB, input_fb, FBall_events, FB_IDs, "FB"), "FB")
h_zz <- plot_dev_heatmap(dev_heatmap_data(avg_event_zz, event_seqsZZ, input_zz, ZZall_events, ZZ_IDs, "ZZ",
                                          drop_events = c(9, 14)), "ZZ")

figure3 <- (h_w3 / h_fb / h_zz) + plot_layout(guides = "collect") +
  plot_annotation(
    caption = "Coloured cells = sensor activated (rank vs. consensus). Filled dot = stayed flowing, open dot = stayed dry (a gap in the partial sequence), X = no data.",
    theme = theme(plot.caption = element_text(size = 10)))
figure3
ggsave("figure3_deviation_heatmap.pdf", figure3, width = 6.5, height = 9)
```

```{r}
# output the unique events and their year, so that I can make a line on big chart
w3_events_all %>% 
  group_by(event_id) %>% 
  summarise(year = year(datetime)) %>% 
  unique()

FBall_events%>% 
  group_by(event_id) %>% 
  summarise(year = year(datetime)) %>% 
  unique() %>% as.data.frame()

ZZall_events%>% 
  group_by(event_id) %>% 
  summarise(year = year(datetime)) %>% 
  unique()%>% as.data.frame()
```

New Version, comparing all sequences, differentiating between wetting and drying, and showing only 2 for each watershed
```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 9
# One event per watershed. Y-axis = that event's OBSERVED order (wetting; drying
# reversed so 1 = dries last). Columns = each TEST sequence (Flow Persistence,
# TWI, Consensus) compared to the observed WETTING and DRYING order.
# Cell colour = test rank - observed rank, BOTH re-ranked on their shared nodes.

# >>> pick ONE event per watershed to display <
pick_event <- list(W3 = 12, FB = 14, ZZ = 8)

# ---- print the eligible (kept) events from the earlier curation -------------
wetting_eligible <- function(spec, u) sort(setdiff(u, c(spec$remove, spec$drying_only)))
drying_eligible  <- function(spec, u) sort(setdiff(u, c(spec$remove, spec$wetting_only)))
keep_trans <- list(
  W3 = events_with_transitions(input_w3, w3_events_all, W3_IDs, 5),
  FB = events_with_transitions(input_fb, FBall_events,  FB_IDs, 5),
  ZZ = events_with_transitions(input_zz, ZZall_events,  ZZ_IDs, 5))
keep_both <- Map(function(s, u) sort(setdiff(u, c(s$remove, s$wetting_only, s$drying_only))),
                 dir_spec, universe)
keep_both <- Map(intersect, keep_both, keep_trans)
for (s in names(keep_both))
  message(s, " events with BOTH valid wetting & drying (choose from these): ",
          paste(keep_both[[s]], collapse = ", "))

# ---- observed ranks for one event (wetting; drying reversed) ----------------
obs_rank <- function(dir_long, e, reverse = FALSE){
  r <- dir_long %>% filter(event_id == e) %>%
    mutate(rank = row_number()) %>% select(ID = sensor_id, rank)
  if (reverse && nrow(r) > 0) r <- r %>% mutate(rank = max(rank) - rank + 1)
  r
}

# ---- build deviation data for one watershed's chosen event ------------------
# Both the observed order and the test sequence are RE-RANKED on their shared
# nodes (1..k) before differencing, so dev reflects relative position only.
test_vs_event_data <- function(models, ds, e, shed){
  obs_wet <- obs_rank(ds$wetting, e)                  # 1 = wets first
  obs_dry <- obs_rank(ds$drying,  e, reverse = TRUE)  # 1 = dries last
  refs <- list(wet = obs_wet, dry = obs_dry)

  dev_df <- purrr::imap_dfr(models, function(mseq, mname){
    ms <- mseq %>% transmute(ID, mrank = sequence)
    purrr::imap_dfr(refs, function(obs, dir){
      obs %>% inner_join(ms, by = "ID") %>%
        mutate(obs_r  = rank(rank,  ties.method = "first"),   # re-rank on shared set
               test_r = rank(mrank, ties.method = "first")) %>%
        transmute(ID, dev = test_r - obs_r,
                  col = paste0(mname, "\n", ifelse(dir == "wet", "wetting", "drying")),
                  ord = match(mname, names(models)) * 2 - (dir == "wet"))
    })
  })

  # y-axis order = this event's observed wetting order, re-ranked 1..k ---------
  yaxis <- obs_wet %>% mutate(obs_rank = rank(rank, ties.method = "first")) %>%
    select(ID, obs_rank)

  dev_df %>% left_join(yaxis, by = "ID") %>%
    mutate(shed = shed, event = e)
}

plot_test_vs_event <- function(d, shed, e){
  ref_order <- d %>% distinct(ID, obs_rank) %>% arrange(obs_rank) %>%
    pull(ID) %>% as.character()
  col_order <- d %>% distinct(col, ord) %>% arrange(ord) %>% pull(col)
  d <- d %>% mutate(ID = factor(as.character(ID), levels = ref_order),
                    col = factor(col, levels = col_order))
  ggplot(d, aes(col, ID, fill = dev)) +
    geom_tile(colour = "white") +
    scale_fill_gradient2(low = "#2166AC", mid = "grey95", high = "#B2182B",
                         midpoint = 0, name = "test −\nobserved") +
    labs(x = "", y = "Node (observed wetting order; wets first at bottom)",
         title = paste0(shed, " — event ", e)) +
    theme_minimal(base_size = 11) +
    theme(panel.grid = element_blank(),
          axis.text.x = element_text(size = 8, angle = 30, hjust = 1),
          axis.text.y = element_text(size = 8),
          axis.title = element_text(size = 11),
          plot.title = element_text(size = 12, face = "bold"),
          legend.title = element_text(size = 10), legend.text = element_text(size = 10))
}

models_w3 <- list("Flow Persistence" = W3_pk_seq, "TWI" = w3_twi_sequence, "Consensus" = avg_event_w3)
models_fb <- list("Flow Persistence" = FB_pk_seq, "TWI" = fb_twi_sequence, "Consensus" = avg_event_fb)
models_zz <- list("Flow Persistence" = ZZ_pk_seq, "TWI" = zz_twi_sequence, "Consensus" = avg_event_zz)

h_w3 <- plot_test_vs_event(test_vs_event_data(models_w3, ds_w3, pick_event$W3, "W3"), "W3", pick_event$W3)
h_fb <- plot_test_vs_event(test_vs_event_data(models_fb, ds_fb, pick_event$FB, "FB"), "FB", pick_event$FB)
h_zz <- plot_test_vs_event(test_vs_event_data(models_zz, ds_zz, pick_event$ZZ, "ZZ"), "ZZ", pick_event$ZZ)

figure3 <- (h_w3 / h_fb / h_zz) + plot_layout(guides = "collect") +
  plot_annotation(
    caption = "Each panel: one event. Columns = test sequences (Flow Persistence, TWI, Consensus) vs. that event's observed wetting and (reversed) drying order, both re-ranked on shared nodes. Colour = test rank minus observed rank; grey = agreement, blue = test places node earlier, red = later.",
    theme = theme(plot.caption = element_text(size = 9)))
figure3
ggsave("figure3_test_vs_event_heatmap.pdf", figure3, width = 8, height = 9)
```
Version with slightly different symbology:
```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 9
# One event per watershed. Y-axis = that event's OBSERVED order (wetting; drying
# reversed so 1 = dries last). Columns = each TEST sequence vs. observed wetting
# and drying order, re-ranked on shared nodes.
# Fill = |test rank - observed rank| on a black->white greyscale (black = exact
# agreement). Direction shown by a texture overlay:
#   diagonal "/"  = test ranks the node LATER than observed (test_r > obs_r)
#   diagonal "\"  = test ranks the node EARLIER than observed (test_r < obs_r)
#   (no line)     = exact agreement (dev = 0)

# >>> pick ONE event per watershed to display <
pick_event <- list(W3 = 12, FB = 14, ZZ = 8)

# ---- print the eligible (kept) events from the earlier curation -------------
wetting_eligible <- function(spec, u) sort(setdiff(u, c(spec$remove, spec$drying_only)))
drying_eligible  <- function(spec, u) sort(setdiff(u, c(spec$remove, spec$wetting_only)))
keep_trans <- list(
  W3 = events_with_transitions(input_w3, w3_events_all, W3_IDs, 5),
  FB = events_with_transitions(input_fb, FBall_events,  FB_IDs, 5),
  ZZ = events_with_transitions(input_zz, ZZall_events,  ZZ_IDs, 5))
keep_both <- Map(function(s, u) sort(setdiff(u, c(s$remove, s$wetting_only, s$drying_only))),
                 dir_spec, universe)
keep_both <- Map(intersect, keep_both, keep_trans)
for (s in names(keep_both))
  message(s, " events with BOTH valid wetting & drying (choose from these): ",
          paste(keep_both[[s]], collapse = ", "))

# ---- observed ranks for one event (wetting; drying reversed) ----------------
obs_rank <- function(dir_long, e, reverse = FALSE){
  r <- dir_long %>% filter(event_id == e) %>%
    mutate(rank = row_number()) %>% select(ID = sensor_id, rank)
  if (reverse && nrow(r) > 0) r <- r %>% mutate(rank = max(rank) - rank + 1)
  r
}

# ---- deviation data; both sequences re-ranked on shared nodes ---------------
test_vs_event_data <- function(models, ds, e, shed){
  obs_wet <- obs_rank(ds$wetting, e)                  # 1 = wets first
  obs_dry <- obs_rank(ds$drying,  e, reverse = TRUE)  # 1 = dries last
  refs <- list(wet = obs_wet, dry = obs_dry)

  dev_df <- purrr::imap_dfr(models, function(mseq, mname){
    ms <- mseq %>% transmute(ID, mrank = sequence)
    purrr::imap_dfr(refs, function(obs, dir){
      obs %>% inner_join(ms, by = "ID") %>%
        mutate(obs_r  = rank(rank,  ties.method = "first"),
               test_r = rank(mrank, ties.method = "first")) %>%
        transmute(ID,
                  dev     = test_r - obs_r,
                  abs_dev = abs(test_r - obs_r),
                  dirn    = dplyr::case_when(dev > 0 ~ "later",
                                             dev < 0 ~ "earlier",
                                             TRUE    ~ "equal"),
                  col = paste0(mname, "\n", ifelse(dir == "wet", "wetting", "drying")),
                  ord = match(mname, names(models)) * 2 - (dir == "wet"))
    })
  })

  yaxis <- obs_wet %>% mutate(obs_rank = rank(rank, ties.method = "first")) %>%
    select(ID, obs_rank)
  dev_df %>% left_join(yaxis, by = "ID") %>% mutate(shed = shed, event = e)
}

# ---- plot: greyscale |dev| fill + diagonal line texture for direction -------
plot_test_vs_event <- function(d, shed, e){
  ref_order <- d %>% distinct(ID, obs_rank) %>% arrange(obs_rank) %>%
    pull(ID) %>% as.character()
  col_order <- d %>% distinct(col, ord) %>% arrange(ord) %>% pull(col)
  d <- d %>% mutate(ID  = factor(as.character(ID), levels = ref_order),
                    colf = factor(col, levels = col_order),
                    x    = as.integer(colf),
                    y    = as.integer(ID))

  seg <- d %>% filter(dirn != "equal") %>%
    mutate(x0 = x - 0.45, x1 = x + 0.45,
           # "/" for later (bottom-left to top-right), "\" for earlier
           y0 = ifelse(dirn == "later", y - 0.45, y + 0.45),
           y1 = ifelse(dirn == "later", y + 0.45, y - 0.45))

  ggplot(d, aes(x, y)) +
    geom_tile(aes(fill = abs_dev), colour = "grey85") +
    geom_segment(data = seg, aes(x = x0, xend = x1, y = y0, yend = y1),
                 colour = "grey55", linewidth = 0.3) +
    scale_fill_gradient(low = "black", high = "white", name = "|test −\nobserved|",
                        limits = c(0, max(d$abs_dev, na.rm = TRUE))) +
    scale_x_continuous(breaks = seq_along(col_order), labels = col_order,
                       expand = c(0, 0)) +
    scale_y_continuous(breaks = seq_along(ref_order), labels = ref_order,
                       expand = c(0, 0)) +
    labs(x = "", y = "Node (observed wetting order; wets first at bottom)",
         title = paste0(shed, " — event ", e)) +
    theme_minimal(base_size = 11) +
    theme(panel.grid = element_blank(),
          axis.text.x = element_text(size = 8, angle = 30, hjust = 1),
          axis.text.y = element_text(size = 8),
          axis.title = element_text(size = 11),
          plot.title = element_text(size = 12, face = "bold"),
          legend.title = element_text(size = 10), legend.text = element_text(size = 10))
}

models_w3 <- list("Flow Persistence" = W3_pk_seq, "TWI" = w3_twi_sequence, "Consensus" = avg_event_w3)
models_fb <- list("Flow Persistence" = FB_pk_seq, "TWI" = fb_twi_sequence, "Consensus" = avg_event_fb)
models_zz <- list("Flow Persistence" = ZZ_pk_seq, "TWI" = zz_twi_sequence, "Consensus" = avg_event_zz)

h_w3 <- plot_test_vs_event(test_vs_event_data(models_w3, ds_w3, pick_event$W3, "W3"), "W3", pick_event$W3)
h_fb <- plot_test_vs_event(test_vs_event_data(models_fb, ds_fb, pick_event$FB, "FB"), "FB", pick_event$FB)
h_zz <- plot_test_vs_event(test_vs_event_data(models_zz, ds_zz, pick_event$ZZ, "ZZ"), "ZZ", pick_event$ZZ)

figure3 <- (h_w3 / h_fb / h_zz) + plot_layout(guides = "collect") +
  plot_annotation(
    caption = "Columns = test sequences vs. one event's observed wetting and (reversed) drying order, re-ranked on shared nodes. Fill = absolute offset (black = exact match, lighter = larger miss). Diagonal '/' = test ranks node later than observed; '\\' = earlier; no line = exact match.",
    theme = theme(plot.caption = element_text(size = 9)))
figure3
ggsave("figure3_test_vs_event_heatmap.pdf", figure3, width = 8, height = 9)
```


# Figure 4 - Grid of TAUS wetting vs wetting for each event

Grid of taus of observed sequences; FINAL FIGURE
```{r}
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 8
# Heatmap of pairwise Kendall's tau between observed event sequences, per
# watershed x direction. Diagonal left blank. Events restricted to the curated
# subset (both-direction + direction-specific, >= 5 cycling sensors).

reorder_events <- FALSE

# define set of events to use for each watershed
dir_spec <- list(
  W3 = list(remove = c(13, 18), wetting_only = c(8, 10),            drying_only = integer(0)),
  FB = list(remove = c(1),      wetting_only = c(3, 5, 9),          drying_only = c(7, 10, 12)),
  ZZ = list(remove = c(),     wetting_only = c(4, 11, 12, 13), drying_only = c(1))
)
universe <- list(W3 = 1:18, FB = 1:19, ZZ = 1:18)
# curated event sets (from earlier) ------------------------------------------
wetting_eligible <- function(spec, u) sort(setdiff(u, c(spec$remove, spec$drying_only)))
drying_eligible  <- function(spec, u) sort(setdiff(u, c(spec$remove, spec$wetting_only)))
keep_trans <- list(
  W3 = events_with_transitions(input_w3, w3_events_all, W3_IDs, 3),
  FB = events_with_transitions(input_fb, FBall_events,  FB_IDs, 3),
  ZZ = events_with_transitions(input_zz, ZZall_events,  ZZ_IDs, 3))
keep_wet <- Map(intersect, Map(wetting_eligible, dir_spec, universe), keep_trans)
keep_dry <- Map(intersect, Map(drying_eligible,  dir_spec, universe), keep_trans)

# calculate drying and wetting pairs
# ---- wetting pairs

# ---- all-pairs wetting-vs-wetting within a watershed ------------------------
pairwise_wetting <- function(wet_long, shed, keep_events, min_shared = 3){
  rl <- event_rank_list(wet_long)
  evs <- intersect(names(rl), as.character(keep_events))
  if (length(evs) < 2) return(tibble())
  out <- vector("list", 0)
  for (cc in combn(evs, 2, simplify = FALSE)){
    m <- inner_join(rl[[cc[1]]], rl[[cc[2]]], by = "sensor_id", suffix = c("_i", "_j"))
    if (nrow(m) < min_shared) next
    kt <- suppressWarnings(cor.test(m$rank_i, m$rank_j, method = "kendall"))
    out[[length(out) + 1]] <- tibble(
      shed = shed, event_i = as.integer(cc[1]), event_j = as.integer(cc[2]),
      n_shared = nrow(m), kendall_tau = unname(kt$estimate), tau_p = kt$p.value)
  }
  bind_rows(out)
}

wetting_pairs <- bind_rows(
  pairwise_wetting(ds_w3$wetting, "W3", keep_wet$W3),
  pairwise_wetting(ds_fb$wetting, "FB", keep_wet$FB),
  pairwise_wetting(ds_zz$wetting, "ZZ", keep_wet$ZZ)
) %>%
  mutate(shed = factor(shed, levels = c("W3", "FB", "ZZ"))) %>%
  arrange(shed, event_i, event_j)

# ---- all-pairs drying-vs-drying within a watershed --------------------------
pairwise_drying <- function(dry_long, shed, keep_events, min_shared = 3){
  rl <- event_rank_list(dry_long)
  evs <- intersect(names(rl), as.character(keep_events))
  if (length(evs) < 2) return(tibble())
  out <- vector("list", 0)
  for (cc in combn(evs, 2, simplify = FALSE)){
    m <- inner_join(rl[[cc[1]]], rl[[cc[2]]], by = "sensor_id", suffix = c("_i", "_j"))
    if (nrow(m) < min_shared) next
    kt <- suppressWarnings(cor.test(m$rank_i, m$rank_j, method = "kendall"))
    out[[length(out) + 1]] <- tibble(
      shed = shed, event_i = as.integer(cc[1]), event_j = as.integer(cc[2]),
      n_shared = nrow(m), kendall_tau = unname(kt$estimate), tau_p = kt$p.value)
  }
  bind_rows(out)
}

drying_pairs <- bind_rows(
  pairwise_drying(ds_w3$drying, "W3", keep_dry$W3),
  pairwise_drying(ds_fb$drying, "FB", keep_dry$FB),
  pairwise_drying(ds_zz$drying, "ZZ", keep_dry$ZZ)
) %>%
  mutate(shed = factor(shed, levels = c("W3", "FB", "ZZ"))) %>%
  arrange(shed, event_i, event_j)

# long, symmetric tau (NO diagonal), filtered to the eligible events ----------
tau_matrix_long <- function(pairs, direction, keep_list){
  purrr::imap_dfr(keep_list, function(evs, shed_sel){
    p <- pairs %>% filter(shed == shed_sel, event_i %in% evs, event_j %in% evs)
    if (nrow(p) == 0) return(tibble())
    bind_rows(
      p %>% transmute(shed = shed_sel, a = event_i, b = event_j, kendall_tau),
      p %>% transmute(shed = shed_sel, a = event_j, b = event_i, kendall_tau)
    )
  }) %>% mutate(direction = direction)
}

tau_long <- bind_rows(
  tau_matrix_long(wetting_pairs, "Wetting", keep_wet),
  tau_matrix_long(drying_pairs,  "Drying",  keep_dry)
) %>%
  mutate(shed = factor(shed, levels = c("W3", "FB", "ZZ")),
         direction = factor(direction, levels = c("Wetting", "Drying")))

# per-panel clustering order (on 1 - tau, diagonal = 1 only for the distance) --
order_one <- function(df){
  evs <- sort(unique(c(df$a, df$b)))
  if (!reorder_events || length(evs) < 3) return(as.character(evs))
  m <- matrix(0, length(evs), length(evs), dimnames = list(evs, evs))
  for (i in seq_len(nrow(df))) m[as.character(df$a[i]), as.character(df$b[i])] <- df$kendall_tau[i]
  diag(m) <- 1
  as.character(evs[hclust(as.dist(1 - m))$order])
}

panels <- tau_long %>% group_split(shed, direction)
plot_one <- function(df){
  ord <- order_one(df)
  df <- df %>% mutate(a = factor(as.character(a), levels = ord),
                      b = factor(as.character(b), levels = ord))
  ggplot(df, aes(a, b, fill = kendall_tau)) +
    geom_tile(colour = "white") +
    scale_fill_gradient2(low = "#2166AC", mid = "grey95", high = "#B2182B",
                         midpoint = 0, limits = c(-1, 1), name = "Kendall\ntau") +
    coord_equal() +
    labs(x = "Event", y = "Event",
         title = paste0(df$shed[1], " — ", df$direction[1])) +
    theme_minimal(base_size = 10) +
    theme(panel.grid = element_blank(),
          axis.text = element_text(size = 7),
          plot.title = element_text(size = 11, face = "bold"))
}

fig_tau_hm <- patchwork::wrap_plots(lapply(panels, plot_one), ncol = 2, byrow = TRUE) +
  plot_layout(guides = "collect")
fig_tau_hm
ggsave("figure_event_tau_heatmap.pdf", fig_tau_hm, width = 12, height = 8)
ggsave("figure_event_tau_heatmap.pdf", fig_tau_hm, width = 12, height = 8)

```
# Figure 5 - Grid of TAUS comparing wetting to drying
squares the same size, but no different than above
```{r}
#| message: false
#| warning: false
#| fig-width: 13
#| fig-height: 5
# Event x event Kendall's tau: each event's WETTING sequence (rows) vs each
# event's DRYING sequence (cols), curated events per axis. Tile size held
# constant across panels by sizing panels to their event counts (no coord_equal).

reverse_drying <- FALSE

dir_spec <- list(
  W3 = list(remove = c(13, 18), wetting_only = c(8, 10),            drying_only = integer(0)),
  FB = list(remove = c(1),      wetting_only = c(3, 5, 9),          drying_only = c(7, 10, 12)),
  ZZ = list(remove = c(),       wetting_only = c(4, 11, 12, 13),    drying_only = c(1))
)
universe <- list(W3 = 1:18, FB = 1:19, ZZ = 1:18)
wetting_eligible <- function(spec, u) sort(setdiff(u, c(spec$remove, spec$drying_only)))
drying_eligible  <- function(spec, u) sort(setdiff(u, c(spec$remove, spec$wetting_only)))
keep_trans <- list(
  W3 = events_with_transitions(input_w3, w3_events_all, W3_IDs, 3),
  FB = events_with_transitions(input_fb, FBall_events,  FB_IDs, 3),
  ZZ = events_with_transitions(input_zz, ZZall_events,  ZZ_IDs, 3))
keep_wet <- Map(intersect, Map(wetting_eligible, dir_spec, universe), keep_trans)
keep_dry <- Map(intersect, Map(drying_eligible,  dir_spec, universe), keep_trans)

event_rank_list <- function(dir_long, reverse = FALSE){
  dir_long %>% group_by(event_id) %>% mutate(rank = row_number()) %>%
    { if (reverse) mutate(., rank = max(rank) - rank + 1) else . } %>%
    ungroup() %>% { split(select(., sensor_id, rank), .$event_id) }
}

wet_dry_grid <- function(wet_long, dry_long, shed, keep_w, keep_d, min_shared = 3){
  wl <- event_rank_list(wet_long)
  dl <- event_rank_list(dry_long, reverse = reverse_drying)
  wl <- wl[names(wl) %in% as.character(keep_w)]
  dl <- dl[names(dl) %in% as.character(keep_d)]
  out <- vector("list", 0)
  for (wi in names(wl)) for (dj in names(dl)){
    m <- inner_join(wl[[wi]], dl[[dj]], by = "sensor_id", suffix = c("_w", "_d"))
    if (nrow(m) < min_shared) next
    kt <- suppressWarnings(cor(m$rank_w, m$rank_d, method = "kendall"))
    out[[length(out)+1]] <- tibble(shed = shed,
                                   wet_event = as.integer(wi),
                                   dry_event = as.integer(dj),
                                   kendall_tau = kt)
  }
  bind_rows(out)
}

grid_all <- bind_rows(
  wet_dry_grid(ds_w3$wetting, ds_w3$drying, "W3", keep_wet$W3, keep_dry$W3),
  wet_dry_grid(ds_fb$wetting, ds_fb$drying, "FB", keep_wet$FB, keep_dry$FB),
  wet_dry_grid(ds_zz$wetting, ds_zz$drying, "ZZ", keep_wet$ZZ, keep_dry$ZZ)
) %>% mutate(shed = factor(shed, levels = c("W3", "FB", "ZZ")))

plot_one <- function(df, keep_w, keep_d){
  df <- df %>% mutate(wet_event = factor(wet_event, levels = sort(keep_w)),
                      dry_event = factor(dry_event, levels = sort(keep_d)))
  diag_df <- df %>% filter(as.character(wet_event) == as.character(dry_event))
  ggplot(df, aes(dry_event, wet_event, fill = kendall_tau)) +
    geom_tile(colour = "white") +
    geom_tile(data = diag_df, fill = NA, colour = "black", linewidth = 0.4) +
    scale_fill_gradient2(low = "#2166AC", mid = "grey95", high = "#B2182B",
                         midpoint = 0, limits = c(-1, 1), name = "Kendall\ntau") +
    scale_x_discrete(drop = FALSE) + scale_y_discrete(drop = FALSE) +
    labs(x = "Event (drying)", y = "Event (wetting)", title = df$shed[1]) +
    theme_minimal(base_size = 10) +
    theme(panel.grid = element_blank(),
          aspect.ratio = length(keep_w) / length(keep_d),  # square tiles per panel
          axis.text = element_text(size = 7),
          plot.title = element_text(size = 11, face = "bold"))
}

plots <- list(
  plot_one(filter(grid_all, shed == "W3"), keep_wet$W3, keep_dry$W3),
  plot_one(filter(grid_all, shed == "FB"), keep_wet$FB, keep_dry$FB),
  plot_one(filter(grid_all, shed == "ZZ"), keep_wet$ZZ, keep_dry$ZZ))

# widths proportional to number of drying-axis events -> constant tile size ----
panel_widths <- c(length(keep_dry$W3), length(keep_dry$FB), length(keep_dry$ZZ))

figure1 <- patchwork::wrap_plots(plots, ncol = 3, widths = panel_widths) +
  plot_layout(guides = "collect") +
  plot_annotation(
    caption = paste(
      "Kendall's tau: each event's wetting order (rows) vs each event's drying order (columns), curated per axis.",
      "Black-outlined diagonal = same event. All grid squares equal size.",
      if (reverse_drying) "Drying = 1 dries last; red = same order."
      else "Drying = 1 dries first; blue (negative) = reversed."))
figure1
ggsave("figure1_wet_vs_dry_tau_heatmap_equal.pdf", figure1, width = 13, height = 5)
```
# Figure 6 - boxplots of Taus for test sequences

Boxplots, but making sure I am using the correct subset of events
```{r}
#| message: false
#| warning: false
#| fig-width: 11
#| fig-height: 6
# For each event, compare its WETTING and DRYING sequence to the two MODEL
# sequences (Flow Persistence, TWI) via Kendall's tau on shared nodes.
# Drying ranks are REVERSED (1 = dries last) so that, like wetting and like the
# model sequences, rank 1 = most persistent -> positive tau = model agrees.

dir_spec <- list(
  W3 = list(remove = c(13, 18), wetting_only = c(8, 10),            drying_only = integer(0)),
  FB = list(remove = c(1),      wetting_only = c(3, 5, 9),          drying_only = c(7, 10, 12)),
  ZZ = list(remove = c(),       wetting_only = c(4, 11, 12, 13),    drying_only = c(1))
)
universe <- list(W3 = 1:18, FB = 1:19, ZZ = 1:18)
wetting_eligible <- function(spec, u) sort(setdiff(u, c(spec$remove, spec$drying_only)))
drying_eligible  <- function(spec, u) sort(setdiff(u, c(spec$remove, spec$wetting_only)))
keep_trans <- list(
  W3 = events_with_transitions(input_w3, w3_events_all, W3_IDs, 3),
  FB = events_with_transitions(input_fb, FBall_events,  FB_IDs, 3),
  ZZ = events_with_transitions(input_zz, ZZall_events,  ZZ_IDs, 3))
keep_wet <- Map(intersect, Map(wetting_eligible, dir_spec, universe), keep_trans)
keep_dry <- Map(intersect, Map(drying_eligible,  dir_spec, universe), keep_trans)

tau_event_model <- function(ev_ranks, model_seq, min_shared = 3){
  m <- ev_ranks %>%
    inner_join(model_seq %>% transmute(sensor_id = ID, mrank = sequence),
               by = "sensor_id")
  if (nrow(m) < min_shared) return(tibble(tau = NA_real_, p = NA_real_, n = nrow(m)))
  ct <- suppressWarnings(cor.test(m$rank, m$mrank, method = "kendall"))
  tibble(tau = unname(ct$estimate), p = ct$p.value, n = nrow(m))
}

compare_dir <- function(dir_long, keep_events, pk_seq, twi_seq, shed, direction){
  rl <- event_rank_list(dir_long)
  evs <- intersect(names(rl), as.character(keep_events))
  purrr::map_dfr(evs, function(e){
    er <- rl[[e]]
    # reverse drying ranks: 1 = dries last (most persistent), matching model order
    if (direction == "Drying") er <- er %>% mutate(rank = max(rank) - rank + 1)
    bind_rows(
      tau_event_model(er, pk_seq)  %>% mutate(model = "Flow Persistence"),
      tau_event_model(er, twi_seq) %>% mutate(model = "TWI")
    ) %>% mutate(shed = shed, direction = direction, event_id = as.integer(e))
  })
}

event_model_tau <- bind_rows(
  compare_dir(ds_w3$wetting, keep_wet$W3, W3_pk_seq, w3_twi_sequence, "W3", "Wetting"),
  compare_dir(ds_w3$drying,  keep_dry$W3, W3_pk_seq, w3_twi_sequence, "W3", "Drying"),
  compare_dir(ds_fb$wetting, keep_wet$FB, FB_pk_seq, fb_twi_sequence, "FB", "Wetting"),
  compare_dir(ds_fb$drying,  keep_dry$FB, FB_pk_seq, fb_twi_sequence, "FB", "Drying"),
  compare_dir(ds_zz$wetting, keep_wet$ZZ, ZZ_pk_seq, zz_twi_sequence, "ZZ", "Wetting"),
  compare_dir(ds_zz$drying,  keep_dry$ZZ, ZZ_pk_seq, zz_twi_sequence, "ZZ", "Drying")
) %>%
  filter(!is.na(tau)) %>%
  mutate(shed      = factor(shed, levels = c("W3", "FB", "ZZ")),
         direction = factor(direction, levels = c("Wetting", "Drying")),
         model     = factor(model, levels = c("Flow Persistence", "TWI")))

model_cols <- c("Flow Persistence" = "#7ebdc2", "TWI" = "#f3dfa2")

fig_event_model <- ggplot(event_model_tau,
                          aes(x = direction, y = tau, fill = model, color = model)) +
  geom_hline(yintercept = 0, linetype = 2, color = "grey60") +
  geom_boxplot(width = 0.55, alpha = 0.25, outlier.shape = NA,
               position = position_dodge(0.65)) +
  geom_point(size = 2, alpha = 0.8,
             position = position_jitterdodge(jitter.width = 0.12, dodge.width = 0.65)) +
  facet_wrap(~ shed) +
  scale_fill_manual(values = model_cols, name = "Model sequence") +
  scale_color_manual(values = model_cols, name = "Model sequence") +
  coord_cartesian(ylim = c(-0.5, 1)) +
  labs(x = "", y = "Kendall's tau (event order vs. model)") +
  theme_bw(base_size = 12) +
  theme(strip.background = element_blank(), legend.position = "right",
        panel.grid.major = element_blank(), # Removes major grid lines
    panel.grid.minor = element_blank())  # Removes minor grid lines)

fig_event_model
ggsave("figure_event_vs_model_tau.pdf", fig_event_model, width = 8.5, height = 3)
ggsave("figure_event_vs_model_tau.png", fig_event_model, width = 8.5, height = 3)

event_model_summary <- event_model_tau %>%
  group_by(shed, direction, model) %>%
  summarise(n_events = n(),
            mean_tau = mean(tau), median_tau = median(tau), sd_tau = sd(tau),
            frac_sig = mean(p < 0.05, na.rm = TRUE), .groups = "drop")
knitr::kable(
  event_model_summary %>%
    mutate(across(c(mean_tau, median_tau, sd_tau, frac_sig), ~ round(.x, 3))),
  caption = "Per-event agreement (Kendall's tau) of each event's wetting and (reversed) drying order with the Flow Persistence and TWI model sequences.")
```

# Figure 7 — Flow Persistence vs. TWI accuracy through time (events only)

```{r fig7-accuracy-time}
#| message: false
#| warning: false
#| fig-width: 11
#| fig-height: 12
# RQ3. For each watershed: backdrop cropped to the two STIC seasons (no winter
# gap), grey bands = identified events, point at each EVENT timestep coloured by
# whether Flow Persistence or TWI is more accurate (or a tie). All rows share the
# SAME season x-limits (anchored to W3) so 2024 ends in July for every panel.
fp_twi_cols <- c("Flow Persistence" = "#7ebdc2",
                 "TWI"              = "#f3dfa2",
                 "Tie"              = "gray10")

# shared season windows, anchored to the W3 record ----------------------------
w3_2024_end <- max(filter(w3_q_seasonal, datetime >= start24)$datetime)
season_lims <- tibble(
  season = c("Summer 2023", "Summer 2024"),
  xmin   = c(start_23, start24),
  xmax   = c(stop_23,  w3_2024_end))

clip_seasons <- function(df) df %>% add_season() %>%
  left_join(season_lims, by = "season") %>%
  filter(datetime >= xmin, datetime <= xmax) %>%
  select(-xmin, -xmax)

# event bands, tagged by season and clipped to the shared windows -------------
season_bands <- function(event_df) event_df %>%
  filter(!is.na(event_id)) %>%
  group_by(event_id) %>%
  summarise(xmin = min(datetime), xmax = max(datetime), .groups = "drop") %>%
  mutate(season = if_else(xmax <= stop_23, "Summer 2023", "Summer 2024")) %>%
  left_join(season_lims, by = "season", suffix = c("", "_lim")) %>%
  filter(xmax >= xmin_lim, xmin <= xmax_lim) %>%
  mutate(xmin = pmax(xmin, xmin_lim), xmax = pmin(xmax, xmax_lim)) %>%
  select(event_id, xmin, xmax, season)

# winner points (FP vs TWI, or tie) at event timesteps ------------------------
winner_pts <- function(df_long, event_df, backdrop, ycol){
  df_long %>%
    filter(model %in% c("pk_accuracy", "twi_accuracy")) %>%
    pivot_wider(names_from = model, values_from = accuracy) %>%
    filter(datetime %in% event_dt(event_df)) %>%
    mutate(winner = case_when(pk_accuracy  > twi_accuracy ~ "Flow Persistence",
                              twi_accuracy > pk_accuracy  ~ "TWI",
                              TRUE                          ~ "Tie"),
           winner = factor(winner, levels = c("Flow Persistence", "TWI", "Tie"))) %>%
    inner_join(backdrop %>% select(datetime, all_of(ycol)), by = "datetime") %>%
    clip_seasons()
}

# one watershed row -----------------------------------------------------------
fig7_panel <- function(backdrop, bands_df, pts, ycol, ylab, show_strip = TRUE){
  ggplot() +
    geom_rect(data = bands_df, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
              fill = "grey85", alpha = 0.5) +
    geom_line(data = backdrop, aes(datetime, .data[[ycol]]),
              color = "grey55", linewidth = 0.4) +
    geom_point(data = pts, aes(datetime, .data[[ycol]], color = winner), size = 0.5) +
    geom_blank(data = season_lims, aes(x = xmin)) +
    geom_blank(data = season_lims, aes(x = xmax)) +
    facet_wrap(~ season, scales = "free_x", nrow = 1) +
    scale_color_manual(values = fp_twi_cols, name = "More accurate", drop = FALSE) +
    scale_x_datetime(date_labels = "%b", date_breaks = "1 month", timezone = "UTC") +
    labs(x = "", y = ylab) +
    theme_classic() +
    theme(strip.background = element_blank(),
          strip.text = if (show_strip) element_text() else element_blank(),
          legend.position = "right",
          text = element_text(size = 12))
}

# backdrops (clipped to shared season windows) --------------------------------
w3_backdrop <- clip_seasons(w3_q_seasonal)
fb_backdrop <- clip_seasons(FBall_events %>% select(datetime, roll_mean))
zz_backdrop <- clip_seasons(ZZall_events %>% select(datetime, roll_mean))

p_w3 <- fig7_panel(w3_backdrop, season_bands(w3_events_all),
                   winner_pts(df_long_w3, w3_events_all, w3_q_seasonal, "Q_mm_day"),
                   "Q_mm_day", "W3 discharge (mm/day)", show_strip = TRUE)
p_fb <- fig7_panel(fb_backdrop, season_bands(FBall_events),
                   winner_pts(df_long_fb, FBall_events, FBall_events, "roll_mean"),
                   "roll_mean", "FB stage (cm)", show_strip = FALSE)
p_zz <- fig7_panel(zz_backdrop, season_bands(ZZall_events),
                   winner_pts(df_long_zz, ZZall_events, ZZall_events, "roll_mean"),
                   "roll_mean", "ZZ stage (cm)", show_strip = FALSE)

figure7 <- p_w3 / p_fb / p_zz + plot_layout(guides = "collect")
figure7
ggsave("figure7_accuracy_time.pdf", figure7, width = 6.5, units = "in")
```

```{r just discharge}
#| message: false
#| warning: false
#| fig-width: 11
#| fig-height: 12
# RQ3. For each watershed: backdrop cropped to the two STIC seasons (no winter
# gap), grey bands = identified events, point at each EVENT timestep coloured by
# whether Flow Persistence or TWI is more accurate (or a tie). All rows share the
# SAME season x-limits (anchored to W3) so 2024 ends in July for every panel.
fp_twi_cols <- c("Flow Persistence" = "#7ebdc2",
                 "TWI"              = "#f3dfa2",
                 "Tie"              = "gray10")

# shared season windows, anchored to the W3 record ----------------------------
w3_2024_end <- max(filter(w3_q_seasonal, datetime >= start24)$datetime)
season_lims <- tibble(
  season = c("Summer 2023", "Summer 2024"),
  xmin   = c(start_23, start24),
  xmax   = c(stop_23,  w3_2024_end))

clip_seasons <- function(df) df %>% add_season() %>%
  left_join(season_lims, by = "season") %>%
  filter(datetime >= xmin, datetime <= xmax) %>%
  select(-xmin, -xmax)

# event bands, tagged by season and clipped to the shared windows -------------
season_bands <- function(event_df) event_df %>%
  filter(!is.na(event_id)) %>%
  group_by(event_id) %>%
  summarise(xmin = min(datetime), xmax = max(datetime), .groups = "drop") %>%
  mutate(season = if_else(xmax <= stop_23, "Summer 2023", "Summer 2024")) %>%
  left_join(season_lims, by = "season", suffix = c("", "_lim")) %>%
  filter(xmax >= xmin_lim, xmin <= xmax_lim) %>%
  mutate(xmin = pmax(xmin, xmin_lim), xmax = pmin(xmax, xmax_lim)) %>%
  select(event_id, xmin, xmax, season)



# one watershed row -----------------------------------------------------------
fig7_panel <- function(backdrop, bands_df, pts, ycol, ylab, show_strip = TRUE){
  ggplot() +
    geom_rect(data = bands_df, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
              fill = "lightblue", alpha = 0.5) +
    geom_line(data = backdrop, aes(datetime, .data[[ycol]]),
              color = "black", linewidth = 0.4) +
    #geom_point(data = pts, aes(datetime, .data[[ycol]], color = winner), size = 0.5) +
    geom_blank(data = season_lims, aes(x = xmin)) +
    geom_blank(data = season_lims, aes(x = xmax)) +
    facet_wrap(~ season, scales = "free_x", nrow = 1) +
    scale_color_manual(values = fp_twi_cols, name = "More accurate", drop = FALSE) +
    scale_x_datetime(date_labels = "%b", date_breaks = "1 month", timezone = "UTC") +
    labs(x = "", y = ylab) +
    theme_classic() +
    theme(strip.background = element_blank(),
          strip.text = if (show_strip) element_text() else element_blank(),
          legend.position = "right",
          text = element_text(size = 12))
}

# backdrops (clipped to shared season windows) --------------------------------
w3_backdrop <- clip_seasons(w3_q_seasonal)


p_w3 <- fig7_panel(w3_backdrop, season_bands(w3_events_all),
                   winner_pts(df_long_w3, w3_events_all, w3_q_seasonal, "Q_mm_day"),
                   "Q_mm_day", "W3 discharge (mm/day)", show_strip = TRUE)

p_w3
ggsave("just_discharge.png", p_w3, width = 7.5, height = 4, units = "in")
```



# Figure 8 — Maps of node-level model accuracy (Flow Persistence & TWI)

```{r fig8-accuracy-maps}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 12
# Per-node accuracy over the whole record, Flow Persistence (pk) and TWI only.
# Rotated for display, greyscale points, one row per watershed faceted by model.
comparison_w3 <- w3_mod$comparison
comparison_fb <- fb_mod$comparison
comparison_zz <- zz_mod$comparison

id_accuracy <- function(comparison){
  comparison %>%
    group_by(ID) %>%
    summarize(total        = n(),
              pk_accuracy  = sum(pk_eval  == "correct") / total,
              twi_accuracy = sum(twi_eval == "correct") / total,
              .groups = "drop") %>%
    select(ID, pk_accuracy, twi_accuracy) %>%
    pivot_longer(c(pk_accuracy, twi_accuracy),
                 names_to = "model", values_to = "accuracy") %>%
    mutate(model = case_when(
             model == "pk_accuracy"  ~ "Flow Persistence",
             model == "twi_accuracy" ~ "Topographic Wetness Index"),
           model = fct_relevel(model, c("Flow Persistence",
                                        "Topographic Wetness Index")))
}
ID_acc_w3 <- id_accuracy(comparison_w3)
ID_acc_fb <- id_accuracy(comparison_fb)
ID_acc_zz <- id_accuracy(comparison_zz)

acc_grey <- c("white", "lightgrey", "darkgrey", "black")  # 0 -> 1

# W3 — not rotated; keeps the model strip labels (top row) --------------------
W3_rot <- ggplot() + theme_void() +
  geom_sf(data = w3_outline, fill = NA, color = "#FFD166", alpha = 0.3, lwd = 1) +
  geom_sf(data = w3_net, colour = "grey", lwd = 1) +
  geom_sf(data = w3_stic_locs_r %>% inner_join(ID_acc_w3, by = "ID"),
          aes(fill = accuracy), size = 3, pch = 21) +
  scale_fill_gradientn(colors = acc_grey, limits = c(0, 1)) +
  theme(rect = element_rect(fill = "transparent", color = NA),
        legend.position = "none") +
  facet_grid(~ model)

# FB — rotated 90 degrees ------------------------------------------------------
FB_rot <- ggplot() + theme_void() +
  geom_sf(data = terra::spin(fb_outline, 90, 279933.5, 4869173),
          fill = NA, color = "#397367", alpha = 0.3, lwd = 1) +
  geom_sf(data = terra::spin(fb_net, 90, 279933.5, 4869173),
          colour = "grey", lwd = 1) +
  geom_sf(data = terra::spin(lcc_fb, 90, 279933.5, 4869173) %>%
            inner_join(ID_acc_fb, by = "ID"),
          aes(fill = accuracy), size = 3, pch = 21) +
  scale_fill_gradientn(colors = acc_grey, limits = c(0, 1)) +
  theme(rect = element_rect(fill = "transparent", color = NA),
        legend.position = "none", strip.text.x = element_blank()) +
  facet_grid(~ model)

# ZZ — rotated 180 degrees; carries the shared legend (bottom) -----------------
ZZ_rot <- ggplot() + theme_void() +
  geom_sf(data = terra::spin(zz_outline, 180, 277401.8, 4867071),
          fill = NA, color = "#7E6B8F", alpha = 0.3, lwd = 1) +
  geom_sf(data = terra::spin(zz_net, 180, 277401.8, 4867071),
          colour = "grey", lwd = 1) +
  geom_sf(data = terra::spin(lcc, 180, 277401.8, 4867071) %>%
            inner_join(ID_acc_zz, by = "ID"),
          aes(fill = accuracy), size = 3, pch = 21) +
  scale_fill_gradientn(colors = acc_grey, limits = c(0, 1), name = "Accuracy") +
  theme(rect = element_rect(fill = "transparent", color = NA),
        legend.position = "bottom",
        panel.spacing = unit(5, "lines"),
        strip.text.x = element_blank()) +
  facet_grid(~ model)

figure8 <- W3_rot / FB_rot / ZZ_rot
figure8
ggsave("figure8_accuracy_maps.pdf", figure8, width = 8, height = 12)
# Note: add north arrows and a scale bar in Canva/Illustrator afterwards.
```

```{r fig8-accuracy-maps}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 12
# Per-node accuracy over the whole record, Flow Persistence (pk) and TWI only.
# Rotated for display, greyscale points, one row per watershed faceted by model.
comparison_w3 <- w3_mod$comparison
comparison_fb <- fb_mod$comparison
comparison_zz <- zz_mod$comparison

id_accuracy <- function(comparison){
  comparison %>%
    group_by(ID) %>%
    summarize(total        = n(),
              pk_accuracy  = sum(pk_eval  == "correct") / total,
              twi_accuracy = sum(twi_eval == "correct") / total,
              .groups = "drop") %>%
    select(ID, pk_accuracy, twi_accuracy) %>%
    pivot_longer(c(pk_accuracy, twi_accuracy),
                 names_to = "model", values_to = "accuracy") %>%
    mutate(model = case_when(
             model == "pk_accuracy"  ~ "Flow Persistence",
             model == "twi_accuracy" ~ "Topographic Wetness Index (TWI)"),
           model = fct_relevel(model, c("Flow Persistence",
                                        "Topographic Wetness Index (TWI)")))
}
ID_acc_w3 <- id_accuracy(comparison_w3)
ID_acc_fb <- id_accuracy(comparison_fb)
ID_acc_zz <- id_accuracy(comparison_zz)

acc_grey <- c("white", "lightgrey", "darkgrey", "black")  # 0 -> 1

# W3 — not rotated; keeps the model strip labels (top row) --------------------
W3_rot <- ggplot() + theme_void() +
  geom_sf(data = w3_outline, fill = NA, color = "#FFD166", alpha = 0.3, lwd = 1) +
  geom_sf(data = w3_net, colour = "grey", lwd = 1) +
  geom_sf(data = w3_stic_locs_r %>% inner_join(ID_acc_w3, by = "ID"),
          aes(fill = accuracy), size = 3, pch = 21) +
  scale_fill_stepsn(colors = c("white", "#ffffff", "#c0c0c0", "#808080", "#2b2b2b", "#000001"),
                  breaks = c(0.5, 0.6, 0.7, 0.8, 0.9), limits = c(0.5, 1),
                  show.limits = TRUE, name = "Accuracy")+
  theme(rect = element_rect(fill = "transparent", color = NA),
        legend.position = "right",
        element_text(size = 12)) +
  facet_grid(~ model)

# FB — rotated 90 degrees ------------------------------------------------------
FB_rot <- ggplot() + theme_void() +
  geom_sf(data = terra::spin(fb_outline, 90, 279933.5, 4869173),
          fill = NA, color = "#397367", alpha = 0.3, lwd = 1) +
  geom_sf(data = terra::spin(fb_net, 90, 279933.5, 4869173),
          colour = "grey", lwd = 1) +
  geom_sf(data = terra::spin(lcc_fb, 90, 279933.5, 4869173) %>%
            inner_join(ID_acc_fb, by = "ID"),
          aes(fill = accuracy), size = 3, pch = 21) +
  scale_fill_stepsn(colors = c("white", "#ffffff", "#c0c0c0", "#808080", "#2b2b2b", "#000001"),
                  breaks = c(0.5, 0.6, 0.7, 0.8, 0.9), limits = c(0.5, 1),
                  show.limits = TRUE, name = "Accuracy")+
  theme(rect = element_rect(fill = "transparent", color = NA),
        legend.position = "right", strip.text.x = element_blank(),
        element_text(size = 12)) +
  facet_grid(~ model)

# ZZ — rotated 180 degrees; carries the shared legend (bottom) -----------------
ZZ_rot <- ggplot() + theme_void() +
  geom_sf(data = terra::spin(zz_outline, 180, 277401.8, 4867071),
          fill = NA, color = "#7E6B8F", alpha = 0.3, lwd = 1) +
  geom_sf(data = terra::spin(zz_net, 180, 277401.8, 4867071),
          colour = "grey", lwd = 1) +
  geom_sf(data = terra::spin(lcc, 180, 277401.8, 4867071) %>%
            inner_join(ID_acc_zz, by = "ID"),
          aes(fill = accuracy), size = 3, pch = 21) +
  scale_fill_stepsn(colors = c("white", "#ffffff", "#c0c0c0", "#808080", "#2b2b2b", "#000001"),
                  breaks = c(0.5, 0.6, 0.7, 0.8, 0.9), limits = c(0.5, 1),
                  show.limits = TRUE, name = "Accuracy")+
  theme(rect = element_rect(fill = "transparent", color = NA),
        legend.position = "right",
        panel.spacing = unit(5, "lines"),
        strip.text.x = element_blank(),
        element_text(size = 12)) +
  facet_grid(~ model)

figure8 <- (W3_rot / FB_rot / ZZ_rot) + plot_layout(guides = "collect")
figure8
ggsave("figure8_accuracy_maps.pdf", figure8, width = 6.5, units = "in")
# Note: add north arrows and a scale bar in Canva/Illustrator afterwards.
```



# Table 2 - average tau values across wetting vs wetting and drying vs drying
```{r}
#| message: false
#| warning: false
# Combines three analyses per watershed:
#   (1) Wetting consistency = pairwise tau among wetting sequences across events
#   (2) Drying consistency  = pairwise tau among drying sequences across events
#   (3) Wetting vs drying   = within-event tau between an event's wetting & drying,
#       tested with a one-sample Wilcoxon signed-rank test (H0: median tau = 0).

summ_consistency <- function(df, label){
  df %>% group_by(shed) %>%
    summarise(n        = n(),
              mean_tau = mean(kendall_tau, na.rm = TRUE),
              sd_tau   = sd(kendall_tau, na.rm = TRUE),
              frac_sig = mean(tau_p < 0.05, na.rm = TRUE),
              wilcox_p = NA_real_,
              .groups = "drop") %>%
    mutate(comparison = label)
}

# within-event wetting-vs-drying: Wilcoxon signed-rank on the per-event taus ----
wet_dry_wilcox <- wet_dry_events %>%
  group_by(shed) %>%
  summarise(n        = n(),
            mean_tau = mean(kendall_tau, na.rm = TRUE),
            sd_tau   = sd(kendall_tau, na.rm = TRUE),
            frac_sig = NA_real_,
            wilcox_p = tryCatch(
              wilcox.test(kendall_tau, mu = 0)$p.value,
              error = function(e) NA_real_),
            .groups = "drop") %>%
  mutate(comparison = "Wetting vs drying (within event)")

consistency <- bind_rows(
  summ_consistency(wetting_pairs, "Wetting consistency (across events)"),
  summ_consistency(drying_pairs,  "Drying consistency (across events)"),
  wet_dry_wilcox
) %>%
  mutate(shed = factor(shed, levels = c("W3", "FB", "ZZ")),
         comparison = factor(comparison, levels = c(
           "Wetting consistency (across events)",
           "Drying consistency (across events)",
           "Wetting vs drying (within event)"))) %>%
  mutate(interpretation = case_when(
    comparison != "Wetting vs drying (within event)" & mean_tau >= 0.5 ~ "Highly consistent",
    comparison != "Wetting vs drying (within event)" & mean_tau >= 0.3 ~ "Moderately consistent",
    comparison != "Wetting vs drying (within event)"                   ~ "Weakly consistent",
    mean_tau <= -0.3 & wilcox_p < 0.05 ~ "Significantly reversed",
    mean_tau <    0  & wilcox_p < 0.05 ~ "Reversed (weak but sig.)",
    wilcox_p >= 0.05 | is.na(wilcox_p) ~ "No sig. difference",
    TRUE ~ "Same order both directions")) %>%
  arrange(shed, comparison) %>%
  select(shed, comparison, n, mean_tau, sd_tau, frac_sig, wilcox_p, interpretation)

knitr::kable(
  consistency %>%
    mutate(across(c(mean_tau, sd_tau, frac_sig), ~ round(.x, 3)),
           wilcox_p = ifelse(is.na(wilcox_p), "—",
                             format.pval(wilcox_p, digits = 2, eps = 1e-3))),
  col.names = c("Watershed", "Comparison", "n", "Mean tau", "SD tau",
                "Frac. p<0.05", "Wilcoxon p", "Interpretation"),
  align = c("l", "l", "r", "r", "r", "r", "r", "l"),
  caption = "Combined sequence analysis. Rows 1-2: internal consistency of wetting and drying sequences across events (Frac. p<0.05 = share of event pairs with significant positive tau). Row 3: within-event wetting-vs-drying order, tested with a one-sample Wilcoxon signed-rank test (H0: median tau = 0); negative mean tau = drying reverses wetting. n = event pairs (rows 1-2) or events (row 3).")
```

# Table 3 - Accuracy of test sequences
```{r}
#| message: false
#| warning: false
# Accuracy of the Flow Persistence (pk) and TWI sequences at predicting the
# observed wet/dry pattern, split into WETTING vs DRYING and SUMMARIZED ACROSS
# EVENTS. Wetting = timesteps up to and including each event's peak network
# extent (most flowing nodes); drying = after the peak. Counts are pooled over
# events, so metrics are observation-weighted. Positive class = flowing (1):
# Sensitivity = flowing correctly predicted, Specificity = dry correctly predicted.

event_dir_metrics <- function(comparison, event_df, shed){
  dat <- comparison %>%
    select(datetime, ID, binary, pk, twi) %>%
    inner_join(event_df %>% select(datetime, event_id) %>% distinct(),
               by = "datetime") %>%
    filter(!is.na(event_id))

  peaks <- dat %>%
    group_by(event_id, datetime) %>%
    summarise(nwet = sum(binary == 1), .groups = "drop") %>%
    group_by(event_id) %>%
    slice_max(nwet, n = 1, with_ties = FALSE) %>%
    transmute(event_id, peak_t = datetime)

  dat %>%
    left_join(peaks, by = "event_id") %>%
    mutate(direction = if_else(datetime <= peak_t, "Wetting", "Drying")) %>%
    pivot_longer(c(pk, twi), names_to = "method", values_to = "pred") %>%
    mutate(method = recode(method,
                           pk  = "Flow Persistence",
                           twi = "Topographic Wetness Index")) %>%
    group_by(shed = shed, event_id, method, direction) %>%
    summarise(tp = sum(pred == 1 & binary == 1),
              tn = sum(pred == 0 & binary == 0),
              fp = sum(pred == 1 & binary == 0),
              fn = sum(pred == 0 & binary == 1),
              .groups = "drop")
}

# pool the per-event counts across events -------------------------------------
direction_summary <- bind_rows(
  event_dir_metrics(w3_mod$comparison, w3_events_all, "W3"),
  event_dir_metrics(fb_mod$comparison, FBall_events,  "FB"),
  event_dir_metrics(zz_mod$comparison, ZZall_events,  "ZZ")
) %>%
  group_by(shed, method, direction) %>%
  summarise(n_events = n_distinct(event_id),
            tp = sum(tp), tn = sum(tn), fp = sum(fp), fn = sum(fn),
            .groups = "drop") %>%
  mutate(n           = tp + tn + fp + fn,
         Accuracy    = (tp + tn) / n,
         Sensitivity = tp / (tp + fn),
         Specificity = tn / (tn + fp),
         shed      = factor(shed, levels = c("W3", "FB", "ZZ")),
         method    = factor(method, levels = c("Flow Persistence",
                                               "Topographic Wetness Index")),
         direction = factor(direction, levels = c("Wetting", "Drying"))) %>%
  arrange(shed, method, direction) %>%
  select(shed, method, direction, n_events, n, Accuracy, Sensitivity, Specificity)

# nicely formatted table (percentages, 1 decimal) -----------------------------
direction_summary_fmt <- direction_summary %>%
  mutate(across(c(Accuracy, Sensitivity, Specificity), ~ round(. * 100, 1)))

knitr::kable(
  direction_summary_fmt,
  col.names = c("Watershed", "Sequence", "Direction", "Events", "n",
                "Accuracy (%)", "Sensitivity (%)", "Specificity (%)"),
  align = c("l", "l", "l", "r", "r", "r", "r", "r"),
  caption = "Predictive accuracy of the Flow Persistence and TWI sequences during events, split into wetting (up to peak network extent) and drying (after peak) phases and summarized across all events. Counts are pooled over events (observation-weighted). Positive class = flowing: Sensitivity = fraction of flowing observations correctly predicted; Specificity = fraction of dry observations correctly predicted."
)


```
```{r}
#| message: false
#| warning: false
library(gt)

direction_summary_gt <- direction_summary_fmt %>%
  select(shed, method, direction, Accuracy, Sensitivity, Specificity)

tbl_dir <- direction_summary_gt %>%
  gt(groupname_col = "shed") %>%                       # row groups by watershed
  # ── Column labels ───────────────────────────────────────────────────────────
  cols_label(
    method      = "Sequence",
    direction   = "Direction",
    Accuracy    = "Accuracy",
    Sensitivity = "Sensitivity",
    Specificity = "Specificity"
  ) %>%
  # ── Title ───────────────────────────────────────────────────────────────────
  tab_header(
    title    = md("**Wetting vs. drying prediction accuracy during events**"),
    subtitle = md("Flow Persistence and TWI sequences, summarized across all events")
  ) %>%
  # ── Source note ─────────────────────────────────────────────────────────────
  tab_source_note(
    source_note = md(
      "Wetting = timesteps up to peak network extent; drying = after peak. Positive class = flowing: Sensitivity = fraction of flowing observations correctly predicted, Specificity = fraction of dry observations correctly predicted. Values are percentages, pooled (observation-weighted) across events."
    )
  ) %>%
  # ── Spanner over the metrics ─────────────────────────────────────────────────
  tab_spanner(
    label   = md("**Accuracy (%)**"),
    columns = c(Accuracy, Sensitivity, Specificity)
  ) %>%
  # ── Number formatting ────────────────────────────────────────────────────────
  fmt_number(columns = c(Accuracy, Sensitivity, Specificity), decimals = 1) %>%
  # ── Column formatting ───────────────────────────────────────────────────────
  cols_align(align = "center",
             columns = c(direction, Accuracy, Sensitivity, Specificity)) %>%
  cols_align(align = "left",
             columns = c(method)) %>%
  cols_width(
    method      ~ px(220),
    direction   ~ px(110),
    Accuracy    ~ px(110),
    Sensitivity ~ px(110),
    Specificity ~ px(110)
  ) %>%
  # ── Styling ─────────────────────────────────────────────────────────────────
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_row_groups()
  ) %>%
  tab_style(
    style = cell_text(style = "italic"),
    locations = cells_body(columns = method)
  ) %>%
  tab_style(
    style = list(cell_fill(color = "#f7f7f7"),
                 cell_text(weight = "bold")),
    locations = cells_column_labels()
  ) %>%
  tab_style(
    style = cell_borders(sides = "bottom", color = "#cccccc", weight = px(1)),
    locations = cells_body()
  ) %>%
  tab_style(
    style = cell_borders(sides = "top", color = "#333333", weight = px(2)),
    locations = cells_column_labels()
  ) %>%
  # Shade the drying rows so the two directions read as paired blocks
  tab_style(
    style = cell_fill(color = "#f0f4f8"),
    locations = cells_body(rows = direction == "Drying")
  ) %>%
  opt_table_font(font = google_font("Source Sans Pro")) %>%
  opt_row_striping(row_striping = FALSE) %>%
  tab_options(
    table.width                       = px(720),
    table.border.top.color            = "white",
    table.border.bottom.color         = "white",
    heading.border.bottom.color       = "#333333",
    row_group.border.top.color        = "#cccccc",
    row_group.border.bottom.color     = "#cccccc",
    column_labels.border.top.color    = "#333333", 
    column_labels.border.bottom.color = "#333333",
    data_row.padding                  = px(6),
    source_notes.font.size            = 11,
    heading.title.font.size           = 14,
    heading.subtitle.font.size        = 11
  )

tbl_dir
gtsave(tbl_dir, "./wetting_drying_acc.png")

```




