---
title: "JustProcessingStage"
format: html
editor_options: 
  chunk_output_type: console
---

11-17-25
Markdown just for processing stage discharge measurements, and preparing discharge datasets for FB and ZZ. Because it is in one place I can just create a finalized .csv for each watershed, and remove the stage/discharge processing part of other scripts.

```{r setup}
library(pacman)
p_load(tidyverse)
```

# Process salt dilutions
Use previously created script with functions for processing salt dilutions.
```{r process-dilutions}
source("calcQ.R")
q_combined <- rbind(
calcQ("./salt_dilutions/FB/FB_8_1") %>% mutate("shed" = "FB"),
# calcQ_noK("./salt_dilutions/FB/FB_8_1", mean_k)%>% mutate("shed" = "FB"),

#calcQ("./salt_dilutions/FB/FB_6_23_fail") %>% mutate("shed" = "FB"),
calcQ("./salt_dilutions/ZZ/ZZ_8_1")%>% mutate("shed" = "ZZ"),
calcQ("./salt_dilutions/ZZ/ZZ_7_1")%>% mutate("shed" = "ZZ"),
#calcQ("./salt_dilutions/ZZ/ZZ_6_23") %>% mutate("shed" = "ZZ"),#different timestep, after fixing gave negative number
calcQ("./salt_dilutions/ZZ/ZZ_6_21")%>% mutate("shed" = "ZZ"),

#no k
calcQ_noK("./salt_dilutions/FB/FB_7_1", mean(c(0.0001120716, 2.621858e-06))) %>% mutate("shed" = "FB"),
calcQ_noK("./salt_dilutions/FB/FB_6_23_Success", mean_k) %>% mutate("shed" = "FB"),

#missing FB 8_5, no k and no upstream logger. The k used is not really based on anything I think...
# ZZ_8_5, no k and no upstream logger
calcQ_noK_oneL("./salt_dilutions/FB/FB_8_5", 0.0002120716) %>% mutate("shed" = "FB"),
calcQ_noK_oneL("./salt_dilutions/ZZ/ZZ_8_5", mean(c(9.453437e-05, 4.017994e-06,
                                                    3.962093e-05, 9.693616e-05)))%>% mutate("shed" = "ZZ")
)
```

# Falls Brook

Potential big problem with my pressure transducer measurements. At some point during summer of 2024, the pressure changes really dramatically, like increases by one order of magnitude. It remains this way in 2025 as well. I think it affects the salt dilutions that Logan conducted.
```{r read-csv}
#### 2023 ####
FB_air23 <- read_csv("./PressureTransducers_11_14_23/FB_air.csv", 
                     skip = 1) %>% 
  select(2:3) %>% 
  rename(DATETIME = 1,
         pressure_psi_air = 2) %>% 
  mutate(DATETIME = mdy_hms(DATETIME))

FB_water23 <- read_csv("./PressureTransducers_11_14_23/FB_water.csv", 
                       skip = 1) %>% 
  select(2:3) %>% 
  rename(DATETIME = 1,
         pressure_psi_water = 2) %>% 
  mutate(DATETIME = mdy_hms(DATETIME))

FB_water25 %>% 
  left_join(FB_air25, by = "DATETIME") %>%
  #filter(DATETIME >= ymd_hms("2024-07-30 14:00:00 UTC")) %>% 
  mutate(diff_psi = pressure_psi_water - pressure_psi_air) %>%
  #mutate(stage_cm = ((diff_psi*6894.76) / (997 * 9.8)) * 100 * 0.393701) %>%
  ggplot(aes(x = DATETIME))+
  geom_line(aes(y = diff_psi), col = "blue")+
    #geom_line(aes(y = pressure_psi_air))+

  theme_classic()+
  labs(title = "Falls Brook Stage",
       x = "",
       y = "Stage (in)")

#exclude windows where they were not deployed
FB_water %>%
  left_join(FB_air, by = "DATETIME") %>%
  mutate(diff_psi = pressure_psi_water - pressure_psi_air) %>%
  mutate(stage_cm = ((diff_psi*6894.76) / (997 * 9.8)) * 100 * 0.393701) %>%
  filter(stage_cm < 5)

#7/22/2023 00:00:00
#9/20/23 00 - 2023-09-22 00:00:00
#2023-11-12 00:30:00

# FB_water %>% 
#   left_join(FB_air, by = "DATETIME") %>% 
#   mutate(diff_psi = pressure_psi_water - pressure_psi_air) %>% 
#   mutate(stage_cm = ((diff_psi*6894.76) / (997 * 9.8)) * 100 * 0.393701) %>% 
#   filter(DATETIME > ymd_hms("2023-07-22 00:00:00"),
#          DATETIME < ymd_hms("2023-11-12 00:00:00")) %>% 
#   filter(!(DATETIME >= ymd_hms("2023-09-20 00:00:00") & 
#           DATETIME <= ymd_hms("2023-09-22 00:00:00"))) %>% 
#   mutate(minutes = minute(DATETIME)) %>% 
#   filter(minutes %in% c(0, 30)) %>% 
#   mutate(roll_mean = rollapply(stage_cm ,48,mean,align='center',fill=NA)) %>%
#   ggplot()+
#   geom_line(aes(x = DATETIME, y = stage_cm))+
#     geom_line(aes(x = DATETIME, y = roll_mean), color = "grey")+
#   theme_classic()+
#   labs(title = "Falls Brook Stage",
#        x = "",
#        y = "Stage (in)")

processed_fb_stage23 <- FB_water23 %>% 
  left_join(FB_air23, by = "DATETIME") %>% 
  mutate(diff_psi = pressure_psi_water - pressure_psi_air,
         stage_cm = ((diff_psi*6894.76) / (997 * 9.8)) * 100 * 0.393701,
         minutes = minute(DATETIME)) %>% 
  filter(DATETIME > ymd_hms("2023-07-22 00:00:00"),
         DATETIME < ymd_hms("2023-11-12 00:00:00")) %>% 
  filter(!(DATETIME >= ymd_hms("2023-09-20 00:00:00") & 
          DATETIME <= ymd_hms("2023-09-22 00:00:00"))) %>% 
  filter(minutes %in% c(0, 30)) %>% 
  mutate(roll_mean = rollapply(stage_cm ,48,mean,align='center',fill=NA),
         deployment = 23) 


processed_fb_stage23 %>% 
  ggplot(aes(x = DATETIME, y = roll_mean))+
  geom_line()+
  theme_classic()
#output processed salt dilutions
# q_combined %>% filter(shed == "FB") %>% 
#   mutate(datetime = round_date(datetime, unit = minutes(30))) %>% 
#   rename(DATETIME = datetime) %>% 
#   left_join(processed_fb_stage, by = "DATETIME")

#### 2024 ####
FB_air24 <- read_csv("./PressureTransducers_summer24/FB_air.csv", skip = 1) %>% 
  select(2:3) %>% 
  rename(DATETIME = 1,
         pressure_psi_air = 2) %>% 
  mutate(DATETIME = mdy_hms(DATETIME)) %>% 
  rbind(fb_air_25)

#FB_water24 <- read_csv("./PressureTransducers_summer24/FB_water.csv", skip = 1) %>% 

FB_water24 <- read_csv("./PressureTransducers_summer24/FB_water.csv", skip = 1) %>% 
  select(2:3) %>% 
  rename(DATETIME = 1,
         pressure_psi_water = 2) %>% 
  mutate(DATETIME = mdy_hms(DATETIME)) %>% 
  rbind(fb_water_25)

processed_fb_stage24 <- FB_water24 %>% 
  left_join(FB_air24, by = "DATETIME") %>% 
  mutate(diff_psi = pressure_psi_water - pressure_psi_air) %>% 
  mutate(stage_cm = ((diff_psi*6894.76) / (997 * 9.8)) * 100 * 0.393701) %>% 
    filter(DATETIME <= ymd_hms("2024-07-20 16:00:00 UTC")) %>% 
  mutate(minutes = minute(DATETIME)) %>% 
  filter(minutes %in% c(0, 30)) %>% 
  mutate(roll_mean = rollapply(stage_cm ,48,mean,align='center',fill=NA),
         deployment = 24) %>%
  drop_na()

processed_fb_stage24 %>% 
  ggplot(aes(x = DATETIME, y = roll_mean))+
  geom_line()+
  theme_classic()

#max(processed_fb_stage_24$DATETIME)
#### 2025 ####
#in 2025 units are kPa, not psi
FB_air25 <- read_csv("./pducers_summer25/fb_air.csv", skip = 1) %>% 
  select(2:3) %>% 
  rename(DATETIME = 1,
         pressure_psi_air = 2) %>% 
  mutate(DATETIME = mdy_hms(DATETIME))

FB_water25 <- read_csv("./pducers_summer25/fb_water.csv", skip = 1) %>% 
  select(2:3) %>% 
  rename(DATETIME = 1,
         pressure_psi_water = 2) %>% 
  mutate(DATETIME = mdy_hms(DATETIME))

processed_fb_stage25 <- FB_water25 %>% 
  left_join(FB_air25, by = "DATETIME") %>% 
  mutate(diff_psi = pressure_psi_water - pressure_psi_air) %>% 
  mutate(stage_cm = ((diff_psi*1000) / (997 * 9.8)) * 100 * 0.393701) %>% 
    filter(DATETIME >= ymd_hms("2024-07-20 04:00:00 UTC"),
           DATETIME <= ymd_hms("2024-08-10 19:30:00 UTC")) %>% 
  mutate(minutes = minute(DATETIME)) %>% 
  filter(minutes %in% c(0, 30)) %>% 
  mutate(roll_mean = rollapply(stage_cm ,24,mean,align='center',fill=NA),
         deployment = 25) %>%
  drop_na()

#### Combine all years ####
fb_all <- rbind(processed_fb_stage25, processed_fb_stage24, processed_fb_stage23) %>% 
  rename("datetime" = DATETIME) %>% 
  drop_na()

processed_fb_stage25 %>% 
  ggplot(aes(x = DATETIME, y = roll_mean))+
  geom_line()+
  theme_classic()

fb_all %>% 
  ggplot(aes(x = datetime))+
    geom_point(aes(y = stage_cm))+
  geom_line(aes(y = roll_mean), color = "grey")+
  theme_classic()

```
```{r fit-stage-discharge}

fb_all %>% 
  filter(datetime == "2024-08-01 11:00:00 UTC")
# processed salt dilutions
dilutionsFB <- 
  q_combined %>% filter(shed == "FB") %>% 
  mutate(datetime = ymd_hms(datetime),
         datetime = round_date(datetime, unit = minutes(30))) %>% 
  #rename(DATETIME = datetime) %>% 
  left_join(fb_all, by = "datetime")
dilutionsFB

plot((dilutionsFB$roll_mean), (dilutionsFB$discharge))

# fitting linear equation
fb_stage_Q_model <- lm(dilutions$discharge ~ dilutions$roll_mean)
summary(fb_stage_Q_model)
#try exponential model
model_exp <- lm(log(dilutionsFB$discharge)~ log(dilutionsFB$roll_mean))
summary(model_exp)

#mutate(Q = roll_mean * fb_stage_Q_model$coefficients[2] +  fb_stage_Q_model$coefficients[1])

# applying exponential equation
# L/s discharge units
fb_all %>% 
  mutate(Q = exp(log(roll_mean) * model_exp$coefficients[2] +  model_exp$coefficients[1])) %>%
# convert Q to mm/day
  ggplot(aes(x = datetime, y = Q))+
  geom_line()

### Convert L/s to mm/day
LPers_to_mmPerDay <- function(Q, area){
  # anticipating L/s, watershed area in km^2
  Q1 <- Q * 86400 # convert days to seconds
  Q2 <- Q1 * 0.001 # convert L to m^3
  Q3 <- Q2 * 1E9 # convert m^3 to mm^3
  
  area_mm2 <- area * 1E12 # convert area to mm^2
  Q4 <- Q3 / area_mm2 # divide mm^3/day by area mm^2 for mm/day
  return(Q4)
}
# discharge not really the same order of magnitude...

fb_all %>% 
  mutate(Q = exp(log(roll_mean) * model_exp$coefficients[2] +  model_exp$coefficients[1])) %>%
  mutate(FB_Q_mm_day = LPers_to_mmPerDay(Q, area_fb)) %>% 
inner_join(w3_limbs) %>% 
  select(datetime, FB_Q_mm_day, Q_mm_day) %>% 
  ggplot(aes(x = Q_mm_day, y = FB_Q_mm_day))+
  geom_point()

# plot of discharge
discharge_df$event_type[is.na(discharge_df$event_type)] <- "baseflow"

stage_df_24 <- discharge_df

discharge_df %>% 
  ggplot(aes(x = DATETIME, y = roll_mean, color = event_type))+
  geom_point()

```

```{r identify-events}
#identifying events
bf <- baseflowB(fb_all$roll_mean, alpha = 0.99)


#subtract baseflow from discharge
#PoT_res = eventPOT(q_23_plotting$Q_mm_day - bf$bf, threshold = 1, min.diff = 1)

PoT_res <- eventPOT(fb_all$roll_mean - bf$bf, threshold = 0.4, min.diff = 20)
#plot the events
plotEvents(data = fb_all$roll_mean, 
           events = PoT_res, 
           xlab = "Index", 
           ylab = "Flow (ML/day)", 
           colpnt = "#E41A1C", 
           colline = "#377EB8", main = "eventPOT")

limbs(data = fb_all$roll_mean, 
               dates =NULL, 
               events = PoT_res, 
               to.plot = TRUE)#now, extract just the start column, then get each window and run through scoring algorithm


ranges <- data.frame("starts" = PoT_res$srt) %>% 
  mutate("stops" = lead(starts)) %>% 
  mutate(event_ID = row_number(starts))


# Example: discharge dataframe with datetime and discharge
discharge_df <- processed_fb_stage

# Copy of your event table
events <- limbs(data = processed_fb_stage$roll_mean, 
               dates =NULL, 
               events = PoT_res, 
               to.plot = FALSE)#now, extract just the start column, then get each window and run through scoring algorithm

# Initialize the event_type column
discharge_df$event_type <- NA_character_
discharge_df$event_id <- NA_integer_

# Assign "rising" and "falling" from event definitions
for (i in seq_len(nrow(events))) {
  ev <- events[i, ]
  discharge_df$event_id[ev$ris.srt:ev$fal.end] <- i  # Event ID for rising+falling
  
  # Rising limb
  if (!is.na(ev$ris.srt) && !is.na(ev$ris.end) && ev$ris.srt <= ev$ris.end) {
    discharge_df$event_type[ev$ris.srt:ev$ris.end] <- "rising"
  }
  
  # Falling limb
  if (!is.na(ev$fal.srt) && !is.na(ev$fal.end) && ev$fal.srt <= ev$fal.end) {
    discharge_df$event_type[ev$fal.srt:ev$fal.end] <- "falling"
  }
  
  # Baseflow between events
  if (i < nrow(events)) {
    this_fal_end <- ev$fal.end
    next_ris_srt <- events$ris.srt[i + 1]
    if (this_fal_end + 1 <= next_ris_srt - 1) {
      baseflow_idx <- (this_fal_end + 1):(next_ris_srt - 1)
      discharge_df$event_type[baseflow_idx] <- "baseflow"
      discharge_df$event_id[baseflow_idx] <- NA  # baseflow is not part of any event
    }
  }
}

# Optional: label all remaining NA values as "baseflow" (e.g., before first or after last event)
discharge_df$event_type[is.na(discharge_df$event_type)] <- "baseflow"

stage_df_23 <- discharge_df

discharge_df %>% 
  ggplot(aes(x = DATETIME, y = roll_mean, color = event_type))+
  geom_point()


```


# Zig Zag

Zig zag has the same problem, but only for the 2025 observations.
```{r read-csv}
#### 2023 ####
ZZ_air23 <- read_csv("./PressureTransducers_11_14_23/ZZ_air.csv", skip = 1) %>% 
  select(2:3) %>% 
  rename(DATETIME = 1,
         pressure_psi_air = 2) %>% 
  mutate(DATETIME = mdy_hms(DATETIME))

ZZ_water23 <- read_csv("./PressureTransducers_11_14_23/ZZ_water.csv", skip = 1) %>% 
  select(2:3) %>% 
  rename(DATETIME = 1,
         pressure_psi_water = 2) %>% 
  mutate(DATETIME = mdy_hms(DATETIME))

processed_zz_stage23 <- ZZ_water23 %>% 
  left_join(ZZ_air23, by = "DATETIME") %>% 
  mutate(diff_psi = pressure_psi_water - pressure_psi_air) %>% 
  mutate(stage_cm = ((diff_psi*6894.76) / (997 * 9.8)) * 100 * 0.393701) %>%
  #filter(DATETIME <= ymd_hms("2024-07-21 10:30:00 UTC")) %>% 
   filter(DATETIME > ymd_hms("2023-07-22 00:00:00"),
         DATETIME < ymd_hms("2023-11-12 00:00:00")) %>% 
  filter(!(DATETIME >= ymd_hms("2023-09-20 00:00:00") & 
          DATETIME <= ymd_hms("2023-09-22 00:00:00"))) %>% 
  mutate(minutes = minute(DATETIME)) %>% 
  filter(minutes %in% c(0, 30)) %>% 
  mutate(roll_mean = rollapply(stage_cm ,48,mean,align='center',fill=NA)) %>% 
  drop_na()


#### 2024 ####
ZZ_air24 <- read_csv("./PressureTransducers_summer24/ZZ_air.csv", skip = 1) %>% 
  select(2:3) %>% 
  rename(DATETIME = 1,
         pressure_psi_air = 2) %>% 
  mutate(DATETIME = mdy_hms(DATETIME))

ZZ_water24 <- read_csv("./PressureTransducers_summer24/ZZ_water.csv", skip = 1) %>% 
  select(2:3) %>% 
  rename(DATETIME = 1,
         pressure_psi_water = 2) %>% 
  mutate(DATETIME = mdy_hms(DATETIME))


processed_zz_stage24 <- ZZ_water24 %>% 
  left_join(ZZ_air24, by = "DATETIME") %>% 
  mutate(diff_psi = pressure_psi_water - pressure_psi_air) %>% 
  mutate(stage_cm = ((diff_psi*6894.76) / (997 * 9.8)) * 100 * 0.393701) %>%
  filter(DATETIME <= ymd_hms("2024-07-21 10:30:00 UTC")) %>% 
  mutate(minutes = minute(DATETIME)) %>% 
  filter(minutes %in% c(0, 30)) %>% 
  mutate(roll_mean = rollapply(stage_cm ,24,mean,align='center',fill=NA)) %>% 
  drop_na()

processed_zz_stage24 %>% 
  ggplot(aes(x = DATETIME))+
    geom_point(aes(y = stage_cm))+
  geom_line(aes(y = roll_mean), color = "grey")+
  theme_classic()
#### 2025 ####
# in 2025, units are kPa, not psi
ZZ_air25 <-
  rbind(read_csv("./pducers_summer25/zz_air.csv", skip = 1) %>% 
  select(2:3) %>% 
  rename(DATETIME = 1,
         pressure_psi_air = 2) %>% 
  mutate(DATETIME = mdy_hms(DATETIME)))

ZZ_water25 <- 
  read_csv("./pducers_summer25/zz_water.csv", skip = 1) %>% 
  select(2:3) %>% 
  rename(DATETIME = 1,
         pressure_psi_water = 2) %>% 
  mutate(DATETIME = mdy_hms(DATETIME))

     
processed_zz_stage25 <- ZZ_water25 %>% 
  left_join(ZZ_air25, by = "DATETIME") %>% 
  mutate(diff_psi = pressure_psi_water - pressure_psi_air) %>% 
  mutate(stage_cm = ((diff_psi*1000) / (997 * 9.8)) * 100 * 0.393701) %>% 
  filter(DATETIME >= ymd_hms(max(processed_zz_stage24$DATETIME)),
         DATETIME <= ymd_hms("2024-08-10 19:30:00 UTC")) %>%
  mutate(minutes = minute(DATETIME)) %>% 
  filter(minutes %in% c(0, 30)) %>% 
  mutate(roll_mean = rollapply(stage_cm ,24,mean,align='center',fill=NA)) %>% 
  drop_na()

#### combined ####
#View(processed_zz_stage)
rbind(processed_zz_stage23,
      processed_zz_stage24,
      processed_zz_stage25) %>%
  ggplot()+
    geom_line(aes(x = DATETIME, y = stage_cm))+
    geom_line(aes(x = DATETIME, y = roll_mean), color = "grey")+  theme_classic()+
  labs(title = "ZZ Stage",
       x = "",
       y = "Stage (in)")
```

```{r fit-stage-discharge}
#Read in measurements from 2024

# processed salt dilutions
dilutionsZZ <- q_combined %>% filter(shed == "ZZ") %>% 
  mutate(datetime = round_date(datetime, unit = minutes(30))) %>% 
  rename(DATETIME = datetime) %>% 
  left_join(processed_zz_stage, by = "DATETIME")
dilutionsZZ

ZZ_water %>% 
  left_join(ZZ_air, by = "DATETIME") %>% 
  mutate(diff_psi = pressure_psi_water - pressure_psi_air) %>% 
  mutate(stage_cm = ((diff_psi*6894.76) / (997 * 9.8)) * 100 * 0.393701,
         #apply a rolling mean to get rid of temperature artifacts
         roll_mean = rollapply(stage_cm ,48,mean,align='center',fill=NA)) %>% 
  ggplot()+
  geom_line(aes(x = DATETIME, y = stage_cm))+
    geom_line(aes(x = DATETIME, y = roll_mean), color = "grey")+
  
  theme_classic()+
  labs(title = "ZZ Stage",
       x = "",
       y = "Stage (in)")+
  geom_point(data = dilutions,
             aes(x = DATETIME, y = stage_cm),
             color = "blue")

dilutions %>% 
  ggplot(aes(x = roll_mean, y = discharge))+
  geom_point()+
  labs(x = "Stage (cm)", y = "Discharge (L/s)",
       title = "ZZ Stage/Q")+
  theme_classic()+
  geom_smooth(method = "lm")

exponential.model <- lm(log(dilutions$discharge)~ dilutions$roll_mean)
summary(exponential.model)

linear <- lm((dilutions$discharge)~ dilutions$roll_mean)
summary(linear)


#now apply model to convert to discharge across time
processed_zz_stage %>% 
  mutate(prediction = (roll_mean^exponential.model$coefficients[2])-exponential.model$coefficients[1]) %>% 
  ggplot(aes(x = DATETIME, y = prediction))+
  geom_line()
  
```