---
title: "Lab 6 R4Geo"
output:
  html_document: default
  pdf_document: default
date: "2022-11-19"
---
John Morgan

For my research, I am planning on utilizing the hydrological model Topmodel to explore the impact of different soil units on streamflow in the Hubbard Brook experimental forest in New Hampshire.

There are two R package implementations of Topmodel as packages, "topmodel" and "dynatop". Neither is exactly how I want to implement it, so at this time I am planning on making my own implementation, and comparing it to these pre-existing ones for my final project. 

For this assignment, I am attempting to run a very basic iteration of the "topmodel" implementation. First, I will make a map of the study watershed, then I will input parameters and run the model.
  

```{r}
#setup
library(pacman)

p_load(dynatopGIS, tidyverse, raster, rgdal, sp, mapview,
       rgeos, spdplyr, whitebox, RColorBrewer, topmodel, lubridate)
```
  
Below is a map of my study watershed, showing field mapped stream flowlines for Paradise Brook, and a hillshade generated from a 1 meter DEM.
```{r}
#reading in dem and watershed shapefiles
setwd("C:/Users/John/Documents/HBTopModel/HB/knb-lter-hbr.211.2")
dem <- raster("dem1m.tif")
#plot(dem)

setwd("C:/Users/John/Documents/HBTopModel/HB/hbef_wsheds")
w3 <- readOGR(dsn='.',layer = 'hbef_wsheds')
wall <- w3

w3 <- w3 %>% filter(WS == "WS3")
#plot(w3)

w3_p <- spTransform(w3, crs(dem))
dem_crop <- crop(dem, w3_p)
w3_dem <- mask(dem_crop, w3_p)
#raster::writeRaster(w3_dem, "w3_dem.tif", format = "GTiff")

#stream shapefile, from Jensen et al. 2017
setwd("C:/Users/John/Documents/HBTopModel/HB/hbstream/HB_Shapefiles")
stream <- readOGR("hb42_master.shp")
s <- raster::shapefile("hb42_master.shp")
stream_p <- spTransform(stream, crs(dem))

#make a hillshade
wbt_hillshade(dem = "w3_dem.tif",
              output = "w3_hillshade.tif",
              azimuth = 115)

hill <- raster("w3_hillshade.tif")

#mapview(hill, col.regions = brewer.pal(6, "Greys"))+
#  mapview(stream_p)

#make a nice map of the HB LTER, highlighting watershed 3
ggplot()

```
  
Below I calculate the TWI using one of the implementations of topmodel, the "topmodel" package.
```{r, warning=FALSE}
#generating topographic index

#convert DEM to matrix
mat <- raster::as.matrix(dem_crop)
mat[mat==-9999] <- NA
#already generated topidx table before
topindex <- topidx(mat, resolution=1)
#write.table(topindex, 'HB_topindex.txt')

#topidx <- read.csv("C:/Users/John/Documents/HBTopModel/HB_topindex.txt", sep = " ")
topidx1 <- make.classes(topindex$atb,16)
topidx2 <- make.classes(topindex$area,16)

# the delay function is a bit more tricky because this requires cumulative fractions, but you generate it as follows:

n <- 5 # number of classes; a higher number will result in a smoother histogram
which(mat == min(mat), arr.ind = TRUE)

delay <- flowlength(mat, c(837, 176))*1 #TOD: add the outlet coordinates; otherwise the flowlength will be calculated to the edge of the map.
delay <- make.classes(delay, n)
delay <- delay[n:1,]
delay[,2] <- c(0, cumsum(delay[1:(n-1),2]))
```
  
And here I attempt to implement the model for the first time:

```{r}
data(huagrahuma)

h <- huagrahuma
#using a combination of variables from example data and HB
#precip
rain <- read.csv("HB/dailyWatershedPrecip1956-2021.csv")
rain$DATE <- ymd(rain$DATE)
rain2 <- filter(rain, watershed == "W3") %>% 
  filter(DATE > ymd("2020-01-01") & DATE < ymd("2021-01-01"))

#actual observed discharge
q <- read.csv("HB/w3_stmflow_2013-2022_5min.csv")
q$DATETIME <- ymd_hms(q$DATETIME)
q2 <- q %>%
  mutate(date = floor_date(DATETIME, unit = "day")) %>%
  group_by(date) %>%
  summarise(
    n = n(),
    q_total = sum(Discharge_ls),
    q_average = mean(Discharge_ls),
    .groups = "drop"
  ) %>% 
  filter(date > ymd("2020-01-01") & date < ymd("2021-01-01"))

#generating PET data
#lazy bad method: normal distribution of PET through year
y <- rnorm(seq(1, 365, 1), mean = 0.00226, sd = 1)

test <- topmodel(h$parameters, topidx1, h$delay, rain2$Precip, y, verbose = F, Qobs = NA)

#plot(test, type="l", col="red", ylim = c(0, 30))
#points(q2$q_average, type = 'l')

final <- data.frame("Date" = q2$date,
                       "Actual_Q" = q2$q_average,
                       "Modeled_Q" = test)

final2 <- gather(final, key = "key", value = "value", -Date)
ggplot(final2)+
  geom_line(aes(x = Date, y = value, color = key))+
  theme_classic()+
  labs(y = "Discharge (L/s)",
       title = "Simulated discharge using topmodel for Paradise Brook, 2020")
```
  
Based on the model results, the simulated discharge was not very close at all. It failed to capture the magnitude of the peaks observed. In future attempts, I will use a more robust method to estimate or parameterize ET.
