---
title: "QPM_analysis_9_29_23"
author: "John Morgan"
date: "2023-09-29"
output: html_document
---
After meeting with Kevin and JP, I am doing an analysis that will be more useful for everyone else at Hubbard Brook. Things I need to do:
- Define a normal channel for each watershed
  - Start with W3, use known extent to develop an area threshold, then use that on other watersheds
- Determine how frequently streams connect with biogeochemical hotspots (rocky areas, shallow soils)
  - Use BOSS to get rocky areas and shallow soils
  - How often does the channel expand up into these areas?
  - What stream expansion threshold connects with hotspots?
- How much does stream area change during the growing season?
  - Map of stream channel, color scale by log(days flowing)
  - What area of stream is flowing over time? How does this translate to biogeochemical fluxes (CO2) from the flowing surface of water?
    - Find CO2 fluxes in NE headwater streams
- The streams are dynamic in the middle ranges- how do I express this? (Less important than others)

```{r}
#defining a normal channel

```

```{r}
#find where expanded channels reach into rocky outcrops and shallow soil

```


